### 2025-03-26 analysis.Rmd
save diagnostics

---
title: "Lapse & Vision Loss PS Analysis"
author: "Erik Westlund"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("settings.R")

library(cobalt)
library(dplyr)
library(gee)
library(ggplot2)
library(kableExtra)
library(margins)
library(MatchIt)
library(purrr)
library(readr)
library(survey)
library(twang)
library(WeightIt)

shouldCache <- TRUE

data <- readr::read_csv(file.path(s_root, "Gina's Project/df_Final_240909.csv")) |> 
  mutate(
    outcome_va_vi = outcome_VA_logMAR >= 0.3,
    outcome_va_blind = outcome_VA_logMAR >= 1,
    outcome_va_vi_binary = if_else(outcome_va_vi, 1, 0),
    outcome_va_blind_binary = if_else(outcome_va_blind, 1, 0),
    ever_lapse_cat = factor(person_ever_lapse),
    ever_lapse_binary = if_else(ever_lapse_cat == TRUE, 1, 0),
    MRN = dplyr::min_rank(e_mrn_deidentified),
    gender_cat = factor(gender),
    race_ethnic_cat = factor(race_ethnicity_gp),
    insurance_cat = factor(insurance_gp),
    age_cat = cut(age, breaks = c(0, 20, 45, 65, Inf), laebsl = c("0-20", "21-45", "46-65", "65+")),
    person_dr = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 1,
      NPDR == "Severe" ~ 1,
      PDR == "Present" ~ 2,
    ),
    person_dr_severity = case_when(
      No_DR == TRUE ~ 0,
      NPDR == "Mild" ~ 1,
      NPDR == "Moderate" ~ 2,
      NPDR == "Severe" ~ 3,
      PDR == "Present" ~ 4,
    ),
    person_ever_treat = case_when(
      other_inject == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & PRP_flag == 0 ~ FALSE,
      other_inject == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | PRP_flag == 1 ~ TRUE
    ),
    glaucoma_bef_hitplus_cat = case_when(
      Glaucoma_bef_hitplus != "Present" | is.na(Glaucoma_bef_hitplus) ~ FALSE,
      Glaucoma_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Glaucoma_after_hitplus != "Present" | is.na(Glaucoma_after_hitplus) ~ FALSE,
      Glaucoma_after_hitplus == "Present" ~ TRUE
    ),
    otherretina_bef_hitplus_cat = case_when(
      Otherretina_bef_hitplus != "Present" | is.na(Otherretina_bef_hitplus) ~ FALSE,
      Otherretina_bef_hitplus == "Present" ~ TRUE
    ),
    glaucoma_after_hitplus_cat = case_when(
      Otherretina_after_hitplus != "Present" | is.na(Otherretina_after_hitplus) ~ FALSE,
      Otherretina_after_hitplus == "Present" ~ TRUE
    ),
    other_inject = factor(other_inject),
    anti_VEGF = factor(anti_VEGF),
    focal_laser_flag = factor(focal_laser_flag),
    PRP_flag = factor(PRP_flag),
    glaucoma_bef_hitplus_cat = factor(glaucoma_bef_hitplus_cat),
    otherretina_bef_hitplus_cat = factor(otherretina_bef_hitplus_cat),
    catsurg_before_hitplus_cat = factor(cataract_surgery_bef_hitplus)
  )

save_og_image <- function(plot, filename, width=800, height=1000) {
  png(filename=filename, width=width, height=height)
  plot
  dev.off()
}

```

## Introduction

* `Treatment`: We consider here the "treatment" to be having a lapsed episode within our observation window.
* `Outcome`: We want to estimate the effect of the treatment on vision loss, measured separately as vision impairment and blindness.
* `Estimand`: We will estimate the ATT, or average treatment effect in the treated.

In summary, we will attempt to estimate the effect of lapsing vs. not lapsing among those who lapsed. Put differently, one could think of this as the impact of attending clinic, counter to fact (lapsing), on vision impairment.

## Process

We will use propensity score matching techniques to estimate the probability of treatment (i.e., the probability of lapsing), given a vector of covariates we think predicts lapsing. These covariates must be measured before treatment. We will then use these probabilities, or propensity scores, to construct a cohort of patients assembled or weighted such that it allows for the estimation of the desired estimand (here, the `ATT`).

Since we will be estimating the `ATT`, all patients who lapsed will be kept in the study. We will use various techniques to choose a matched cohort of patients, with the end goal of matching lapsers with patients who do not lapse, despite having a similar propensity to do so.

## Issues

10 People are missing CCI and DCSI:

```{r cci}
missing <- data |> 
  select(e_mrn_deidentified, CCI, DCSI) |> 
  filter(is.na(CCI) | is.na(DCSI))

missing |> count()
missing |> kable()

```

We will proceed with a data set that drops those patients.

```{r get_m_data}
m_data <- data |> filter(
  !is.na(!CCI) & !is.na(DCSI)
)

m_data |> count()
```

We will model `p(lapse)` using this formula:

```{r init_ps, echo = FALSE}
matchingFormula <- as.formula("ever_lapse_binary ~ 
  baseline_VA_logMAR +
  gender_cat +
  race_ethnic_cat +
  insurance_cat +
  age_cat +
  CCI +
  DCSI +
  other_inject +
  anti_VEGF +
  focal_laser_flag +
  PRP_flag +
  glaucoma_bef_hitplus_cat +
  otherretina_bef_hitplus_cat +
  catsurg_before_hitplus_cat")

print(matchingFormula)

```

## Initial Imbalance

We first want to assess the imbalance before weighting or matching to get a sense for how imbalanced our sample is.

```{r initial_imbalance}
m.initialImbalance <- matchit(
    formula = matchingFormula,
    data = m_data,
    method = NULL,
    distance = "glm" 
  )
```

```{r initial_imablance_img, echo=FALSE}

save_og_image(
  plot = plot(summary(m.initialImbalance), main="Summary of Initial Imbalance", sub="distance = p(lapse)"),
  filename = "images/m.initialImbalance_summary.png",
  width = 800,
  height = 1000
)
```

![Initial Imbalance](images/m.initialImbalance_summary.png)

A rule of thumb is <0.1 SMD is good, < 0.15 is acceptable, and over that is problematic.

Accounting:

* 25 measures
* 13 > 0.1

## Propensity Score Construction

We will use a function that lets us run a number of different propensity score designs.

All propensity score models use a binomial family with a logit link. Matching is done with replacement because we have more treated than control units.

* `nearest_glm`: Nearest neighbor using propensity scores generated with GLM (generalized linear model)
* `nearest_gam`: Nearest neighbor using propensity scores generated with a GAM (generalized additive model)
* `nearest_gbm`: Nearest neighbor using propensity scores generated with GBM (generalized boosted model)
* `nearest_lasso`: Nearest neighbor using propensity scores generated with lasso regression
* `nearest_rpart`: Nearest neighbor using propensity scores generated with regression trees
* `optimal_glm`: Optimal matching is similar to nearest but optimizes on the absolute pair distance
* `subclass_glm`: Uses N number of bins (set to 5) to create strata and weights
* `nearest_mahalanobis`: Matches using the mahalanobis distance between the set of covariates.
* `genetic`: Nearest neighbor matching using a genetic algorithm to generate a distance scores.
* `coarsened`: Coarsens the covariates into bins and uses exact matching afterwards.
* `twang_gbm`: Use weighting with GBM (gradient boosting model)

```{r ps_f, echo=FALSE}

# Define function to run matching with disk storage and selective re-running
run_matching_methods <- function(data, formula, methods, output_dir = "ps_results") {
  if (!dir.exists(output_dir)) {
    dir.create(output_dir, recursive = TRUE)
  }

  all_results <- list()

  for (params in methods) {
    filename <- file.path(output_dir, paste0(params$name, ".rds"))
    diagnostics_filename <- file.path(output_dir, paste0(params$name, "_diagnostics.rds"))

    print(paste0("Starting method: ", params$name, "."))

    result <- NULL

    if (file.exists(filename) && !params$rerun_model) {
      print("Loading cached results...")
      result <- readRDS(filename)$result
    }

    if (is.null(result)) {
      print("No cached results found or rerunning model. Running method...")

      result <- tryCatch({
        if (params$method == "twang") {
          ps(formula = formula, data = data.frame(data), 
             n.trees = 10000, interaction.depth = 3,
             shrinkage = 0.01, estimand = params$estimand,
             stop.method = c("es.mean", "ks.max"))
        } else if (!is.null(params$use) && params$use == "weightit") {
          do.call(weightit, c(list(formula = formula, data = data), params))
        } else {
          do.call(matchit, c(list(formula = formula, data = data), params))
        }
      }, error = function(e) {
        message("Error in ", params$name, ": ", e)
        return(NULL)
      })

      if (!is.null(result)) {
        saveRDS(list(result = result), file = filename)
      }
    }

    if (is.null(result)) {
      next
    }

    love_plot <- NULL
    balance_plot <- NULL
    balance_summary <- NULL

    if (params$rerun_diagnostics || !file.exists(diagnostics_filename)) {
      print("Running diagnostics...")

      love_plot <- if (params$method != "twang") { 
        tryCatch({
          love.plot(result, stat = "mean.diffs", var.order = "adjusted", stars = "std")
        }, error = function(e) {
          message("Error generating balance plot for ", params$name, ": ", e)
          return(NULL)
        })
      } else if(params$method == "twang") {
       tryCatch({
          plot(result, plots=3)
        }, error = function(e) {
          message("Error generating balance plot for ", params$name, ": ", e)
          return(NULL)
        })
      } else {
        NULL
      }

      balance_plot <- tryCatch({
        if (params$method == "twang") {
          bal.plot(result, var.name = "prop.score.es.mean", which = "both")
        } else if("distance" %in% names(result)) {
          bal.plot(result, var.name = "distance", which = "both")
        } else {
          NULL
        }
      }, error = function(e) {
        message("Error generating common support plot for ", params$name, ": ", e)
        return(NULL)
      })

      p_dist_plot <- tryCatch({
        if (params$method == "twang") {
          plot(result, plots=2)
        } else if("distance" %in% names(result)) {
          plot(result, type = "jitter", interactive=FALSE)
        } else {
          NULL
        }
      }, error = function(e) {
        message("Error generating p-score distribution support plot for ", params$name, ": ", e)
        return(NULL)
      })

      balance_summary <- tryCatch({
        if (params$method == "twang") {
          summary(result)
        } else {
          bal.tab(result)
        }
      }, error = function(e) {
        message("Error generating balance summary for ", params$name, ": ", e)
        return(NULL)
      })

      saveRDS(list(love_plot = love_plot, balance_plot = balance_plot, p_dist_plot = p_dist_plot, summary = balance_summary), file = diagnostics_filename)
    } else {
      print("Loading cached diagnostics...")
      diagnostics <- readRDS(diagnostics_filename)
      love_plot <- diagnostics$love_plot
      balance_plot <- diagnostics$balance_plot
      balance_summary <- diagnostics$summary
    }

    all_results[[params$name]] <- list(
      result = result,
      love_plot = love_plot,
      p_dist_plot = p_dist_plot,
      balance_plot = balance_plot,
      summary = balance_summary
    )
  }

  return(all_results)
}

# These methods had issues with computation/time; revisit later.
# list(name = "genetic", method = "genetic", rerun = FALSE),
# list(name = "optimal_pair", method = "optimal", rerun = FALSE),
# list(name = "full", method = "full", rerun = FALSE),

# Define a list of matching methods and parameters, each with a unique name
methods <- list(
  list(name = "nearest_glm", method = "nearest", distance = "glm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gam", method = "nearest", distance = "gam", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_gbm", method = "nearest", distance = "gbm", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_lasso", method = "nearest", distance = "lasso", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_rpart", method = "nearest", distance = "rpart", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "subclass_glm", method = "subclass", distance = "glm", subclasses = 5, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "nearest_mahalanobis", method = "nearest", distance = "mahalanobis", replace = TRUE, estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE),
  list(name = "cbps", use="weightit", method="cbps", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "entropy", use="weightit", method="ebal", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "bart", use="weightit", method="bart", estimand="ATT", rerun_model=FALSE, rerun_diagnostics = TRUE),
  list(name = "twang_gbm", use="twanng", method = "twang", estimand="ATT", rerun_model = FALSE, rerun_diagnostics = TRUE)
)

results <- run_matching_methods(
  data = m_data,
  formula = matchingFormula,
  methods = methods,
  output_dir = ps_dir
)

```

Note that optimal pair matching would not run on our data. I have not tried optimal full matching. Genetic may run but over 18 hours later it still hadn't matched. I may debug this later, but for now we have 9 methods.

## Results Summary

```{r analysis_setup}
analysisFormulaSimple <- formula(
  outcome_va_vi_binary ~
    ever_lapse_binary
)

analysisFormulaMiddle <- formula(
  outcome_va_vi_binary ~
    ever_lapse_binary + 
    gender_cat +
    race_ethnic_cat +
    insurance_cat +
    age_cat
)

analysisFormulaFull <- formula(
  outcome_va_vi_binary ~
    ever_lapse_binary + 
    gender_cat +
    race_ethnic_cat +
    insurance_cat +
    age_cat +
    CCI +
    DCSI +
    other_inject +
    anti_VEGF +
    focal_laser_flag +
    PRP_flag +
    glaucoma_bef_hitplus_cat +
    otherretina_bef_hitplus_cat +
    catsurg_before_hitplus_cat
)

run_logistic_models <- function(design, formulas, method_name) {
  print(paste0("Estimating models: ", method_name))

  results <- list()
  roundTo <- 2
  roundP <- 3

  for (i in seq_along(formulas)) {
    if (method_name %in% c("cbps", "entropy", "bart")) {
      package <- "WeightIt"
      approach <- "ipw"
    } else if (method_name %in% c("twang_gbm")) {
      package <- "twang"
      approach <- "ipw"
    } else if (method_name == "subclass_glm") {
      package <- "MatchIt"
      approach <- "subclassification"
    } else {
      package <- "MatchIt"
      approach <- "match w/ replacement"
    }

    if (method_name == "subclass_glm") {
      # Extract subclass weights
      subclass_weights <- weights(design, type = "sampling")

      # Run weighted survey logistic regression
      model <- svyglm(formula = formulas[[i]], design = design, family = quasibinomial(), contrasts = NULL)
    } else {
      # Standard survey logistic regression
      model <- svyglm(formula = formulas[[i]], design = design, family = quasibinomial(), contrasts = NULL)
    }

    model_summary <- summary(model)
    coef_table <- model_summary$coefficients
    estimate <- coef_table[2, "Estimate"]

    or <- formatC(exp(estimate), format = "f", digits = roundTo)
    estimate <- formatC(estimate, format = "f", digits = roundTo)
    se <- formatC(coef_table[2, "Std. Error"], format = "f", digits = roundTo)
    t_value <- formatC(coef_table[2, "t value"], format = "f", digits = roundTo)
    p_value <- formatC(coef_table[2, "Pr(>|t|)"], format = "f", digits = roundP)

    confint_values <- confint(model, level = 0.95)[2, ]
    ci_lb_or <- formatC(exp(confint_values[1]), format = "f", digits = roundTo)
    ci_ub_or <- formatC(exp(confint_values[2]), format = "f", digits = roundTo)

    result_summ <- data.frame(
      package = package,
      method = method_name,
      approach = approach,
      estimand = "ATT",
      formula = names(formulas)[i],
      estimate = estimate,
      estimate_or = or,
      se = se,
      t_value = t_value,
      p_value = p_value,
      lb_95_or = ci_lb_or,
      ub_95_or = ci_ub_or
    )

    rownames(result_summ) <- NULL
    results[[i]] <- result_summ
  }

  do.call(rbind, results)
}
```

Below are results for the simple (treatment only), middle (demographics only), and full (everything). See the formulas above.

I excluded CEM (coarsened exact matching) because in the process it ended up dropping patients with certain factor levels and so estimation wouldn't run because of lack of contrasts. We could respecify models to exclude those covariates but I would not go that route.

```{r results, echo=FALSE}

formulas <- list(
  tx_only = analysisFormulaSimple,
  # middle = analysisFormulaMiddle,
  full = analysisFormulaFull
)

# Start with naive model results
naive_design <- svydesign(ids = ~e_mrn_deidentified, data = m_data, weights=~1)
naive_results <- run_logistic_models(naive_design, formulas, "Naive")

all_results <- list(naive_results)

# Removing CEM because matching process drops levels of certain factors due to unavailable matches.
method_names <- names(results)
method_names <- method_names[! method_names %in% c("cem")]

for (method_name in method_names) {
  print(paste0("Running model for method: ", method_name))
  method_result <- results[[method_name]]

  if(method_name %in% c("cbps", "entropy", "energy", "bart")) {
    # Weightit methods
    matched_data <- m_data
    weights <- method_result$result$weights
    matched_data$weights <- weights
    design <- svydesign(ids = ~e_mrn_deidentified, weights = ~weights, data = matched_data)
  } else if (method_name == "twang_gbm") {
    # For TWANG model with weights
    m_data$twang_att_w <- get.weights(method_result$result, stop.method = "es.mean")
    design <- svydesign(ids = ~e_mrn_deidentified, weights = ~twang_att_w, data = m_data)
  } else if(method_name == "subclass_glm") {
    matched_data <- MatchIt::match_data(method_result$result, subclass = "subclass")
    design <- svydesign(ids = ~e_mrn_deidentified, data = matched_data, weights=~weights)
    design <- postStratify(design, ~subclass, xtabs(~subclass, data = matched_data))
  }  else {
    matched_data <- MatchIt::get_matches(method_result$result)
    design <- svydesign(ids = ~e_mrn_deidentified, data = matched_data, weights=~1)
  }

  # Run logistic models and store results
  method_results <- run_logistic_models(design, formulas, method_name)
  all_results[[method_name]] <- method_results
}

# Combine all results into a single data frame
final_results <- do.call(rbind, all_results)

# Remove row names
rownames(final_results) <- NULL

final_results |> 
  filter(formula == "full") |> 
  arrange(estimate) |> 
  kable(format = "markdown", caption = "Logistic Regression Results Across Matching Methods (Full Model)")

saveRDS(final_results, file.path(ps_dir, paste0("ps_results_summary.rds")))
```

## Diagnostics

```{r diag_plots}

for (method_name in method_names) {
  # Get balance and common support plots for the method
  method_results <- results[[method_name]]
  love_plot <- method_results$love_plot
  balance_plot <- method_results$balance_plot
  p_dist_plot <- method_results$p_dist_plot

  # Output section header for balance plot
  if (!is.null(love_plot)) {
    cat("### Love Plot for", method_name, "\n\n")
    print(love_plot)
    cat("\n\n")
  }

  # Output section header for balanace plot of pscore (if exists)
  if (!is.null(balance_plot)) {
    cat("### Balance Plot of distance measure for", method_name, "\n\n")
    print(balance_plot)
    cat("\n\n")
  }

    # Output section header for box plot of p-scores (if exists)
  if (!is.null(p_dist_plot)) {
    cat("### Balance Plot of distance measure for", method_name, "\n\n")
    print(p_dist_plot)
    cat("\n\n")
  }
}
```

```{r save_diag_plots}

if(! dir.exists("ps_diagnostics")) {
  dir.create("ps_diagnostics")
}

for (method_name in method_names) {
  # Get balance and common support plots for the method
  print(paste0("saving plots for ", method_name))
  method_results <- results[[method_name]]
  method_result <- method_results$result

  if(method_name == "subclass_glm") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_results$p_dist_plot, type="jitter", interactive=FALSE)
    dev.off()
  } else if(method_name == "nearest_mahalanobis") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)
  } else if(class(method_results$result) == "matchit") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_results$p_dist_plot, type="jitter", interactive=FALSE)
    dev.off()
  } else if(class(method_results$result) == "ps") {
    # love plot
    png(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_result, plots = 3)
    dev.off()

    # balance plot
    ps_data <- method_result$data
    ps_data$p_score <- method_result$ps$es.mean.ATT

    bplot <- ggplot(ps_data, aes(x=p_score, fill=person_ever_lapse)) + 
      geom_density(alpha=0.4)
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), bplot, dpi=300)

    # jitter plot
    png(file.path("ps_diagnostics", paste0(method_name, "_p_dist_plot.png")), res = 300, width = 8, height = 6, units = "in")
    plot(method_result, plots = 2)
    dev.off()
  } else if(class(method_result) == "weightit") {
    # love plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_love_plot.png")), method_results$love_plot, dpi=300)

    # balance plot
    ggsave(file.path("ps_diagnostics", paste0(method_name, "_balance_plot.png")), method_results$balance_plot, dpi=300)
  }
}
```

## Margins Model

From here on out, we'll use the twang model with IPW.

A key research question we had is the relationship between lapse and vision impairment for certain subpopulations, including those with:

* PDR on any treatment
* PDR on anti-VEGF only
* PDR on PRP only

```{r post-analysis}
design <- svydesign(ids = ~e_mrn_deidentified, weights = ~twang_att_w, data = m_data)
model <- svyglm(formula = formulas[["full"]], design = design, family = quasibinomial(), contrasts=NULL)

avg_margfx <- margins(model, design=design)

all_mfx <- summary(avg_margfx)

# PDR on any treatment
design_pdr_any_tx <- subset(design, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_pdr_any_tx <- subset(m_data, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_any_tx <- data_pdr_any_tx |> nrow()
mfx_any_tx <- summary(margins(model, design = design_pdr_any_tx, data = data_pdr_any_tx, variables=c("ever_lapse_binary")))

# PDR on anti-VEGF only
design_pdr_anti_vegf <- subset(design, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
data_pdr_anti_vegf <- subset(m_data, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
n_anti_vegf <- data_pdr_anti_vegf |> nrow()
mfx_anti_vegf <- summary(margins(model, design = design_pdr_anti_vegf, data = data_pdr_anti_vegf, variables=c("ever_lapse_binary")))

# PDR on PRP only
design_pdr_prp <- subset(design, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
data_pdr_prp <- subset(m_data, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
n_pdr_prp <- data_pdr_prp |> nrow()
mfx_pdr_prp <- summary(margins(model, design = design_pdr_prp, data = data_pdr_prp, variables=c("ever_lapse_binary")))

```

### Twang: Average Marginal Effect

```{r avg_ame}
all_mfx |> kable()
```

### Twang: PDR on any treatment

```{r avg_ame_any_tx}
print(n_any_tx)
mfx_any_tx |> kable()
```

### Twang: PDR on AntiVEGF

```{r avg_ame_antivegf}
print(n_anti_vegf)
mfx_anti_vegf |> kable()
```

### Twang: PDR on PRP

```{r avg_ame_prp}
print(n_pdr_prp)
mfx_pdr_prp |> kable()
```

### Gut check on similar results: Nearest Neighbor GLM matching

```{r gut_check}

nearest_glm_data <- match.data(results[["nearest_glm"]]$result)
nearest_glm_design <- svydesign(ids = ~e_mrn_deidentified, data = nearest_glm_data, weights=~1)
nearest_glm_model <- svyglm(formula = formulas[["full"]], design = nearest_glm_design, family = binomial())
nearest_glm_avg_mfx <- summary(margins(nearest_glm_model, design = nearest_glm_design, variables=c("ever_lapse_binary")))

# PDR on any treatment
nearest_glm_design_pdr_any_tx <- subset(nearest_glm_design, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
nearest_glm_data_pdr_any_tx <- subset(nearest_glm_data, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
nearest_glm_n_any_tx <- nearest_glm_data_pdr_any_tx |> nrow()
nearest_glm_mfx_any_tx <- summary(margins(model, design = nearest_glm_design_pdr_any_tx, data = nearest_glm_data_pdr_any_tx, variables=c("ever_lapse_binary")))

# PDR on anti-VEGF only
nearest_glm_design_pdr_anti_vegf <- subset(nearest_glm_design, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
nearest_glm_data_pdr_anti_vegf <- subset(nearest_glm_data, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
nearest_glm_n_anti_vegf <- nearest_glm_data_pdr_anti_vegf |> nrow()
nearest_glm_mfx_anti_vegf <- summary(margins(model, design = nearest_glm_design_pdr_anti_vegf, data = nearest_glm_data_pdr_anti_vegf, variables=c("ever_lapse_binary")))

# PDR on PRP only
nearest_glm_design_pdr_prp <- subset(nearest_glm_design, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
nearest_glm_data_pdr_prp <- subset(nearest_glm_data, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
nearest_glm_n_pdr_prp <- nearest_glm_data_pdr_prp |> nrow()
nearest_glm_mfx_pdr_prp <- summary(margins(model, design = nearest_glm_design_pdr_prp, data = nearest_glm_data_pdr_prp, variables=c("ever_lapse_binary")))

```

### Nearest GLM: Average Marginal Effect

```{r nearest_glm_avg_ame}
nearest_glm_avg_mfx |> kable()
```

### Nearest GLM: PDR on any treatment

```{r nearest_glm_avg_ame_any_tx}
print(nearest_glm_n_any_tx)
nearest_glm_mfx_any_tx |> kable()
```

### Nearest GLM: PDR on AntiVEGF

```{r nearest_glm_avg_ame_anti_vegf}
print(nearest_glm_n_anti_vegf)
nearest_glm_mfx_anti_vegf |> kable()
```

### Nearest GLM: PDR on PRP

```{r nearest_glm_avg_ame_prp}
print(nearest_glm_n_pdr_prp)
nearest_glm_mfx_pdr_prp |> kable()
```