### 2025-03-27 results-summary.Rmd
refactor models to have access to full model

---
title: "Results Summary"
author: "Erik Westlund"
date: "2025-02-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(dplyr)
library(ggplot2)
library(kableExtra)
source("settings.R")

# Get results summary from the pipeline; no PHI in here
results <- readRDS("ps_results_summary.rds")
# results <- readRDS(file.path(ps_dir, "ps_results_summary.rds"))

fullResults <- results |> filter(formula == "full")

fullResultsForHumans <- fullResults |> 
  mutate(
    estimate_or = estimate_or |> as.numeric(),
    se = se |> as.numeric(),
    t_value = t_value |> as.numeric(),
    p_value = p_value |> as.numeric(),
    lb_95_or = lb_95_or |> as.numeric(),
    ub_95_or = ub_95_or |> as.numeric(),
    method = case_when(
      method == "Naive" ~ "No Propensity Score Balancing",
      method == "nearest_glm" ~ "Nearest Neighbor Matching, GLM (MatchIt)",
      method == "nearest_gam" ~ "Nearest Neighbor Matching, GAM (MatchIt)",
      method == "nearest_gbm" ~ "Nearest Neighbor Matching, GBM (MatchIt)",
      method == "nearest_lasso" ~ "Nearest Neighbor Matching, LASSO (MatchIt)",
      method == "nearest_mahalanobis" ~ "Nearest Neighbor Matching, Mahalanobis (MatchIt)",
      method == "nearest_rpart" ~ "Nearest Neighbor Matching, Recursive Partinioning (MatchIt)",
      method == "subclass_glm" ~ "Subclassification, GLM (MatchIt)",
      method == "cbps" ~ "IPTW, Covariate Balancing Propensity Score (WeightIt)",
      method == "entropy" ~ "IPTW, Entropy Balancing (WeightIt)",
      method == "bart" ~ "IPTW, Bayesian Additive Regression Trees (WeightIt)",
      method == "twang_gbm" ~ "IPTW, GBM (twang)",
      TRUE ~ "Unknown"
    )
  ) |> 
  select(
    method, estimand, estimate_or, se, t_value, p_value, lb_95_or, ub_95_or
  ) |> 
  rename(
    "OR" = estimate_or,
    "SE" = se,
    "t" = t_value,
    "Lower" = lb_95_or,
    "Upper" = ub_95_or
  ) |> 
  arrange(OR) |>
  mutate(method = factor(method, levels = method))

# Define colors
fullResultsForHumans$color <- "black"
fullResultsForHumans$color[fullResultsForHumans$method == "No Propensity Score Balancing"] <- "red"
fullResultsForHumans$color[fullResultsForHumans$method == "IPTW, GBM (twang)"] <- "blue"
```

```{r, echo = FALSE, fig.width = 10, fig.height = 8}
# Plot
plot <- ggplot(fullResultsForHumans, aes(x = reorder(method, OR), y = OR, color = color)) +
  geom_pointrange(aes(ymin = Lower, ymax = Upper), size = 0.8) +
  geom_point(size = 3) +
  geom_text(aes(label = sprintf("%.2f", OR), color = color), vjust = -1.2, hjust = 0.5, size = 4) +
  coord_flip() +
  scale_color_identity() +
  scale_y_continuous(breaks = seq(1, max(fullResultsForHumans$Upper), by = 0.1)) +
  scale_x_discrete(expand = c(0.075, 0.075)) +
  theme_minimal(base_size = 14) +
  labs(
    x = element_blank(),
    y = "Odds Ratio",
    title = "Doubly Robust Propensity Score Adjusted\n Model Estimates (ATT)",
    caption = "\n\nOdds Ratio of 1.0 = No Association.
               MatchIt, WeightIt, and twang R packages are used for estimation.
               IPTW = Inverse probability of treatment weighting.
               IPTW Twang GBM results (shown in blue) reported throughout paper.
               No Adjustment (shown in red) reflects results with no propensity score adjustment."
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", margin = margin(b = 15)),
    legend.position = "none",
    axis.line = element_line(color = "gray80"),
    axis.text = element_text(color = "gray30"),
    axis.ticks = element_line(color = "gray80"),
    axis.title.x = element_text(margin = margin(t = 10)),
    plot.caption = element_text(margin = margin(t = 15)),
    plot.margin = margin(20, 20, 30, 20)
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray70") +  
  geom_hline(yintercept = 1, linetype = "dashed",  color = "gray30")

plot
ggsave("results_summary.png", plot, width = 10, height = 8, dpi = 300)

tableOutput <- fullResultsForHumans |> 
  select(-color) |> 
  rename(
    Method = method,
    Estimand = estimand,
  )

tableOutput$Method <- factor(
  fullResultsForHumans$method, 
  levels = c(
    "IPTW, GBM (twang)",  # Move to top
    "No Adjustment",      # Move to second
    setdiff(levels(factor(fullResultsForHumans$method)), c("IPTW, GBM (twang)", "No Adjustment"))  # Keep existing order for others
  )
)

tableOutput <- tableOutput |> arrange(Method)

tableOutput |> 
  kable()
```