### 2025-06-23 twang-analysis.Rmd

---
title: "Twang only analysis summary"
author: "Erik Westlund"
date: "`r Sys.Date()`"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

source("settings.R")

library(dplyr)
library(kableExtra)
library(MatchIt)
library(twang)
library(survey)
library(margins)

# Note this depends on output from analysis.rmd

m_data <- readRDS(file.path(ps_dir, "m_data.rds"))
twang <- readRDS(file.path(ps_dir, "twang_gbm.rds"))
formulas <- readRDS(file.path(ps_dir, "formulas.rds"))
```

# Model

This model adjusts for multiple eyes by using model-robust standard errors (specifically, Horvitz-Thompson standard errors, which are generalized from model-robust sandwich estimators). Details are in the documentation for the survey package.

The below output is before exponentiation. To exponentiate any coefficient, just do `e^b` where `b` is the coefficient and `e` is Euler's constant. In `R` or `Stata`, type `exp(b)` for short-hand

```{r analysis}
m_data$twang_att_w <- get.weights(twang$result, stop.method = "es.mean")
design <- svydesign(ids = ~e_mrn_deidentified, weights = ~twang_att_w, data = m_data)
model <- svyglm(formula = formulas[["full"]], design = design, family = quasibinomial())

summary(model)
```

# Marginal effects

Inspect logic for bugs given quickness of this turnaround

```{r mfx}
avg_margfx <- margins(model, design=design)

all_mfx <- summary(avg_margfx)

# PDR on any treatment
design_pdr_any_tx <- subset(design, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_pdr_any_tx <- subset(m_data, PDR == "Present" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_any_tx <- data_pdr_any_tx |> nrow()
mfx_any_tx <- summary(margins(model, design = design_pdr_any_tx, data = data_pdr_any_tx, variables=c("ever_lapse_binary")))

# PDR on anti-VEGF only
design_pdr_anti_vegf <- subset(design, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
data_pdr_anti_vegf <- subset(m_data, PDR == "Present" & PRP_flag == 0 & anti_VEGF == 1 & focal_laser_flag == 0 & other_inject == 0)
n_anti_vegf <- data_pdr_anti_vegf |> nrow()
mfx_anti_vegf <- summary(margins(model, design = design_pdr_anti_vegf, data = data_pdr_anti_vegf, variables=c("ever_lapse_binary")))

# PDR on PRP only
design_pdr_prp <- subset(design, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
data_pdr_prp <- subset(m_data, PDR == "Present" & PRP_flag == 1 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0)
n_pdr_prp <- data_pdr_prp |> nrow()
mfx_pdr_prp <- summary(margins(model, design = design_pdr_prp, data = data_pdr_prp, variables=c("ever_lapse_binary")))

# NPDR or NoDR not on treatment
design_npdr_nodr_no_tx <- subset(design, (!is.na(NPDR) | No_DR == TRUE) & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
data_npdr_nodr_no_tx <- subset(m_data, (!is.na(NPDR) | No_DR == TRUE) & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
n_npdr_nodr_no_tx <- data_npdr_nodr_no_tx |> nrow()
mfx_npdr_nodr_no_tx <- summary(margins(model, design = design_npdr_nodr_no_tx, data = data_npdr_nodr_no_tx, variables=c("ever_lapse_binary")))

# NPDR or NoDR any tx
design_npdr_nodr_any_tx <- subset(design, (!is.na(NPDR) | No_DR == TRUE) & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_npdr_nodr_any_tx <- subset(m_data, (!is.na(NPDR) | No_DR == TRUE) & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_npdr_nodr_any_tx <- data_npdr_nodr_any_tx |> nrow()
mfx_npdr_nodr_any_tx <- summary(margins(model, design = design_npdr_nodr_any_tx, data = data_npdr_nodr_any_tx, variables=c("ever_lapse_binary")))

# NPDR not on treatment
design_npdr_no_tx <- subset(design, !is.na(NPDR) & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
data_npdr_no_tx <- subset(m_data, !is.na(NPDR) & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
n_npdr_no_tx <- data_npdr_no_tx |> nrow()
mfx_npdr_no_tx <- summary(margins(model, design = design_npdr_no_tx, data = data_npdr_no_tx, variables=c("ever_lapse_binary")))

# NPDR  on any treatment
design_npdr_any_tx <- subset(design, !is.na(NPDR) & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_npdr_any_tx <- subset(m_data, !is.na(NPDR) & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_npdr_any_tx <- data_npdr_any_tx |> nrow()
mfx_npdr_any_tx <- summary(margins(model, design = design_npdr_any_tx, data = data_npdr_any_tx, variables=c("ever_lapse_binary")))

# NPDR not on treatment: Mild
design_npdr_mild_no_tx <- subset(design, NPDR == "Mild" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
data_npdr_mild_no_tx <- subset(m_data, NPDR == "Mild" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
n_npdr_mild_no_tx <- data_npdr_mild_no_tx |> nrow()
mfx_npdr_mild_no_tx <- summary(margins(model, design = design_npdr_mild_no_tx, data = data_npdr_mild_no_tx, variables=c("ever_lapse_binary")))

# NPDR  on any treatment: Mild
design_npdr_mild_any_tx <- subset(design, NPDR == "Mild" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_npdr_mild_any_tx <- subset(m_data, NPDR == "Mild" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_npdr_mild_any_tx <- data_npdr_mild_any_tx |> nrow()
mfx_npdr_mild_any_tx <- summary(margins(model, design = design_npdr_mild_any_tx, data = data_npdr_mild_any_tx, variables=c("ever_lapse_binary")))

# NPDR not on treatment: Moderate
design_npdr_moderate_no_tx <- subset(design, NPDR == "Moderate" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
data_npdr_moderate_no_tx <- subset(m_data, NPDR == "Moderate" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
n_npdr_moderate_no_tx <- data_npdr_moderate_no_tx |> nrow()
mfx_npdr_moderate_no_tx <- summary(margins(model, design = design_npdr_moderate_no_tx, data = data_npdr_moderate_no_tx, variables=c("ever_lapse_binary")))

# NPDR  on any treatment: Moderate
design_npdr_moderate_any_tx <- subset(design, NPDR == "Moderate" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_npdr_moderate_any_tx <- subset(m_data, NPDR == "Moderate" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_npdr_moderate_any_tx <- data_npdr_moderate_any_tx |> nrow()
mfx_npdr_moderate_any_tx <- summary(margins(model, design = design_npdr_moderate_any_tx, data = data_npdr_moderate_any_tx, variables=c("ever_lapse_binary")))

# NPDR not on treatment: Severe
design_npdr_severe_no_tx <- subset(design, NPDR == "Severe" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
data_npdr_severe_no_tx <- subset(m_data, NPDR == "Severe" & (PRP_flag == 0 & anti_VEGF == 0 & focal_laser_flag == 0 & other_inject == 0))
n_npdr_severe_no_tx <- data_npdr_severe_no_tx |> nrow()
mfx_npdr_severe_no_tx <- summary(margins(model, design = design_npdr_severe_no_tx, data = data_npdr_severe_no_tx, variables=c("ever_lapse_binary")))

# NPDR  on any treatment: Severe
design_npdr_severe_any_tx <- subset(design, NPDR == "Severe" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
data_npdr_severe_any_tx <- subset(m_data, NPDR == "Severe" & (PRP_flag == 1 | anti_VEGF == 1 | focal_laser_flag == 1 | other_inject == 1))
n_npdr_severe_any_tx <- data_npdr_severe_any_tx |> nrow()
mfx_npdr_severe_any_tx <- summary(margins(model, design = design_npdr_severe_any_tx, data = data_npdr_severe_any_tx, variables=c("ever_lapse_binary")))

```

```{r forest_plots}
library(ggplot2)
library(dplyr)

# Extract marginal effects data from all calculated objects
# Each mfx object has columns: factor, AME, SE, z, p, lower, upper
forest_data <- bind_rows(
  # Overall population
  all_mfx |> 
    filter(factor == "ever_lapse_binary") |> 
    mutate(
      label = "Overall Population",
      group_type = "overall",
      n = nrow(m_data)
    ),

  # No DR or NPDR subgroups
  mfx_npdr_nodr_no_tx |> 
    mutate(
      label = "No DR or NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_nodr_no_tx
    ),
  mfx_npdr_nodr_any_tx |> 
    mutate(
      label = "No DR or NPDR, any treatment",
      group_type = "subgroup", 
      n = n_npdr_nodr_any_tx
    ),

  # NPDR overall
  mfx_npdr_no_tx |> 
    mutate(
      label = "NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_no_tx
    ),
  mfx_npdr_any_tx |> 
    mutate(
      label = "NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_any_tx
    ),

  # NPDR by severity
  mfx_npdr_mild_no_tx |> 
    mutate(
      label = "Mild NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_mild_no_tx
    ),
  mfx_npdr_mild_any_tx |> 
    mutate(
      label = "Mild NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_mild_any_tx
    ),
  mfx_npdr_moderate_no_tx |> 
    mutate(
      label = "Moderate NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_moderate_no_tx
    ),
  mfx_npdr_moderate_any_tx |> 
    mutate(
      label = "Moderate NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_moderate_any_tx
    ),
  mfx_npdr_severe_no_tx |> 
    mutate(
      label = "Severe NPDR, no treatment",
      group_type = "subgroup",
      n = n_npdr_severe_no_tx
    ),
  mfx_npdr_severe_any_tx |> 
    mutate(
      label = "Severe NPDR, any treatment",
      group_type = "subgroup",
      n = n_npdr_severe_any_tx
    ),

  # PDR subgroups
  mfx_any_tx |> 
    mutate(
      label = "PDR, any treatment",
      group_type = "subgroup",
      n = n_any_tx
    ),
  mfx_anti_vegf |> 
    mutate(
      label = "PDR, on anti-VEGF only",
      group_type = "subgroup",
      n = n_anti_vegf
    ),
  mfx_pdr_prp |> 
    mutate(
      label = "PDR, on PRP only",
      group_type = "subgroup",
      n = n_pdr_prp
    )
) |>
  mutate(
    # Add sample sizes to labels
    label_with_n = paste0(label, " (n=", n, ")"),
    # Calculate confidence intervals
    Lower = AME - 1.96 * SE,
    Upper = AME + 1.96 * SE
  )

# Define order for plots (reversed to match original plot)
subgroup_order <- rev(c(
  "No DR or NPDR, no treatment", "No DR or NPDR, any treatment", "NPDR, no treatment",
  "NPDR, any treatment", "Mild NPDR, no treatment", "Mild NPDR, any treatment",
  "Moderate NPDR, no treatment", "Moderate NPDR, any treatment",
  "Severe NPDR, no treatment", "Severe NPDR, any treatment",
  "PDR, any treatment", "PDR, on anti-VEGF only", "PDR, on PRP only"
))

# Version 1: Subgroups only (matching original forest plot)
data_subgroups <- forest_data |> 
  filter(group_type == "subgroup") |>
  mutate(Labels = factor(label, levels = subgroup_order))

plot_subgroups <- ggplot(data_subgroups, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  geom_point(size = 3, color = "#2563eb") +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "#2563eb") +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.000, 0.06),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "The Average Marginal Effects of Lapse Across Selected Groups",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("All p-values < ", max(data_subgroups$p, na.rm = TRUE))
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Version 2: With overall population included
all_order <- c("Overall Population", subgroup_order)
data_all <- forest_data |>
  mutate(
    Labels = factor(label, levels = rev(all_order)),
    # Different colors for overall vs subgroups
    point_color = ifelse(group_type == "overall", "#dc2626", "#2563eb"),
    point_size = ifelse(group_type == "overall", 4, 3)
  )

plot_with_overall <- ggplot(data_all, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  # Add horizontal separator line after overall population
  geom_hline(yintercept = length(subgroup_order) + 0.5, color = "gray50", linetype = "dashed", alpha = 0.5) +
  geom_point(aes(color = point_color, size = point_size)) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = point_color), height = 0.2) +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  scale_color_identity() +
  scale_size_identity() +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.000, 0.06),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "The Average Marginal Effects of Lapse Across All Groups",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("All p-values < ", max(data_all$p, na.rm = TRUE), 
                    "\nOverall population shown in red, subgroups in blue")
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = margin(b = 20)),
    plot.margin = margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Display the plots
print(plot_subgroups)
print(plot_with_overall)

# Save the plots
ggsave("forest_plot_subgroups_only.png", plot_subgroups, 
       width = 10, height = 8, dpi = 300, bg = "white")
ggsave("forest_plot_with_overall.png", plot_with_overall, 
       width = 10, height = 8.5, dpi = 300, bg = "white")

# Also save extracted data for reference
write.csv(forest_data, "forest_plot_data.csv", row.names = FALSE)

```

```{r report}

# PDR on any treatment
print("PDR on any treatment:")
print(n_any_tx)
mfx_any_tx |> kable()

# PDR on anti-VEGF only
print("PDR on anti-VEGF only:")
print(n_anti_vegf)
mfx_anti_vegf |> kable()

# PDR on PRP only
print("PDR on PRP only:")
print(n_pdr_prp)
mfx_pdr_prp |> kable()

# NPDR or NoDR not on treatment
print("NPDR or NoDR not on treatment:")
print(n_npdr_nodr_no_tx)
mfx_npdr_nodr_no_tx |> kable()

# NPDR or NoDR any tx
print("NPDR or NoDR any tx:")
print(n_npdr_nodr_any_tx)
mfx_npdr_nodr_any_tx |> kable()

# NPDR not on treatment
print("NPDR not on treatment:")
print(n_npdr_no_tx)
mfx_npdr_no_tx |> kable()

# NPDR  on any treatment
print("NPDR  on any treatment:")
print(n_npdr_any_tx)
mfx_npdr_any_tx |> kable()

# NPDR not on treatment: Mild
print("NPDR not on treatment: Mild:")
print(n_npdr_mild_no_tx)
mfx_npdr_mild_no_tx |> kable()

# NPDR  on any treatment: Mild
print("NPDR  on any treatment: Mild:")
print(n_npdr_mild_any_tx)
mfx_npdr_mild_any_tx |> kable()

# NPDR not on treatment: Moderate
print("NPDR not on treatment: Moderate:")
print(n_npdr_moderate_no_tx)
mfx_npdr_moderate_no_tx |> kable()

# NPDR  on any treatment: Moderate
print("NPDR  on any treatment: Moderate:")
print(n_npdr_moderate_any_tx)
mfx_npdr_moderate_any_tx |> kable()

# NPDR not on treatment: Severe
print("NPDR not on treatment: Severe:")
print(n_npdr_severe_no_tx)
mfx_npdr_severe_no_tx |> kable()

# NPDR  on any treatment: Severe
print("NPDR  on any treatment: Severe:")
print(n_npdr_severe_any_tx)
mfx_npdr_severe_any_tx |> kable()

```