### 2025-08-08 reviewer-response.Rmd
use margins package

---
title: "Reviewer Response Analysis"
author: "Erik Westlund"
date: "`r Sys.Date()`"
output: 
  html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
    code_folding: hide
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  warning = FALSE,
  message = FALSE,
  fig.width = 10,
  fig.height = 6
)

# Load required libraries and settings
source("settings.R")
source("functions.R")

# Load all required libraries
load_required_libraries()
```

# Data Preparation

```{r load_data}
# Load and prepare data
data_path <- file.path(s_root, "Gina's Project/df_Final_240909.csv")
data <- load_and_prepare_data(data_path)

# Apply the same exclusions as main analysis
m_data <- apply_exclusions(data)

print(sprintf("Total observations after exclusions: %d", nrow(m_data)))
print(sprintf("Total unique patients: %d", length(unique(m_data$e_mrn_deidentified))))
```

## Load IPTW Weights from Analysis

```{r twang_weights}
# Load cached results from the main analysis pipeline
cached_results <- load_cached_results("ps_results")

# Check if twang model exists in cached results
if(!"twang_gbm" %in% names(cached_results)) {
  stop("No cached twang model found. Please run analysis.Rmd first to generate IPTW weights.")
}

print("Loading cached twang model for IPTW weights...")
twang_model <- cached_results$twang_gbm$result

# Extract IPTW weights from twang model (same as in analysis.Rmd)
m_data$twang_att_w <- get.weights(twang_model, stop.method = "es.mean")

# Summary of weights
print("Weight distribution:")
summary(m_data$twang_att_w)
```

## Main Effect of Lapsing (IPTW-Weighted)

```{r main_effect}
# Run the main model with full covariates to get the lapse effect
# Using the same formula as analysisFormulaFull from functions.R
main_formula <- get_analysis_formulas()$full

# Create weighted survey design
survey_design_weighted <- svydesign(
  ids = ~e_mrn_deidentified,
  data = m_data,
  weights = ~twang_att_w
)

# Fit the main model
main_model <- svyglm(
  formula = main_formula,
  design = survey_design_weighted,
  family = quasibinomial()
)

# Extract lapse effect
coef_summary <- summary(main_model)$coefficients
lapse_est <- coef_summary["ever_lapse_binary", "Estimate"]
lapse_se <- coef_summary["ever_lapse_binary", "Std. Error"]
lapse_p <- coef_summary["ever_lapse_binary", "Pr(>|t|)"]

# Calculate confidence intervals
lapse_ci <- confint(main_model)["ever_lapse_binary", ]

# Create results table
lapse_results <- data.frame(
  Log_Odds = round(lapse_est, 4),
  Odds_Ratio = round(exp(lapse_est), 4),
  CI_Lower = round(lapse_ci[1], 4),
  CI_Upper = round(lapse_ci[2], 4),
  P_Value = format.pval(lapse_p, digits = 4, eps = 0.0001)
)

knitr::kable(lapse_results, 
             caption = "Effect of Lapsing on Vision Impairment (IPTW-Weighted)",
             row.names = FALSE)
```

# Reviewer Question 1: Subgroup Testing with Formal Interaction Terms

**Reviewer Comment:** "Inadequate Subgroup Testing: Visual forest‑plot overlap doesn't prove consistent effects across subgroups (e.g., PDR+anti‑VEGF vs. PDR+PRP). Without formal interaction testing, subgroup claims lack statistical backing."

**Response:** We perform formal interaction testing using two models:
1. Interaction between lapse status and DR severity
2. Interaction between lapse status and treatment type

## Create Analysis Variables

```{r create_variables}
# Create the treatment type variable and properly labeled DR severity
m_data <- m_data |>
  mutate(
    treatment_type = case_when(
      anti_VEGF == 1 ~ "anti_VEGF",
      PRP_flag == 1 ~ "PRP",
      other_inject == 1 | focal_laser_flag == 1 ~ "other_treatment",
      TRUE ~ "no_treatment"
    ),
    treatment_type = factor(treatment_type, 
                           levels = c("no_treatment", "anti_VEGF", "PRP", "other_treatment")),
    # Create labeled DR severity factor for clearer interaction terms
    dr_severity = factor(person_dr,
                        levels = c(0, 1, 2),
                        labels = c("No_DR", "NPDR", "PDR"))
  )

# Check distribution of treatment types
print("Treatment Type Distribution:")
table(m_data$treatment_type)

# Display DR severity coding and distribution
print("DR Severity Distribution:")
table(m_data$dr_severity)
```

## Variable Definitions

```{r variable_definitions}
# Create a reference table for all key variables
variable_definitions <- data.frame(
  Variable = c("ever_lapse_binary", "outcome_va_vi_binary", "dr_severity", 
               "treatment_type", "person_dr", "person_dr_severity"),
  Type = c("Binary", "Binary", "Factor", "Factor", "Numeric", "Numeric"),
  Values = c("0 = No lapse, 1 = Lapsed from care",
             "0 = No vision impairment, 1 = Vision impairment (VA ≥ 0.3 logMAR)",
             "No_DR, NPDR, PDR (labeled factor for interactions)",
             "no_treatment, anti_VEGF, PRP, other_treatment",
             "0 = No DR, 1 = NPDR, 2 = PDR (numeric coding)",
             "0 = No DR, 1 = Mild NPDR, 2 = Moderate NPDR, 3 = Severe NPDR, 4 = PDR"),
  stringsAsFactors = FALSE
)

knitr::kable(variable_definitions, 
             caption = "Key Variable Definitions",
             col.names = c("Variable", "Type", "Possible Values/Coding"))
```

## Define Model Formulas

```{r define_formulas}
# Define the base formula (same as analysisFormulaFull but without the outcome)
base_covariates <- c(
  "gender_cat",
  "race_ethnic_cat", 
  "insurance_cat",
  "age_cat",
  "CCI",
  "DCSI",
  "other_inject",
  "anti_VEGF",
  "focal_laser_flag",
  "PRP_flag",
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat",
  "catsurg_before_hitplus_cat"
)

# Model 1: Interaction with DR severity
# dr_severity: No_DR, NPDR, PDR
interaction_dr_formula <- as.formula(paste(
  "outcome_va_vi_binary ~ ever_lapse_binary * dr_severity +",
  paste(base_covariates, collapse = " + ")
))

# Model 2: Interaction with treatment type
interaction_treatment_formula <- as.formula(paste(
  "outcome_va_vi_binary ~ ever_lapse_binary * treatment_type +",
  paste(base_covariates, collapse = " + ")
))

print("DR Interaction Formula:")
print(interaction_dr_formula)

print("Treatment Interaction Formula:")
print(interaction_treatment_formula)
```

## Run Interaction Models (IPTW Weighted)

```{r interaction_models, results='asis'}
# Store results for weighted analyses only
interaction_results <- list()

# Create weighted survey design
survey_design_weighted <- svydesign(
  ids = ~e_mrn_deidentified,
  data = m_data,
  weights = ~twang_att_w
)

### DR Severity Interaction Model

```{r dr_interaction}
# Fit Model 1: DR severity interaction
model_dr_interaction <- svyglm(
  formula = interaction_dr_formula,
  design = survey_design_weighted,
  family = quasibinomial()
)

# Summary of DR interaction model
summary_dr <- summary(model_dr_interaction)

# Extract interaction p-values for DR model
dr_interaction_terms <- grep("ever_lapse_binary:dr_severity", 
                            names(coef(model_dr_interaction)), value = TRUE)
dr_interaction_pvals <- if(length(dr_interaction_terms) > 0) {
  summary_dr$coefficients[dr_interaction_terms, "Pr(>|t|)"]
} else {
  numeric(0)
}

# Display results
dr_results <- data.frame(
  Term = dr_interaction_terms,
  Coefficient = summary_dr$coefficients[dr_interaction_terms, "Estimate"],
  SE = summary_dr$coefficients[dr_interaction_terms, "Std. Error"],
  P_Value = round(dr_interaction_pvals, 4),
  Significant = ifelse(dr_interaction_pvals < 0.05, "Yes", "No")
)

knitr::kable(dr_results, caption = "DR Severity Interaction Terms (Individual Coefficients)")
```

## Understanding the Interaction Model

In the interaction model `outcome ~ lapse * dr_severity`, the coefficients represent:

- **Main effect of lapse (β₁)**: Effect of lapsing for the reference group (No DR)
- **Main effects of DR severity (β₂, β₃)**: Baseline differences in risk for NPDR/PDR vs No DR among non-lapsers
- **Interaction terms (β₄, β₅)**: How much the effect of lapsing differs in NPDR/PDR compared to No DR

The actual effect of lapsing within each DR group is:
- **No DR**: β₁
- **NPDR**: β₁ + β₄ 
- **PDR**: β₁ + β₅

## Joint Test of Interaction Terms

```{r joint_interaction_test}

# Joint Wald test: Are the interaction terms jointly significant?
# H0: ever_lapse_binary:dr_severityNPDR = 0 AND ever_lapse_binary:dr_severityPDR = 0
# This tests whether the effect of lapsing is the same across all DR severity levels

joint_test <- linearHypothesis(
  model_dr_interaction,
  c("ever_lapse_binary:dr_severityNPDR = 0",
    "ever_lapse_binary:dr_severityPDR = 0"),
  vcov = vcov(model_dr_interaction)
)

# Display joint test results
joint_results <- data.frame(
  Test = "Joint interaction test",
  Chisq = round(joint_test$Chisq[2], 3),
  Df = joint_test$Df[2],
  P_Value = format.pval(joint_test$`Pr(>Chisq)`[2], digits = 4, eps = 0.0001)
)

knitr::kable(joint_results, 
             caption = "Joint Wald Test for DR Severity Interactions",
             row.names = FALSE)

# Interpret the joint test
if(joint_test$`Pr(>Chisq)`[2] < 0.05) {
  print("Joint test is significant: The effect of lapsing varies by DR severity")
} else {
  print("Joint test is not significant: The effect of lapsing is similar across DR severity levels")
}
```

## Effect of Lapsing Within Each DR Severity Group

```{r lapse_effects_by_dr}

# Method 1: Use emmeans WITHOUT nuisance factors for true marginalization
# Increase reference grid limit to handle the full factorial
emm_options(rg.limit = 200000)  # Increase limit to handle large reference grid

# Check for cached emmeans results
emmeans_cache_file <- "ps_results/emmeans_dr_interaction.rds"
if(file.exists(emmeans_cache_file)) {
  cat("Loading cached emmeans results...\n")
  emmeans_results <- readRDS(emmeans_cache_file)
  emm <- emmeans_results$emm
  lapse_contrasts <- emmeans_results$lapse_contrasts
  cat("Cached results loaded successfully.\n\n")
} else {
  cat("Computing emmeans (this may take several minutes)...\n")
  start_time <- Sys.time()

  # Calculate marginal means - this will average over all covariate combinations
  emm <- emmeans(model_dr_interaction, 
                 ~ ever_lapse_binary | dr_severity)

  # Get contrasts: effect of lapse (1 vs 0) within each DR severity
  lapse_contrasts <- contrast(emm, method = "revpairwise", by = "dr_severity")

  # Save to cache
  emmeans_results <- list(
    emm = emm,
    lapse_contrasts = lapse_contrasts,
    computed_at = Sys.time()
  )
  saveRDS(emmeans_results, file = emmeans_cache_file)

  end_time <- Sys.time()
  cat("Emmeans computation completed in", round(difftime(end_time, start_time, units = "mins"), 2), "minutes\n")
  cat("Results cached to:", emmeans_cache_file, "\n\n")
}

# Convert to data frame and clean up
lapse_effects <- as.data.frame(lapse_contrasts)

# Create clean results table
lapse_effects_table <- data.frame(
  DR_Severity = lapse_effects$dr_severity,
  Log_Odds = round(lapse_effects$estimate, 4),
  SE = round(lapse_effects$SE, 4),
  CI_Lower = round(lapse_effects$estimate - 1.96 * lapse_effects$SE, 4),
  CI_Upper = round(lapse_effects$estimate + 1.96 * lapse_effects$SE, 4),
  Odds_Ratio = round(exp(lapse_effects$estimate), 4),
  P_Value = format.pval(lapse_effects$p.value, digits = 4, eps = 0.0001)
)

knitr::kable(lapse_effects_table,
             caption = "Effect of Lapsing Within Each DR Severity Group",
             row.names = FALSE)
```

## Alternative: Manual Calculation of Lapse Effects

```{r manual_lapse_effects}
# Note: This manual calculation should produce identical results to emmeans above
# If emmeans still has issues, we can calculate the effects manually
# Extract coefficients and variance-covariance matrix
coef_vals <- coef(model_dr_interaction)
vcov_mat <- vcov(model_dr_interaction)

# Define the linear combinations for each DR group
# Effect in No DR: just the main effect
# Effect in NPDR: main effect + NPDR interaction
# Effect in PDR: main effect + PDR interaction

# Create contrast matrix
contrast_names <- c("Lapse effect in No DR",
                    "Lapse effect in NPDR", 
                    "Lapse effect in PDR")

# Initialize results
manual_effects <- data.frame(
  DR_Severity = c("No_DR", "NPDR", "PDR"),
  Log_Odds = numeric(3),
  SE = numeric(3),
  stringsAsFactors = FALSE
)

# Effect for No DR (just main effect)
manual_effects$Log_Odds[1] <- coef_vals["ever_lapse_binary"]
manual_effects$SE[1] <- sqrt(vcov_mat["ever_lapse_binary", "ever_lapse_binary"])

# Effect for NPDR (main + interaction)
if("ever_lapse_binary:dr_severityNPDR" %in% names(coef_vals)) {
  manual_effects$Log_Odds[2] <- coef_vals["ever_lapse_binary"] + 
                                 coef_vals["ever_lapse_binary:dr_severityNPDR"]
  # SE calculation for sum of coefficients
  manual_effects$SE[2] <- sqrt(
    vcov_mat["ever_lapse_binary", "ever_lapse_binary"] +
    vcov_mat["ever_lapse_binary:dr_severityNPDR", "ever_lapse_binary:dr_severityNPDR"] +
    2 * vcov_mat["ever_lapse_binary", "ever_lapse_binary:dr_severityNPDR"]
  )
}

# Effect for PDR (main + interaction)
if("ever_lapse_binary:dr_severityPDR" %in% names(coef_vals)) {
  manual_effects$Log_Odds[3] <- coef_vals["ever_lapse_binary"] + 
                                 coef_vals["ever_lapse_binary:dr_severityPDR"]
  # SE calculation for sum of coefficients
  manual_effects$SE[3] <- sqrt(
    vcov_mat["ever_lapse_binary", "ever_lapse_binary"] +
    vcov_mat["ever_lapse_binary:dr_severityPDR", "ever_lapse_binary:dr_severityPDR"] +
    2 * vcov_mat["ever_lapse_binary", "ever_lapse_binary:dr_severityPDR"]
  )
}

# Calculate CIs and OR
manual_effects <- manual_effects |>
  mutate(
    CI_Lower = Log_Odds - 1.96 * SE,
    CI_Upper = Log_Odds + 1.96 * SE,
    Odds_Ratio = exp(Log_Odds),
    # Calculate p-values (two-sided test)
    Z_stat = Log_Odds / SE,
    P_Value = 2 * pnorm(-abs(Z_stat))
  ) |>
  mutate(
    Log_Odds = round(Log_Odds, 4),
    SE = round(SE, 4),
    CI_Lower = round(CI_Lower, 4),
    CI_Upper = round(CI_Upper, 4),
    Odds_Ratio = round(Odds_Ratio, 4),
    P_Value = format.pval(P_Value, digits = 4, eps = 0.0001)
  ) |>
  select(-Z_stat)

knitr::kable(manual_effects,
             caption = "Effect of Lapsing Within Each DR Severity Group (Manual Calculation)",
             row.names = FALSE)

# Verify that emmeans and manual calculations match
if(exists("lapse_effects_table") && nrow(lapse_effects_table) > 0) {
  # Compare the two methods
  comparison <- data.frame(
    DR_Severity = manual_effects$DR_Severity,
    Emmeans_LogOdds = lapse_effects_table$Log_Odds,
    Manual_LogOdds = manual_effects$Log_Odds,
    Difference = abs(lapse_effects_table$Log_Odds - manual_effects$Log_Odds)
  )

  if(max(comparison$Difference) < 0.001) {
    cat("\n✓ Verification: Emmeans and manual calculations produce identical results (differences < 0.001)\n")
  } else {
    cat("\n⚠ Warning: Some differences between emmeans and manual calculations:\n")
    print(comparison)
  }
} else {
  cat("\nNote: Using manual calculation method as emmeans encountered issues with reference grid size.\n")
}
```

## Visualization of Interaction Effects

```{r visualize_interactions, fig.height=6, fig.width=10}
# Use the manual_effects data for visualization
plot_data <- manual_effects

# Bar chart showing magnitude of effects with error bars
p_magnitude <- ggplot(plot_data, aes(x = DR_Severity, y = Log_Odds, fill = DR_Severity)) +
  geom_col(alpha = 0.7) +
  geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
                width = 0.2, size = 1) +
  geom_hline(yintercept = 0, linetype = "solid", alpha = 0.5) +
  labs(title = "Magnitude of Lapse Effect Across DR Severity",
       subtitle = "Note: Confidence intervals for No DR and NPDR overlap",
       x = "DR Severity",
       y = "Log Odds (Effect Size)") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        legend.position = "none") +
  scale_fill_manual(values = c("No_DR" = "#1f77b4", 
                               "NPDR" = "#ff7f0e", 
                               "PDR" = "#d62728")) +
  geom_text(aes(label = sprintf("OR: %.2f", Odds_Ratio)),
            vjust = ifelse(plot_data$Log_Odds > 0, -1, 1.5),
            size = 4)

print(p_magnitude)
```

## Predicted Probabilities and Differences by Lapse Status

```{r probability_table_emmeans}
# Try to compute emmeans on probability scale with error handling
compute_prob_table <- tryCatch({
  # Increase reference grid limit to handle all combinations
  emm_options(rg.limit = 200000)

  # Get predicted probabilities on the response scale
  # This will average over all covariate combinations
  emm_probs <- emmeans(model_dr_interaction, 
                       ~ ever_lapse_binary | dr_severity,
                       type = "response")

  # Convert to data frame for manipulation
  prob_summary <- as.data.frame(emm_probs)

  # Check if we have the expected structure
  if(nrow(prob_summary) == 0) {
    stop("emmeans returned empty results")
  }

  # Calculate differences manually from the emmeans object
  prob_contrasts <- contrast(emm_probs, method = "revpairwise", by = "dr_severity")
  prob_diff_emmeans <- as.data.frame(prob_contrasts)

  # Check structure before creating table
  n_dr_levels <- length(unique(prob_summary$dr_severity))
  expected_rows <- n_dr_levels * 2

  if(nrow(prob_summary) != expected_rows) {
    warning(paste("Expected", expected_rows, "rows but got", nrow(prob_summary)))
  }

  # Create a clean table with probabilities for each group
  # Reorganize the data more carefully
  prob_table <- data.frame()
  for(dr_level in unique(prob_summary$dr_severity)) {
    dr_data <- prob_summary[prob_summary$dr_severity == dr_level, ]
    if(nrow(dr_data) == 2) {
      prob_table <- rbind(prob_table,
                          data.frame(
                            DR_Severity = dr_level,
                            Prob_No_Lapse = dr_data$prob[dr_data$ever_lapse_binary == 0],
                            Prob_Lapse = dr_data$prob[dr_data$ever_lapse_binary == 1]
                          ))
    }
  }

  # Add differences if available
  if(nrow(prob_diff_emmeans) > 0 && nrow(prob_table) > 0) {
    prob_table$Difference = prob_diff_emmeans$estimate
    prob_table$CI_Lower = prob_diff_emmeans$asymp.LCL
    prob_table$CI_Upper = prob_diff_emmeans$asymp.UCL
    prob_table$P_Value = prob_diff_emmeans$p.value
  } else {
    prob_table$Difference = NA
    prob_table$CI_Lower = NA
    prob_table$CI_Upper = NA
    prob_table$P_Value = NA
  }

  TRUE  # Success indicator
}, error = function(e) {
  cat("Note: Could not compute emmeans probabilities on response scale.\n")
  cat("Error:", e$message, "\n")
  cat("This may be due to the large reference grid size.\n")
  cat("Using manual calculation method instead (see next section).\n\n")
  FALSE  # Failure indicator
})

# Only create and display table if computation succeeded
if(compute_prob_table == TRUE && exists("prob_table") && nrow(prob_table) > 0) {
  # Format the table for display - check which columns exist
  display_table <- prob_table |>
    mutate(
      Prob_No_Lapse = round(Prob_No_Lapse, 3),
      Prob_Lapse = round(Prob_Lapse, 3)
    )

  # Add optional columns if they exist
  if("Difference" %in% names(prob_table)) {
    display_table <- display_table |>
      mutate(Difference = round(Difference, 3))
  }

  if(all(c("CI_Lower", "CI_Upper") %in% names(prob_table))) {
    display_table <- display_table |>
      mutate(CI = paste0("(", round(CI_Lower, 3), ", ", round(CI_Upper, 3), ")"))
  }

  if("P_Value" %in% names(prob_table)) {
    display_table <- display_table |>
      mutate(P_Value = format.pval(P_Value, digits = 3, eps = 0.001))
  }

  # Select only columns that exist
  cols_to_select <- c("DR_Severity", "Prob_No_Lapse", "Prob_Lapse")
  if("Difference" %in% names(display_table)) cols_to_select <- c(cols_to_select, "Difference")
  if("CI" %in% names(display_table)) cols_to_select <- c(cols_to_select, "CI")
  if("P_Value" %in% names(display_table)) cols_to_select <- c(cols_to_select, "P_Value")

  display_table <- display_table |> select(all_of(cols_to_select))

  # Adjust column names based on what's available
  col_names <- c("DR Severity", "P(VI | No Lapse)", "P(VI | Lapse)")
  if("Difference" %in% names(display_table)) col_names <- c(col_names, "Difference")
  if("CI" %in% names(display_table)) col_names <- c(col_names, "95% CI")
  if("P_Value" %in% names(display_table)) col_names <- c(col_names, "P-value")

  knitr::kable(display_table,
               caption = "Predicted Probability of Vision Impairment by DR Severity and Lapse Status (Averaged Over All Covariate Patterns)",
               col.names = col_names,
               row.names = FALSE,
               digits = 3)
} else {
  cat("Skipping emmeans probability table due to computation issues.\n")
  cat("See manual calculation in the next section for probability differences.\n")
}
```

## Average Marginal Effects by DR Severity Using Margins Package

```{r margins_dr_severity}
# Calculate AMEs for each DR severity group using margins package
# This matches the approach in twang-analysis.Rmd

cat("Calculating Average Marginal Effects using margins package...\n\n")

# Overall AME
avg_margfx_overall <- margins(model_dr_interaction, design = survey_design_weighted, 
                              variables = "ever_lapse_binary")
mfx_overall <- summary(avg_margfx_overall)

# For each DR severity group
# No DR
design_no_dr <- subset(survey_design_weighted, dr_severity == "No_DR")
data_no_dr <- subset(m_data, dr_severity == "No_DR")
n_no_dr <- nrow(data_no_dr)
mfx_no_dr <- summary(margins(model_dr_interaction, 
                             design = design_no_dr, 
                             data = data_no_dr, 
                             variables = "ever_lapse_binary"))

# NPDR
design_npdr <- subset(survey_design_weighted, dr_severity == "NPDR")
data_npdr <- subset(m_data, dr_severity == "NPDR")
n_npdr <- nrow(data_npdr)
mfx_npdr <- summary(margins(model_dr_interaction, 
                            design = design_npdr, 
                            data = data_npdr, 
                            variables = "ever_lapse_binary"))

# PDR
design_pdr <- subset(survey_design_weighted, dr_severity == "PDR")
data_pdr <- subset(m_data, dr_severity == "PDR")
n_pdr <- nrow(data_pdr)
mfx_pdr <- summary(margins(model_dr_interaction, 
                           design = design_pdr, 
                           data = data_pdr, 
                           variables = "ever_lapse_binary"))

# Combine results into a table
ame_results <- data.frame(
  DR_Severity = c("Overall", "No_DR", "NPDR", "PDR"),
  N = c(nrow(m_data), n_no_dr, n_npdr, n_pdr),
  AME = c(mfx_overall$AME, mfx_no_dr$AME, mfx_npdr$AME, mfx_pdr$AME),
  SE = c(mfx_overall$SE, mfx_no_dr$SE, mfx_npdr$SE, mfx_pdr$SE),
  CI_Lower = c(mfx_overall$lower, mfx_no_dr$lower, mfx_npdr$lower, mfx_pdr$lower),
  CI_Upper = c(mfx_overall$upper, mfx_no_dr$upper, mfx_npdr$upper, mfx_pdr$upper),
  P_Value = c(mfx_overall$p, mfx_no_dr$p, mfx_npdr$p, mfx_pdr$p)
)

# Format for display
ame_display <- ame_results |>
  mutate(
    AME = round(AME, 4),
    SE = round(SE, 4),
    CI = paste0("(", round(CI_Lower, 4), ", ", round(CI_Upper, 4), ")"),
    P_Value = format.pval(P_Value, digits = 4, eps = 0.0001)
  ) |>
  select(DR_Severity, N, AME, SE, CI, P_Value)

knitr::kable(ame_display,
             caption = "Average Marginal Effects of Lapsing on Vision Impairment by DR Severity",
             col.names = c("DR Severity", "N", "AME", "SE", "95% CI", "P-value"),
             row.names = FALSE,
             digits = 4)

cat("\nNote: AME represents the average change in probability of vision impairment when lapsing occurs,")
cat("\naveraged over the covariate distribution within each DR severity group.")
cat("\nThis approach matches the methodology used in twang-analysis.Rmd.\n")
```

## Visualization of Average Marginal Effects

```{r ame_visualization, fig.height=6, fig.width=10}
# Plot the AMEs from margins package
if(exists("ame_results") && nrow(ame_results) > 0) {
  # Exclude overall from plot, focus on DR severity groups
  plot_data_ame <- ame_results[ame_results$DR_Severity != "Overall", ]

  # Create forest plot of AMEs
  ame_plot <- ggplot(plot_data_ame, aes(x = DR_Severity, y = AME)) +
    geom_point(size = 4, color = "darkblue") +
    geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
                  width = 0.15, size = 1, color = "darkblue") +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5, color = "gray50") +
    labs(title = "Average Marginal Effect of Lapsing on Vision Impairment by DR Severity",
         subtitle = "Change in probability of vision impairment (IPTW-weighted, with 95% CIs)",
         x = "DR Severity",
         y = "Average Marginal Effect (Probability Scale)") +
    theme_minimal() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          plot.subtitle = element_text(size = 11, color = "gray30"),
          axis.text = element_text(size = 11),
          axis.title = element_text(size = 12)) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) +
    # Add sample size annotations
    geom_text(aes(label = paste0("n=", format(N, big.mark=","))),
              vjust = -2.5, size = 3.5, color = "gray50")

  print(ame_plot)

  # Create a second plot comparing AME magnitudes
  ame_bar_plot <- ggplot(plot_data_ame, aes(x = reorder(DR_Severity, AME), y = AME, fill = DR_Severity)) +
    geom_col(alpha = 0.7) +
    geom_errorbar(aes(ymin = CI_Lower, ymax = CI_Upper), 
                  width = 0.2, size = 1) +
    geom_hline(yintercept = 0, linetype = "solid", alpha = 0.5) +
    coord_flip() +
    labs(title = "Magnitude of Lapse Effect Across DR Severity",
         subtitle = "Average marginal effects with 95% confidence intervals",
         x = "",
         y = "Average Marginal Effect") +
    theme_minimal() +
    theme(plot.title = element_text(size = 14, face = "bold"),
          axis.text = element_text(size = 11),
          axis.title = element_text(size = 12),
          legend.position = "none") +
    scale_fill_manual(values = c("No_DR" = "#1f77b4", 
                                 "NPDR" = "#ff7f0e", 
                                 "PDR" = "#d62728")) +
    scale_y_continuous(labels = scales::number_format(accuracy = 0.001)) +
    geom_text(aes(label = sprintf("%.3f", AME)),
              hjust = ifelse(plot_data_ame$AME > 0, -0.2, 1.2),
              size = 3.5)

  print(ame_bar_plot)
} else {
  cat("AME results not available for visualization.\n")
}
```

## Complete Coefficient Interpretation

```{r coefficient_interpretation}
# Extract all coefficients for complete interpretation
all_coefs <- summary_dr$coefficients

# Create interpretation table for main effects and interactions
main_dr_coefs <- c("ever_lapse_binary", "dr_severityNPDR", "dr_severityPDR",
                   dr_interaction_terms)

interp_data <- data.frame(
  Term = character(),
  Estimate = numeric(),
  Interpretation = character(),
  stringsAsFactors = FALSE
)

# Main effect of lapse
if("ever_lapse_binary" %in% rownames(all_coefs)) {
  interp_data <- rbind(interp_data, data.frame(
    Term = "ever_lapse_binary",
    Estimate = round(all_coefs["ever_lapse_binary", "Estimate"], 4),
    Interpretation = "Effect of lapsing for No DR patients (reference group)"
  ))
}

# DR severity main effects
if("dr_severityNPDR" %in% rownames(all_coefs)) {
  interp_data <- rbind(interp_data, data.frame(
    Term = "dr_severityNPDR",
    Estimate = round(all_coefs["dr_severityNPDR", "Estimate"], 4),
    Interpretation = "Baseline difference: NPDR vs No DR (among non-lapsers)"
  ))
}

if("dr_severityPDR" %in% rownames(all_coefs)) {
  interp_data <- rbind(interp_data, data.frame(
    Term = "dr_severityPDR",
    Estimate = round(all_coefs["dr_severityPDR", "Estimate"], 4),
    Interpretation = "Baseline difference: PDR vs No DR (among non-lapsers)"
  ))
}

# Interaction terms
for(term in dr_interaction_terms) {
  if(term %in% rownames(all_coefs)) {
    dr_level <- gsub("ever_lapse_binary:dr_severity", "", term)
    interp_data <- rbind(interp_data, data.frame(
      Term = term,
      Estimate = round(all_coefs[term, "Estimate"], 4),
      Interpretation = paste("Additional effect of lapsing in", dr_level, "vs No DR")
    ))
  }
}

knitr::kable(interp_data,
             caption = "Interpretation of Model Coefficients",
             row.names = FALSE)
```

# Summary

## Key Findings

A formal interaction analysis provides evidence that **==he effect of lapsing from care on vision impairment depends significantly on patients' baseline DR severity (joint interaction test p < 0.05).

### Substantive Results:

1. **Patients with No DR (n=16,130)**: Lapsing from care increases the odds of vision impairment by approximately 17% (OR = 1.17, log odds = 0.160). This effect is statistically significant.

2. **Patients with NPDR (n=5,695)**: Lapsing has a substantially stronger effect, increasing the odds of vision impairment by approximately 47% (OR = 1.47, log odds = 0.386). This effect is also statistically significant and notably larger than in the No DR group (although the error bars do cross over)

3. **Patients with PDR (n=3,187)**: Lapsing shows minimal effect on vision impairment risk (OR ≈ 1.01, log odds = 0.006), and this effect is not statistically significant.

### Response to Reviewer:

This analysis directly addresses the reviewer's concern about "inadequate subgroup testing" by:

- Providing formal statistical tests (joint Wald test) rather than relying on visual overlap alone
- Calculating group-specific effects with proper confidence intervals
- Demonstrating that the interaction is statistically significant (p < 0.05)
- Showing that subgroup effects are meaningfully different, not just statistically different

The significant interaction confirms that a "one-size-fits-all" interpretation of lapsing effects would be misleading. The surprising finding that PDR patients show essentially no harm from lapsing (OR ≈ 1.01) suggests we may need to **reconsider our conceptual understanding of lapsing in advanced disease**. While NPDR patients show substantial risk from care disruption, the null effect in PDR raises questions about whether our retention efforts should be differentially targeted based on disease severity.