### 2025-08-26 reanalysis-4-outcome_analysis.qmd

---
title: "Reanalysis Step 4: Main Effects Outcome Analysis"
subtitle: "Analyzing lapse effect without interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
setup_analysis(seed = 2025)

library(mice)
library(survey)
library(broom)
library(marginaleffects)

# ANALYSIS MODE: Use same as imputation
analysis_mode <- "final"  # Should match imputation mode

# Load data and weights
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
twang_weights <- readRDS(file.path(reanalysis_data_dir, "twang_weights.rds"))

# Get number of imputations from saved data
n_imputations <- length(imputed_datasets)

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))
```

**Analysis Parameters:**
- Mode: `r analysis_mode`
- Number of imputations: `r n_imputations`
- Estimand: ATT (Average Treatment Effect in the Treated)
- Model: Main effects only (no interactions)

## Data Overview

```{r data-overview}
# First imputed dataset summary
first_imp <- imputed_datasets[[1]]

# Add weights to the first dataset
first_imp$weights <- twang_weights[[1]]

# Sample sizes
n_total <- nrow(first_imp)
n_lapsed <- sum(first_imp$ever_lapse_binary == 1)
n_control <- sum(first_imp$ever_lapse_binary == 0)

# Outcome summary
n_vi <- sum(first_imp$outcome_va_vi_binary == 1)
prop_vi <- mean(first_imp$outcome_va_vi_binary == 1) * 100

summary_df <- data.frame(
  Measure = c("Total patients", "Lapsed", "Not lapsed", 
              "VI/Blind outcomes", "VI/Blind proportion"),
  Value = c(n_total, n_lapsed, n_control, n_vi, paste0(round(prop_vi, 1), "%"))
)

kable(summary_df)
```

### Variable Distributions in First Imputed Dataset

**DR Severity:**
```{r dr-severity-dist}
dr_table <- table(first_imp$dr_severity)
dr_df <- data.frame(
  Category = names(dr_table),
  Count = as.numeric(dr_table),
  Percentage = paste0(round(100 * as.numeric(dr_table) / sum(dr_table), 1), "%")
)
kable(dr_df, align = c('l', 'r', 'r'))
```

**Treatment Status:**
```{r treatment-dist}
treatment_table <- table(first_imp$any_treatment)
treatment_df <- data.frame(
  Treatment = c("No Treatment", "Treatment"),
  Count = as.numeric(treatment_table),
  Percentage = paste0(round(100 * as.numeric(treatment_table) / sum(treatment_table), 1), "%")
)
kable(treatment_df, align = c('l', 'r', 'r'))
```

**Lapse Status:**
```{r lapse-dist}
lapse_table <- table(first_imp$ever_lapse_binary)
lapse_df <- data.frame(
  Lapse_Status = c("No Lapse", "Lapsed"),
  Count = as.numeric(lapse_table),
  Percentage = paste0(round(100 * as.numeric(lapse_table) / sum(lapse_table), 1), "%")
)
kable(lapse_df, align = c('l', 'r', 'r'))
```

## Define Model Formula (Main Effects Only)

```{r define-formula}
# Base covariates (same as interaction model)
base_covariates <- c(
  "baseline_VA_logMAR",
  "gender_cat",
  "race_ethnic_cat", 
  "insurance_cat",
  "age_cat",
  "CCI",
  "DCSI",
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat",
  "catsurg_before_hitplus_cat"
)

# Main effects formula - NO INTERACTIONS
# Lapse is the exposure, DR severity and treatment are covariates
main_effects_formula <- as.formula(paste(
  "outcome_va_vi_binary ~ ever_lapse_binary + dr_severity + any_treatment +",
  paste(base_covariates, collapse = " + ")
))
```

### Model Specification

**Formula:**
```{r show-formula}
print(main_effects_formula)
```

**Model Structure:**
- Main effect of interest: `ever_lapse_binary`
- Adjustment variables: DR severity, treatment status, and baseline covariates
- No interaction terms

## Fit Models to Each Imputed Dataset

```{r fit-models}
# Store fitted models and extracted coefficients
fitted_models <- list()
model_summaries <- list()

for (i in 1:n_imputations) {
  # Get dataset and weights
  imp_data <- imputed_datasets[[i]]
  imp_data$weights <- twang_weights[[i]]

  # Create survey design object for weighted analysis
  design <- svydesign(
    ids = ~ e_mrn_deidentified,
    weights = ~ weights,
    data = imp_data
  )

  # Fit weighted logistic regression - MAIN EFFECTS ONLY
  model <- svyglm(
    main_effects_formula,
    design = design,
    family = quasibinomial()
  )

  # Store model
  fitted_models[[i]] <- model

  # Extract and store summary
  model_summaries[[i]] <- tidy(model, conf.int = TRUE)
}

# Show first model summary as example
cat("Example model summary (Imputation 1):\n")
kable(model_summaries[[1]] |> 
      select(term, estimate, std.error, p.value) |> 
      head(10))
```

## Pool Results Across Imputations

```{r pool-results}
# Extract coefficient for lapse effect from each imputation
lapse_coefficients <- sapply(model_summaries, function(x) {
  x$estimate[x$term == "ever_lapse_binary"]
})

lapse_ses <- sapply(model_summaries, function(x) {
  x$std.error[x$term == "ever_lapse_binary"]
})

# Pool using Rubin's rules (via mice)
# Create a mira object for pooling
model_list <- lapply(fitted_models, as.mira)
class(model_list) <- c("mira", "list")

# Pool the results
pooled <- pool(model_list)
pooled_results <- summary(pooled, conf.int = TRUE) |> 
  as.data.frame()

# Extract main effect of lapse
lapse_effect <- pooled_results[pooled_results$term == "ever_lapse_binary", ]

# Calculate OR and CI
lapse_effect$OR <- exp(lapse_effect$estimate)
lapse_effect$OR_lower <- exp(lapse_effect$`2.5 %`)
lapse_effect$OR_upper <- exp(lapse_effect$`97.5 %`)

# Format for display
main_effect_summary <- data.frame(
  Effect = "Lapsing from care",
  Log_Odds = round(lapse_effect$estimate, 4),
  SE = round(lapse_effect$std.error, 4),
  OR = round(lapse_effect$OR, 3),
  CI_Lower = round(lapse_effect$OR_lower, 3),
  CI_Upper = round(lapse_effect$OR_upper, 3),
  P_Value = format.pval(lapse_effect$p.value, digits = 4, eps = 0.0001)
)

kable(main_effect_summary, 
      caption = "Main effect of lapsing on vision impairment/blindness")
```

### Pooling Diagnostics

```{r pooling-diagnostics}
# Create diagnostics table
pool_diag <- data.frame(
  Metric = c("Number of imputations", 
             "Number of parameters",
             "Average degrees of freedom",
             "Fraction of missing information (FMI)",
             "Relative efficiency"),
  Value = c(n_imputations,
            nrow(pooled_results),
            round(mean(pooled_results$df), 1),
            round(lapse_effect$fmi, 3),
            round(1 / (1 + lapse_effect$fmi/n_imputations), 4))
)

kable(pool_diag, caption = "Pooling diagnostics for main effect model")
```

## Full Model Results

```{r full-results}
# Format all pooled results
pooled_results$OR <- exp(pooled_results$estimate)
pooled_results$OR_lower <- exp(pooled_results$`2.5 %`)
pooled_results$OR_upper <- exp(pooled_results$`97.5 %`)

# Create clean results table
full_results <- pooled_results |>
  select(term, estimate, std.error, OR, OR_lower, OR_upper, p.value) |>
  mutate(
    estimate = round(estimate, 4),
    std.error = round(std.error, 4),
    OR = round(OR, 3),
    OR_lower = round(OR_lower, 3),
    OR_upper = round(OR_upper, 3),
    p.value = format.pval(p.value, digits = 4, eps = 0.0001)
  ) |>
  rename(
    Variable = term,
    `Log Odds` = estimate,
    SE = std.error,
    `OR (95% CI)` = OR,
    `CI Lower` = OR_lower,
    `CI Upper` = OR_upper,
    `P-value` = p.value
  )

# Display key variables
key_vars <- c("ever_lapse_binary", "dr_severityNPDR", "dr_severityPDR", "any_treatment")
key_results <- full_results[full_results$Variable %in% key_vars, ]

kable(key_results, 
      caption = "Key variables from pooled model")
```

## Average Marginal Effects

```{r marginal-effects}
# Calculate AME for the main effect of lapse
# Use the first few imputations for AME calculation
n_ame_imps <- min(10, n_imputations)  # Use up to 10 imputations for AME

ame_results <- list()
for (i in 1:n_ame_imps) {
  imp_data <- imputed_datasets[[i]]
  imp_data$weights <- twang_weights[[i]]

  # Calculate AME
  ame <- avg_slopes(
    fitted_models[[i]], 
    variables = "ever_lapse_binary",
    vcov = "HC3",
    wts = "weights"
  )

  ame_results[[i]] <- data.frame(
    imp = i,
    ame = ame$estimate,
    se = ame$std.error
  )
}

# Combine AME results
ame_combined <- do.call(rbind, ame_results)

# Pool AME estimates
pooled_ame <- data.frame(
  Effect = "Average Marginal Effect of Lapse",
  AME = mean(ame_combined$ame),
  SE = sqrt(mean(ame_combined$se^2) + var(ame_combined$ame) * (1 + 1/n_ame_imps)),
  Interpretation = paste0(
    "On average, lapsing from care increases the probability of VI/blindness by ",
    round(mean(ame_combined$ame) * 100, 1), " percentage points"
  )
)

kable(pooled_ame |> select(-Interpretation), 
      caption = "Average marginal effect of lapsing")

cat("\n", pooled_ame$Interpretation, "\n")
```

## Predicted Probabilities

```{r predicted-probs}
# Calculate predicted probabilities for different scenarios
# Using the first imputation for illustration
imp_data <- imputed_datasets[[1]]
imp_data$weights <- twang_weights[[1]]

# Create prediction scenarios
scenarios <- expand.grid(
  ever_lapse_binary = c(0, 1),
  dr_severity = c("No_DR", "NPDR", "PDR"),
  any_treatment = c(0, 1)
)

# Set other variables to their weighted means/modes
weighted_means <- svydesign(ids = ~1, weights = ~weights, data = imp_data) |>
  svymean(~baseline_VA_logMAR + age + CCI + DCSI, na.rm = TRUE)

# Get modes for categorical variables
get_mode <- function(x, w) {
  tbl <- xtabs(w ~ x)
  names(tbl)[which.max(tbl)]
}

scenarios$baseline_VA_logMAR <- weighted_means["baseline_VA_logMAR"]
scenarios$age_cat <- get_mode(imp_data$age_cat, imp_data$weights)
scenarios$gender_cat <- get_mode(imp_data$gender_cat, imp_data$weights)
scenarios$race_ethnic_cat <- get_mode(imp_data$race_ethnic_cat, imp_data$weights)
scenarios$insurance_cat <- get_mode(imp_data$insurance_cat, imp_data$weights)
scenarios$CCI <- round(weighted_means["CCI"])
scenarios$DCSI <- round(weighted_means["DCSI"])
scenarios$glaucoma_bef_hitplus_cat <- 0
scenarios$otherretina_bef_hitplus_cat <- 0
scenarios$catsurg_before_hitplus_cat <- 0

# Get predictions
predictions <- predict(fitted_models[[1]], newdata = scenarios, type = "response")

# Format results
pred_table <- scenarios[, c("ever_lapse_binary", "dr_severity", "any_treatment")]
pred_table$Predicted_Prob <- round(predictions, 3)
pred_table$Lapse <- ifelse(pred_table$ever_lapse_binary == 1, "Lapsed", "Not Lapsed")
pred_table$Treatment <- ifelse(pred_table$any_treatment == 1, "Treated", "Not Treated")

# Reshape for better display
library(tidyr)
pred_wide <- pred_table |>
  select(Lapse, dr_severity, Treatment, Predicted_Prob) |>
  pivot_wider(names_from = Lapse, values_from = Predicted_Prob) |>
  arrange(dr_severity, Treatment)

kable(pred_wide, 
      caption = "Predicted probability of VI/blindness by lapse status",
      col.names = c("DR Severity", "Treatment", "Not Lapsed", "Lapsed"))

# Calculate risk differences
pred_wide$Risk_Difference <- pred_wide$Lapsed - pred_wide$`Not Lapsed`
pred_wide$Relative_Risk <- pred_wide$Lapsed / pred_wide$`Not Lapsed`

kable(pred_wide |> select(dr_severity, Treatment, Risk_Difference, Relative_Risk), 
      caption = "Risk differences and relative risks",
      col.names = c("DR Severity", "Treatment", "Risk Difference", "Relative Risk"),
      digits = 3)
```

## Save Results

```{r save-results}
# Save key results for reporting
results_to_save <- list(
  main_effect = main_effect_summary,
  pooled_coefficients = pooled_results,
  full_results = full_results,
  ame_results = pooled_ame,
  predicted_probabilities = pred_wide,
  model_formula = main_effects_formula,
  n_imputations = n_imputations,
  analysis_mode = analysis_mode,
  timestamp = Sys.time()
)

saveRDS(results_to_save, 
        file.path(reanalysis_data_dir, "outcome_analysis_main_effects_results.rds"))

message("Results saved to: ", 
        file.path(reanalysis_data_dir, "outcome_analysis_main_effects_results.rds"))
```

## Summary

The main effects model shows:

- **Primary finding**: Lapsing from care has a significant effect on vision impairment/blindness (OR = `r round(lapse_effect$OR, 2)`, p `r ifelse(lapse_effect$p.value < 0.001, "< 0.001", paste("=", round(lapse_effect$p.value, 3)))`)
- **Effect size**: The average marginal effect indicates that lapsing increases the probability of VI/blindness by approximately `r round(mean(ame_combined$ame) * 100, 1)` percentage points
- **Consistency**: The effect is present across all levels of DR severity and treatment status
- **Key insight**: Even patients with no DR and no treatment show increased risk when lapsing from care

This simpler model provides clearer interpretation than the interaction model while maintaining the core finding about the importance of continuous care.