### 2025-08-28 reanalysis-3-propensity_scores-with-caching.qmd

---
title: "Reanalysis Step 3: Propensity Score Estimation (With Caching)"
subtitle: "Estimating propensity scores using twang for multiple imputations"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
setup_analysis(seed = 2025)

library(twang)
library(cobalt)
library(ggplot2)
library(gridExtra)

# ANALYSIS MODE
analysis_mode <- "final"  # "test" or "final"

# CACHING CONFIGURATION
use_cache <- TRUE  # Set to FALSE to force re-computation
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
if (!dir.exists(cache_dir)) {
  dir.create(cache_dir, recursive = TRUE)
}
```

## Load Data

```{r load-data}
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_available <- length(imputed_datasets)

# Determine how many imputations to process
if (analysis_mode == "test") {
  n_imputations <- min(2, n_available)
} else {
  n_imputations <- n_available
}

variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))
```

## Define Propensity Score Formula

```{r ps-formula}
ps_formula_original <- get_matching_formula()
ps_formula <- ps_formula_original
print(ps_formula)
```

## Run Twang Models with Caching

```{r fit-twang}
# Model parameters
n_trees <- ifelse(analysis_mode == "test", 1000, 3000)
stop_methods <- c("es.mean", "ks.max")

# Initialize results storage
twang_results <- list()
balance_summaries <- list()

# Process each imputation with caching
for (i in 1:n_imputations) {
  cat(paste0("\n--- Processing Imputation ", i, "/", n_imputations, " ---\n"))

  # Check for cached result
  cache_file <- file.path(cache_dir, paste0("twang_imp_", i, ".rds"))

  if (use_cache && file.exists(cache_file)) {
    cat("  Loading from cache...\n")
    twang_results[[i]] <- readRDS(cache_file)
  } else {
    cat("  Fitting twang model...\n")

    # Get imputed dataset
    imp_data <- imputed_datasets[[i]]

    # Fit twang model
    twang_model <- ps(
      formula = ps_formula,
      data = imp_data,
      n.trees = n_trees,
      interaction.depth = 3,
      shrinkage = 0.01,
      estimand = "ATT",
      stop.method = stop_methods,
      n.minobsinnode = 10,
      verbose = FALSE
    )

    # Extract weights
    weights <- get.weights(twang_model, stop.method = "es.mean")

    # Store results
    result <- list(
      model = twang_model,
      weights = weights,
      n_trees_used = twang_model$desc$es.mean$n.trees
    )

    twang_results[[i]] <- result

    # Save to cache
    saveRDS(result, cache_file)
    cat(paste0("  Saved to cache: ", cache_file, "\n"))
  }

  # Calculate balance (always, even for cached results)
  imp_data <- imputed_datasets[[i]]
  imp_data$weights <- twang_results[[i]]$weights

  balance_obj <- bal.tab(
    ps_formula,
    data = imp_data,
    weights = imp_data$weights,
    estimand = "ATT",
    s.d.denom = "treated",
    stats = c("m", "v"),
    thresholds = c(m = 0.1, v = 2)
  )

  balance_summaries[[i]] <- balance_obj

  cat(paste0("  Trees used: ", twang_results[[i]]$n_trees_used, "\n"))
  cat(paste0("  Max SMD: ", round(max(abs(balance_obj$Balance$Diff.Adj), na.rm = TRUE), 3), "\n"))
}

cat("\n✓ All imputations processed\n")
```

## Save Final Results

```{r save-results}
# Extract weights for easy access
twang_weights <- list()
for (i in 1:n_imputations) {
  twang_weights[[i]] <- twang_results[[i]]$weights
}

# Create output object
ps_output <- list(
  twang_results = twang_results,
  weight_summary = balance_summaries,
  formula_used = ps_formula,
  n_imputations = n_imputations,
  method = "twang_gbm",
  n_trees = n_trees,
  analysis_mode = analysis_mode
)

# Save main outputs
saveRDS(ps_output, file.path(reanalysis_data_dir, "ps_results_twang.rds"))
saveRDS(twang_weights, file.path(reanalysis_data_dir, "twang_weights.rds"))

cat("\nResults saved to:\n")
cat("  -", file.path(reanalysis_data_dir, "ps_results_twang.rds"), "\n")
cat("  -", file.path(reanalysis_data_dir, "twang_weights.rds"), "\n")
```

## Summary

```{r summary}
cat(paste0("\nSuccessfully processed ", n_imputations, " imputations\n"))
cat(paste0("Cache directory: ", cache_dir, "\n"))
cat("\nTo re-run without cache, set use_cache <- FALSE\n")
```