### 2025-08-29 reanalysis-2b-imputation_diagnostics.qmd
enhance

---
title: "Reanalysis Step 2b: Multiple Imputation Outcome Diagnostics"
subtitle: "Assessing plausibility of imputed visual acuity outcomes"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 6
---

## Overview

This notebook examines whether our multiple imputation procedure for visual acuity outcomes is producing clinically plausible values. We're particularly concerned about:

1. **Temporal consistency**: Do imputed 2-year outcomes respect the trajectory from baseline through intermediate timepoints?
2. **Clinical plausibility**: Are changes in vision realistic given what we know about disease progression?
3. **Distribution similarity**: Do imputed values have similar distributions to observed values?

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
setup_analysis(seed = 2025)

library(mice)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(knitr)

# Initialize logging
log_file <- init_log("imputation_diagnostics")
```

## Load Data

```{r load-data}
# Load the original data to identify which values were imputed
df_original <- readRDS(file.path(reanalysis_data_dir, "df_for_imputation.rds"))

# Load the imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

log_message(paste0("Loaded ", n_imputations, " imputed datasets"))
log_message(paste0("Total observations: ", nrow(df_original)))
```

## Identify Imputed vs Observed Outcomes

```{r identify-imputed}
# In the original data, cohort_id tells us who had observed outcomes
# "submitted" = has true 2-year outcome
# "excluded" = lacks true 2-year outcome (these were imputed)

# Create more nuanced grouping
df_original <- df_original |>
  mutate(
    outcome_was_imputed = cohort_id == "excluded",
    had_intermediate = !is.na(intermediate_VA_logMAR),

    # Create detailed imputation groups
    imputation_group = case_when(
      !outcome_was_imputed ~ "Observed",
      outcome_was_imputed & had_intermediate ~ "Imputed (with intermediate)",
      outcome_was_imputed & !had_intermediate ~ "Imputed (no intermediate)",
      TRUE ~ "Unknown"
    ),
    imputation_group = factor(imputation_group, 
                             levels = c("Observed", 
                                       "Imputed (with intermediate)", 
                                       "Imputed (no intermediate)"))
  )

# Summary of imputation status
imputation_summary <- df_original |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    pct = round(100 * n / nrow(df_original), 1),
    had_baseline = sum(!is.na(baseline_VA_logMAR)),
    pct_baseline = round(100 * had_baseline / n, 1),
    .groups = "drop"
  )

kable(imputation_summary,
      col.names = c("Group", "N", "%", "Has Baseline", "% Baseline"),
      caption = "Detailed breakdown of imputation groups")

# Baseline characteristics by group
baseline_chars <- df_original |>
  group_by(imputation_group) |>
  summarise(
    # Demographics
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    pct_male = round(100 * mean(gender == "Male", na.rm = TRUE), 1),

    # Clinical characteristics
    mean_CCI = mean(CCI, na.rm = TRUE),
    sd_CCI = sd(CCI, na.rm = TRUE),
    mean_DCSI = mean(DCSI, na.rm = TRUE),
    sd_DCSI = sd(DCSI, na.rm = TRUE),

    # Baseline vision
    mean_baseline_VA = mean(baseline_VA_logMAR, na.rm = TRUE),
    sd_baseline_VA = sd(baseline_VA_logMAR, na.rm = TRUE),

    # Treatment indicators
    pct_anti_VEGF = round(100 * mean(anti_VEGF == 1, na.rm = TRUE), 1),
    pct_PRP = round(100 * mean(PRP_flag == 1, na.rm = TRUE), 1),

    # Lapse status
    pct_lapsed = round(100 * mean(ever_lapse_binary == 1, na.rm = TRUE), 1),

    .groups = "drop"
  )

kable(baseline_chars |> 
        mutate(across(where(is.numeric), ~round(., 2))),
      caption = "Baseline characteristics by imputation group")
```

## Calculate Change Scores

```{r calculate-changes}
# Function to calculate change scores for a single imputed dataset
calculate_change_scores <- function(imp_data, original_data) {
  imp_data |>
    mutate(
      # Add imputation status from original data
      outcome_was_imputed = original_data$outcome_was_imputed,
      had_intermediate = original_data$had_intermediate,
      imputation_group = original_data$imputation_group,

      # Calculate changes
      change_baseline_to_outcome = outcome_VA_logMAR - baseline_VA_logMAR,
      change_baseline_to_intermediate = intermediate_VA_logMAR - baseline_VA_logMAR,
      change_intermediate_to_outcome = outcome_VA_logMAR - intermediate_VA_logMAR,

      # Categorize changes (negative = improvement, positive = worsening)
      # Using 0.1 logMAR as clinically meaningful change (approximately 1 line on eye chart)
      outcome_direction = case_when(
        change_baseline_to_outcome < -0.1 ~ "Improved",
        change_baseline_to_outcome > 0.1 ~ "Worsened",
        TRUE ~ "Stable"
      ),

      # Flag extreme changes (clinically implausible)
      extreme_improvement = change_baseline_to_outcome < -1.0,  # >10 lines improvement
      extreme_worsening = change_baseline_to_outcome > 1.5,      # >15 lines worsening

      # For those with intermediate values, check trajectory consistency
      trajectory_reversal = case_when(
        !is.na(change_baseline_to_intermediate) & !is.na(change_intermediate_to_outcome) ~
          sign(change_baseline_to_intermediate) != sign(change_intermediate_to_outcome),
        TRUE ~ FALSE
      )
    )
}

# Calculate for all imputed datasets
all_changes <- list()
for (i in 1:n_imputations) {
  all_changes[[i]] <- calculate_change_scores(imputed_datasets[[i]], df_original)
}

# For summary statistics, use the first imputation as representative
changes_df <- all_changes[[1]]
```

## Compare Change Distributions

### Summary Statistics

```{r summary-stats}
# Compare change scores across all three groups
change_summary_detailed <- changes_df |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    q25_change = quantile(change_baseline_to_outcome, 0.25, na.rm = TRUE),
    q75_change = quantile(change_baseline_to_outcome, 0.75, na.rm = TRUE),
    min_change = min(change_baseline_to_outcome, na.rm = TRUE),
    max_change = max(change_baseline_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    iqr_change = q75_change - q25_change
  )

kable(change_summary_detailed |> 
        select(imputation_group, n, mean_change, sd_change, median_change, iqr_change, min_change, max_change),
      digits = 3,
      col.names = c("Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
      caption = "Distribution of vision changes (outcome - baseline) by detailed imputation group")

# Also keep simplified version for compatibility
change_summary <- changes_df |>
  group_by(outcome_was_imputed) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    q25_change = quantile(change_baseline_to_outcome, 0.25, na.rm = TRUE),
    q75_change = quantile(change_baseline_to_outcome, 0.75, na.rm = TRUE),
    min_change = min(change_baseline_to_outcome, na.rm = TRUE),
    max_change = max(change_baseline_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    iqr_change = q75_change - q25_change,
    group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Statistical tests across three groups
anova_test <- aov(change_baseline_to_outcome ~ imputation_group, data = changes_df)
anova_summary <- summary(anova_test)

# Pairwise comparisons
pairwise_tests <- pairwise.t.test(changes_df$change_baseline_to_outcome, 
                                  changes_df$imputation_group,
                                  p.adjust.method = "bonferroni")

# Format ANOVA results
anova_results <- data.frame(
  Test = "One-way ANOVA",
  F_Statistic = anova_summary[[1]]$`F value`[1],
  df1 = anova_summary[[1]]$Df[1],
  df2 = anova_summary[[1]]$Df[2],
  P_Value = anova_summary[[1]]$`Pr(>F)`[1],
  Interpretation = ifelse(anova_summary[[1]]$`Pr(>F)`[1] < 0.05, 
                         "Groups differ significantly", "No significant difference")
)

kable(anova_results, digits = 4, caption = "ANOVA comparing change distributions across three groups")

# Format pairwise results
pairwise_matrix <- pairwise_tests$p.value
pairwise_df <- data.frame(
  Comparison = c("Imputed (with intermediate) vs Observed",
                "Imputed (no intermediate) vs Observed",
                "Imputed (no intermediate) vs Imputed (with intermediate)"),
  P_Value = c(pairwise_matrix[1,1], pairwise_matrix[2,1], pairwise_matrix[2,2]),
  Significant = ifelse(c(pairwise_matrix[1,1], pairwise_matrix[2,1], pairwise_matrix[2,2]) < 0.05,
                      "Yes", "No")
)

kable(pairwise_df, digits = 4, caption = "Pairwise comparisons (Bonferroni adjusted)")

# Also do simplified tests for backward compatibility
t_test <- t.test(change_baseline_to_outcome ~ outcome_was_imputed, data = changes_df)
ks_test <- ks.test(
  changes_df$change_baseline_to_outcome[changes_df$outcome_was_imputed],
  changes_df$change_baseline_to_outcome[!changes_df$outcome_was_imputed]
)
```

### Visualization of Change Distributions

```{r change-distributions}
#| fig-height: 8
#| fig-width: 12

# Prepare data for plotting - using detailed groups
plot_data_detailed <- changes_df |>
  filter(!is.na(change_baseline_to_outcome))

# Also prepare simplified version for some plots
plot_data <- changes_df |>
  mutate(
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed"),
    Group = factor(Group, levels = c("Observed", "Imputed"))
  ) |>
  filter(!is.na(change_baseline_to_outcome))

# 1. Density plots - detailed groups
p1_detailed <- ggplot(plot_data_detailed, aes(x = change_baseline_to_outcome, fill = imputation_group)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(data = change_summary_detailed, 
             aes(xintercept = mean_change, color = imputation_group),
             linetype = "solid", size = 1) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", 
                              "Imputed (with intermediate)" = "#A23B72",
                              "Imputed (no intermediate)" = "#F18F01")) +
  scale_color_manual(values = c("Observed" = "#2E86AB", 
                               "Imputed (with intermediate)" = "#A23B72",
                               "Imputed (no intermediate)" = "#F18F01")) +
  labs(
    title = "Distribution of Vision Changes by Detailed Group",
    subtitle = "Vertical lines show group means",
    x = "Change in logMAR (positive = worsening)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 1b. Simplified density plot (for comparison)
p1 <- ggplot(plot_data, aes(x = change_baseline_to_outcome, fill = Group)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(data = change_summary, 
             aes(xintercept = mean_change, color = group),
             linetype = "solid", size = 1) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = "Distribution of Vision Changes (Simplified)",
    subtitle = "Vertical lines show group means",
    x = "Change in logMAR (positive = worsening)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 2. Boxplots - detailed groups
p2_detailed <- ggplot(plot_data_detailed, aes(x = imputation_group, y = change_baseline_to_outcome, fill = imputation_group)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", 
                              "Imputed (with intermediate)" = "#A23B72",
                              "Imputed (no intermediate)" = "#F18F01")) +
  labs(
    title = "Vision Changes by Detailed Group",
    subtitle = "Box shows IQR, whiskers extend to 1.5*IQR",
    y = "Change in logMAR (positive = worsening)",
    x = ""
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Q-Q plot
observed_changes <- plot_data$change_baseline_to_outcome[plot_data$Group == "Observed"]
imputed_changes <- plot_data$change_baseline_to_outcome[plot_data$Group == "Imputed"]

# Sample to same size for Q-Q plot
n_min <- min(length(observed_changes), length(imputed_changes))
qq_data <- data.frame(
  observed = sort(sample(observed_changes, n_min)),
  imputed = sort(sample(imputed_changes, n_min))
)

p3 <- ggplot(qq_data, aes(x = observed, y = imputed)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Q-Q Plot: Imputed vs Observed Changes",
    subtitle = "Points should follow red line if distributions are similar",
    x = "Observed Change Quantiles",
    y = "Imputed Change Quantiles"
  ) +
  theme_minimal()

# 4. Cumulative distribution
p4 <- ggplot(plot_data, aes(x = change_baseline_to_outcome, color = Group)) +
  stat_ecdf(size = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = "Cumulative Distribution of Changes",
    x = "Change in logMAR",
    y = "Cumulative Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Arrange plots - show detailed version first
grid.arrange(p1_detailed, p2_detailed, ncol = 2)

# Then show additional comparisons
grid.arrange(p3, p4, ncol = 2)
```

## Directional Change Analysis

```{r directional-changes}
# Compare proportions across detailed groups
direction_summary_detailed <- changes_df |>
  filter(!is.na(outcome_direction)) |>
  group_by(imputation_group, outcome_direction) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(imputation_group) |>
  mutate(
    total = sum(n),
    pct = round(100 * n / total, 1)
  )

# Reshape for display
direction_wide_detailed <- direction_summary_detailed |>
  select(imputation_group, outcome_direction, n, pct) |>
  pivot_wider(
    names_from = outcome_direction,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = 0)
  )

kable(direction_wide_detailed,
      caption = "Direction of vision change by detailed imputation group")

# Also create simplified version for backward compatibility
direction_summary <- changes_df |>
  filter(!is.na(outcome_direction)) |>
  group_by(outcome_was_imputed, outcome_direction) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(outcome_was_imputed) |>
  mutate(
    total = sum(n),
    pct = round(100 * n / total, 1),
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Reshape for display
direction_wide <- direction_summary |>
  select(Group, outcome_direction, n, pct) |>
  pivot_wider(
    names_from = outcome_direction,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = 0)
  )

kable(direction_wide,
      caption = "Direction of vision change by imputation status")

# Chi-square test for independence
direction_table <- table(changes_df$outcome_was_imputed, changes_df$outcome_direction)
chi_test <- chisq.test(direction_table)

chi_result <- data.frame(
  Test = "Chi-square test of independence",
  Chi_Square = chi_test$statistic,
  df = chi_test$parameter,
  P_Value = chi_test$p.value,
  Interpretation = ifelse(chi_test$p.value < 0.05, 
                         "Proportions differ between groups",
                         "Proportions similar between groups")
)

kable(chi_result, digits = 4)

# Visualization
direction_plot_data <- direction_summary |>
  mutate(outcome_direction = factor(outcome_direction, 
                                   levels = c("Improved", "Stable", "Worsened")))

p_direction <- ggplot(direction_plot_data, 
                     aes(x = Group, y = pct, fill = outcome_direction)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(pct, "%")), 
           position = position_dodge(width = 0.9),
           vjust = -0.5) +
  scale_fill_manual(values = c("Improved" = "#2ca02c", 
                              "Stable" = "#ff7f0e", 
                              "Worsened" = "#d62728")) +
  labs(
    title = "Direction of Vision Change by Group",
    y = "Percentage",
    x = "",
    fill = "Change Direction"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p_direction)
```

## Baseline Predictors of Change

This section examines whether baseline characteristics (like CCI) predict the magnitude of change, which might explain differences between groups.

```{r baseline-predictors}
# Merge baseline characteristics with change scores
analysis_df <- changes_df |>
  left_join(df_original |> select(CCI, DCSI, age, baseline_VA_logMAR), 
            by = c("baseline_VA_logMAR"))

# Compare baseline predictors by imputation group
predictor_summary <- analysis_df |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    # Correlation between CCI and change
    cor_CCI_change = cor(CCI, change_baseline_to_outcome, use = "complete.obs"),
    # Correlation between baseline VA and change  
    cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
    # Mean change by CCI tercile
    mean_change_low_CCI = mean(change_baseline_to_outcome[CCI <= quantile(CCI, 0.33, na.rm = TRUE)], na.rm = TRUE),
    mean_change_high_CCI = mean(change_baseline_to_outcome[CCI >= quantile(CCI, 0.67, na.rm = TRUE)], na.rm = TRUE),
    .groups = "drop"
  )

kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
      caption = "Relationship between baseline characteristics and vision change")

# Regression model to check if group differences persist after adjustment
adj_model <- lm(change_baseline_to_outcome ~ imputation_group + CCI + DCSI + age + baseline_VA_logMAR,
                data = analysis_df)

adj_summary <- summary(adj_model)
adj_coef <- data.frame(
  Variable = rownames(adj_summary$coefficients),
  Estimate = adj_summary$coefficients[,1],
  SE = adj_summary$coefficients[,2],
  P_Value = adj_summary$coefficients[,4]
) |>
  filter(grepl("imputation_group", Variable))

kable(adj_coef |> mutate(across(where(is.numeric), ~round(., 4))),
      caption = "Group differences after adjusting for baseline characteristics")

# Visualization: Change by CCI tertile and group
analysis_df <- analysis_df |>
  mutate(
    CCI_tertile = cut(CCI, 
                     breaks = quantile(CCI, c(0, 0.33, 0.67, 1), na.rm = TRUE),
                     labels = c("Low", "Medium", "High"),
                     include.lowest = TRUE)
  )

p_cci <- ggplot(analysis_df |> filter(!is.na(CCI_tertile)), 
                aes(x = CCI_tertile, y = change_baseline_to_outcome, fill = imputation_group)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  facet_wrap(~ imputation_group, scales = "free_x") +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", 
                              "Imputed (with intermediate)" = "#A23B72",
                              "Imputed (no intermediate)" = "#F18F01")) +
  labs(
    title = "Vision Change by CCI Tertile and Imputation Group",
    x = "CCI Tertile",
    y = "Change in logMAR"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_cci)
```

## Clinical Plausibility Checks

```{r plausibility}
# Check for extreme changes
extreme_summary <- changes_df |>
  group_by(outcome_was_imputed) |>
  summarise(
    n_extreme_improvement = sum(extreme_improvement, na.rm = TRUE),
    pct_extreme_improvement = round(100 * mean(extreme_improvement, na.rm = TRUE), 2),
    n_extreme_worsening = sum(extreme_worsening, na.rm = TRUE),
    pct_extreme_worsening = round(100 * mean(extreme_worsening, na.rm = TRUE), 2),
    n_any_extreme = sum(extreme_improvement | extreme_worsening, na.rm = TRUE),
    pct_any_extreme = round(100 * mean(extreme_improvement | extreme_worsening, na.rm = TRUE), 2),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(extreme_summary |> select(Group, everything(), -outcome_was_imputed),
      caption = "Frequency of clinically extreme changes (>1.0 logMAR improvement or >1.5 worsening)")

# List specific extreme cases
if (sum(changes_df$extreme_improvement | changes_df$extreme_worsening, na.rm = TRUE) > 0) {
  extreme_cases <- changes_df |>
    filter(extreme_improvement | extreme_worsening) |>
    select(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR, 
           change_baseline_to_outcome, outcome_was_imputed) |>
    mutate(
      Type = ifelse(change_baseline_to_outcome < -1.0, "Extreme Improvement", "Extreme Worsening"),
      Imputed = ifelse(outcome_was_imputed, "Yes", "No")
    ) |>
    arrange(change_baseline_to_outcome)

  if (nrow(extreme_cases) <= 20) {
    kable(extreme_cases |> head(20),
          digits = 3,
          caption = "Examples of extreme changes")
  }
}
```

## Trajectory Analysis

For patients with intermediate values, we check if the imputed outcomes respect the temporal trajectory.

```{r trajectory}
# Focus on those with intermediate values
trajectory_data <- changes_df |>
  filter(had_intermediate) |>
  select(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR,
         change_baseline_to_intermediate, change_intermediate_to_outcome,
         trajectory_reversal, outcome_was_imputed)

# Summary of trajectory consistency
trajectory_summary <- trajectory_data |>
  group_by(outcome_was_imputed) |>
  summarise(
    n = n(),
    n_reversals = sum(trajectory_reversal, na.rm = TRUE),
    pct_reversals = round(100 * mean(trajectory_reversal, na.rm = TRUE), 1),
    mean_slope_1 = mean(change_baseline_to_intermediate, na.rm = TRUE),
    mean_slope_2 = mean(change_intermediate_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(trajectory_summary |> select(Group, everything(), -outcome_was_imputed),
      digits = 3,
      caption = "Trajectory analysis for patients with intermediate values")

# Visualization of trajectories
# Sample some trajectories for visualization
set.seed(2025)
n_sample <- min(100, nrow(trajectory_data))
trajectory_sample <- trajectory_data |>
  sample_n(n_sample) |>
  mutate(
    patient_id = row_number(),
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Reshape for plotting
trajectory_long <- trajectory_sample |>
  select(patient_id, Group, baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR) |>
  pivot_longer(
    cols = c(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR),
    names_to = "timepoint",
    values_to = "VA_logMAR"
  ) |>
  mutate(
    timepoint = factor(timepoint, 
                      levels = c("baseline_VA_logMAR", "intermediate_VA_logMAR", "outcome_VA_logMAR"),
                      labels = c("Baseline", "Intermediate", "2 Years"))
  )

p_trajectory <- ggplot(trajectory_long, aes(x = timepoint, y = VA_logMAR, 
                                           group = patient_id, color = Group)) +
  geom_line(alpha = 0.3) +
  geom_point(alpha = 0.3, size = 0.5) +
  facet_wrap(~ Group) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = paste0("Sample Trajectories (n=", n_sample, " patients with intermediate values)"),
    subtitle = "Lines show individual patient trajectories",
    x = "Timepoint",
    y = "Visual Acuity (logMAR)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_trajectory)
```

## Multiple Imputation Variability

Check consistency across multiple imputations.

```{r mi-variability}
# Calculate change scores for all imputations
all_change_summaries <- list()
for (i in 1:n_imputations) {
  change_data <- all_changes[[i]]

  summary_i <- change_data |>
    group_by(outcome_was_imputed) |>
    summarise(
      mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
      sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(imputation = i)

  all_change_summaries[[i]] <- summary_i
}

mi_summary <- bind_rows(all_change_summaries)

# Calculate between-imputation variability
mi_variability <- mi_summary |>
  group_by(outcome_was_imputed) |>
  summarise(
    mean_of_means = mean(mean_change),
    sd_of_means = sd(mean_change),
    mean_of_sds = mean(sd_change),
    sd_of_sds = sd(sd_change),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(mi_variability |> select(Group, everything(), -outcome_was_imputed),
      digits = 4,
      caption = "Between-imputation variability in change scores")

# Plot variability across imputations
p_mi_var <- ggplot(mi_summary, aes(x = imputation, y = mean_change, 
                                   color = factor(outcome_was_imputed))) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("FALSE" = "#2E86AB", "TRUE" = "#A23B72"),
                    labels = c("Observed", "Imputed")) +
  labs(
    title = "Mean Change Across Imputations",
    x = "Imputation Number",
    y = "Mean Change (logMAR)",
    color = "Group"
  ) +
  theme_minimal()

print(p_mi_var)
```

## Summary and Recommendations

```{r summary}
# Key findings
key_findings <- list()

# 1. Check if distributions are similar
if (ks_test$p.value > 0.05) {
  key_findings$distribution <- "✓ Change distributions are statistically similar between imputed and observed groups"
} else {
  key_findings$distribution <- "⚠ Change distributions differ significantly between groups"
}

# 2. Check mean differences
mean_diff <- abs(change_summary$mean_change[2] - change_summary$mean_change[1])
if (mean_diff < 0.05) {  # Less than 0.5 lines difference
  key_findings$means <- paste0("✓ Mean changes are similar (difference = ", round(mean_diff, 3), " logMAR)")
} else {
  key_findings$means <- paste0("⚠ Mean changes differ by ", round(mean_diff, 3), " logMAR")
}

# 3. Check variability
sd_ratio <- change_summary$sd_change[2] / change_summary$sd_change[1]
if (sd_ratio > 0.8 && sd_ratio < 1.25) {
  key_findings$variability <- paste0("✓ Similar variability (SD ratio = ", round(sd_ratio, 2), ")")
} else {
  key_findings$variability <- paste0("⚠ Different variability (SD ratio = ", round(sd_ratio, 2), ")")
}

# 4. Check extreme values
if (max(extreme_summary$pct_any_extreme) < 5) {
  key_findings$extremes <- "✓ Few extreme values (<5% in both groups)"
} else {
  key_findings$extremes <- paste0("⚠ Notable extreme values (", 
                                  max(extreme_summary$pct_any_extreme), "% in one group)")
}

# Print findings
cat("### Key Findings:\n\n")
for (finding in key_findings) {
  cat("- ", finding, "\n")
}
```

### Potential Concerns

```{r concerns}
concerns <- character()

# Check for specific issues
if (ks_test$p.value < 0.05) {
  concerns <- c(concerns, 
    "The distribution of changes differs between imputed and observed groups, suggesting the imputation may not be capturing the true distribution.")
}

if (mean_diff > 0.1) {
  concerns <- c(concerns,
    "The mean change differs by more than 1 line on the eye chart, which may be clinically meaningful.")
}

if (sd_ratio > 1.5 || sd_ratio < 0.67) {
  concerns <- c(concerns,
    "The variability in imputed values is substantially different from observed values.")
}

if (mean(trajectory_summary$pct_reversals[trajectory_summary$outcome_was_imputed]) > 30) {
  concerns <- c(concerns,
    "Many imputed trajectories show reversals, which may be clinically implausible.")
}

if (length(concerns) > 0) {
  cat("\n### Concerns Identified:\n\n")
  for (i in 1:length(concerns)) {
    cat(i, ". ", concerns[i], "\n\n")
  }

  cat("\n### Recommendations:\n\n")
  cat("Consider alternative imputation approaches that:\n")
  cat("1. Incorporate temporal constraints (e.g., monotone imputation)\n")
  cat("2. Respect clinical boundaries on plausible changes\n")
  cat("3. Use the intermediate values more directly when available\n")
  cat("4. Consider parametric models that can incorporate domain knowledge about vision progression\n")
} else {
  cat("\n### Conclusion:\n")
  cat("The imputation procedure appears to be producing clinically plausible values that are similar to observed outcomes.\n")
}

# Finalize logging
log_message("Imputation diagnostics completed")
final_log <- finalize_log(success = TRUE)
```

**Log file:** `r final_log`