### 2025-08-29 reanalysis-3-imputation_diagnostics.qmd

---
title: "Reanalysis Step 2b: Multiple Imputation Outcome Diagnostics"
subtitle: "Assessing plausibility of imputed visual acuity outcomes"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 6
---

## Overview

This notebook examines whether our multiple imputation procedure for visual acuity outcomes is producing clinically plausible values. We're particularly concerned about:

1. **Temporal consistency**: Do imputed 2-year outcomes respect the trajectory from baseline through intermediate timepoints?
2. **Clinical plausibility**: Are changes in vision realistic given what we know about disease progression?
3. **Distribution similarity**: Do imputed values have similar distributions to observed values?

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
setup_analysis(seed = 2025)

library(mice)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(knitr)

# Initialize logging
log_file <- init_log("imputation_diagnostics")
```

## Load Data

```{r load-data}
# Load the original data to identify which values were imputed
df_original <- readRDS(file.path(reanalysis_data_dir, "df_for_imputation.rds"))

# Load the imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

log_message(paste0("Loaded ", n_imputations, " imputed datasets"))
log_message(paste0("Total observations: ", nrow(df_original)))
```

## Identify Imputed vs Observed Outcomes

```{r identify-imputed}
# In the original data, cohort_id tells us who had observed outcomes
# "submitted" = has true 2-year outcome
# "excluded" = lacks true 2-year outcome (these were imputed)

# Create more nuanced grouping
df_original <- df_original |>
  mutate(
    outcome_was_imputed = cohort_id == "excluded",
    had_intermediate = !is.na(intermediate_VA_logMAR),

    # Create detailed imputation groups
    imputation_group = case_when(
      !outcome_was_imputed ~ "Observed",
      outcome_was_imputed & had_intermediate ~ "Imputed (with intermediate)",
      outcome_was_imputed & !had_intermediate ~ "Imputed (no intermediate)",
      TRUE ~ "Unknown"
    ),
    imputation_group = factor(imputation_group, 
                             levels = c("Observed", 
                                       "Imputed (with intermediate)", 
                                       "Imputed (no intermediate)"))
  )

# Summary of imputation status
imputation_summary <- df_original |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    pct = round(100 * n / nrow(df_original), 1),
    had_baseline = sum(!is.na(baseline_VA_logMAR)),
    pct_baseline = round(100 * had_baseline / n, 1),
    .groups = "drop"
  )

kable(imputation_summary,
      col.names = c("Group", "N", "%", "Has Baseline", "% Baseline"),
      caption = "Detailed breakdown of imputation groups")

# Baseline characteristics by group
baseline_chars <- df_original |>
  group_by(imputation_group) |>
  summarise(
    # Demographics
    mean_age = mean(age, na.rm = TRUE),
    sd_age = sd(age, na.rm = TRUE),
    pct_male = round(100 * mean(gender == "Male", na.rm = TRUE), 1),

    # Clinical characteristics - check if variables exist
    mean_CCI = if("CCI" %in% names(df_original)) mean(CCI, na.rm = TRUE) else NA_real_,
    sd_CCI = if("CCI" %in% names(df_original)) sd(CCI, na.rm = TRUE) else NA_real_,
    mean_DCSI = if("DCSI" %in% names(df_original)) mean(DCSI, na.rm = TRUE) else NA_real_,
    sd_DCSI = if("DCSI" %in% names(df_original)) sd(DCSI, na.rm = TRUE) else NA_real_,

    # Baseline vision
    mean_baseline_VA = mean(baseline_VA_logMAR, na.rm = TRUE),
    sd_baseline_VA = sd(baseline_VA_logMAR, na.rm = TRUE),

    # Treatment indicators
    pct_anti_VEGF = round(100 * mean(anti_VEGF == 1, na.rm = TRUE), 1),
    pct_PRP = round(100 * mean(PRP_flag == 1, na.rm = TRUE), 1),

    # Lapse status - distinguish between observed and missing
    n_lapse_observed = sum(!is.na(person_ever_lapse)),
    n_lapse_missing = sum(is.na(person_ever_lapse)),
    pct_lapse_missing = round(100 * n_lapse_missing / n(), 1),
    # Among those with observed lapse status
    pct_lapsed_observed = round(100 * mean(ever_lapse_binary == 1, na.rm = TRUE), 1),

    .groups = "drop"
  )

kable(baseline_chars |> 
        mutate(across(where(is.numeric), ~round(., 2))),
      caption = "Baseline characteristics by imputation group")
```

## Lapse Status Missingness Analysis

```{r lapse-missingness}
# Detailed analysis of lapse status by group
lapse_analysis <- df_original |>
  group_by(imputation_group) |>
  summarise(
    n_total = n(),

    # Lapse status availability
    n_lapse_known = sum(!is.na(person_ever_lapse)),
    n_lapse_missing = sum(is.na(person_ever_lapse)),
    pct_lapse_missing = round(100 * n_lapse_missing / n_total, 1),

    # Among those with known lapse status
    n_lapsed_if_known = sum(person_ever_lapse == TRUE, na.rm = TRUE),
    n_not_lapsed_if_known = sum(person_ever_lapse == FALSE, na.rm = TRUE),
    pct_lapsed_if_known = round(100 * n_lapsed_if_known / n_lapse_known, 1),

    # Overall lapse rate (including imputed - ever_lapse_binary should be complete)
    pct_lapsed_overall = round(100 * mean(ever_lapse_binary == 1), 1),

    .groups = "drop"
  )

kable(lapse_analysis,
      col.names = c("Group", "N Total", "N Lapse Known", "N Lapse Missing", 
                    "% Lapse Missing", "N Lapsed (if known)", "N Not Lapsed (if known)",
                    "% Lapsed (if known)", "% Lapsed (overall)"),
      caption = "Lapse status availability and rates by imputation group")

# Cross-tabulation of outcome availability vs lapse status
outcome_by_lapse <- df_original |>
  mutate(
    has_outcome = !is.na(outcome_VA_logMAR),
    lapse_known = !is.na(person_ever_lapse)
  ) |>
  group_by(lapse_known, has_outcome) |>
  summarise(
    n = n(),
    pct_of_total = round(100 * n / nrow(df_original), 1),
    # Among those with known lapse status
    pct_lapsed = round(100 * mean(ever_lapse_binary == 1, na.rm = TRUE), 1),
    .groups = "drop"
  )

kable(outcome_by_lapse,
      caption = "Relationship between outcome availability and lapse status knowledge")

# Key insight: Check if missing outcomes correlate with lapse
missing_outcome_lapse <- df_original |>
  filter(!is.na(person_ever_lapse)) |>  # Only those with known lapse status
  group_by(ever_lapse_binary) |>
  summarise(
    n = n(),
    n_missing_outcome = sum(is.na(outcome_VA_logMAR)),
    pct_missing_outcome = round(100 * n_missing_outcome / n, 1),
    .groups = "drop"
  ) |>
  mutate(
    lapse_status = ifelse(ever_lapse_binary == 1, "Lapsed", "Not Lapsed")
  ) |>
  select(lapse_status, n, n_missing_outcome, pct_missing_outcome)

kable(missing_outcome_lapse,
      col.names = c("Lapse Status", "N", "N Missing Outcome", "% Missing Outcome"),
      caption = "Outcome missingness by known lapse status")
```

## Calculate Change Scores

```{r calculate-changes}
# Add row identifier to original data for proper merging
df_original <- df_original |>
  mutate(row_id = row_number())

# Function to calculate change scores for a single imputed dataset
calculate_change_scores <- function(imp_data, original_data) {
  imp_data |>
    mutate(
      row_id = row_number(),
      # Add imputation status from original data
      outcome_was_imputed = original_data$outcome_was_imputed,
      had_intermediate = original_data$had_intermediate,
      imputation_group = original_data$imputation_group,

      # Calculate changes
      change_baseline_to_outcome = outcome_VA_logMAR - baseline_VA_logMAR,
      change_baseline_to_intermediate = intermediate_VA_logMAR - baseline_VA_logMAR,
      change_intermediate_to_outcome = outcome_VA_logMAR - intermediate_VA_logMAR,

      # Categorize changes (negative = improvement, positive = worsening)
      # Using 0.1 logMAR as clinically meaningful change (approximately 1 line on eye chart)
      outcome_direction = case_when(
        change_baseline_to_outcome < -0.1 ~ "Improved",
        change_baseline_to_outcome > 0.1 ~ "Worsened",
        TRUE ~ "Stable"
      ),

      # Flag extreme changes (clinically implausible)
      extreme_improvement = change_baseline_to_outcome < -1.0,  # >10 lines improvement
      extreme_worsening = change_baseline_to_outcome > 1.5,      # >15 lines worsening

      # For those with intermediate values, check trajectory consistency
      trajectory_reversal = case_when(
        !is.na(change_baseline_to_intermediate) & !is.na(change_intermediate_to_outcome) ~
          sign(change_baseline_to_intermediate) != sign(change_intermediate_to_outcome),
        TRUE ~ FALSE
      )
    )
}

# Calculate for all imputed datasets
all_changes <- list()
for (i in 1:n_imputations) {
  all_changes[[i]] <- calculate_change_scores(imputed_datasets[[i]], df_original)
}

# For summary statistics, use the first imputation as representative
changes_df <- all_changes[[1]]
```

## Compare Change Distributions

### Summary Statistics

```{r summary-stats}
# Compare change scores across all three groups
change_summary_detailed <- changes_df |>
  group_by(imputation_group) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    q25_change = quantile(change_baseline_to_outcome, 0.25, na.rm = TRUE),
    q75_change = quantile(change_baseline_to_outcome, 0.75, na.rm = TRUE),
    min_change = min(change_baseline_to_outcome, na.rm = TRUE),
    max_change = max(change_baseline_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    iqr_change = q75_change - q25_change
  )

kable(change_summary_detailed |> 
        select(imputation_group, n, mean_change, sd_change, median_change, iqr_change, min_change, max_change),
      digits = 3,
      col.names = c("Group", "N", "Mean", "SD", "Median", "IQR", "Min", "Max"),
      caption = "Distribution of vision changes (outcome - baseline) by detailed imputation group")

# Also keep simplified version for compatibility
change_summary <- changes_df |>
  group_by(outcome_was_imputed) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    q25_change = quantile(change_baseline_to_outcome, 0.25, na.rm = TRUE),
    q75_change = quantile(change_baseline_to_outcome, 0.75, na.rm = TRUE),
    min_change = min(change_baseline_to_outcome, na.rm = TRUE),
    max_change = max(change_baseline_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(
    iqr_change = q75_change - q25_change,
    group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Statistical tests across three groups
anova_test <- aov(change_baseline_to_outcome ~ imputation_group, data = changes_df)
anova_summary <- summary(anova_test)

# Pairwise comparisons
pairwise_tests <- pairwise.t.test(changes_df$change_baseline_to_outcome, 
                                  changes_df$imputation_group,
                                  p.adjust.method = "bonferroni")

# Format ANOVA results
anova_results <- data.frame(
  Test = "One-way ANOVA",
  F_Statistic = anova_summary[[1]]$`F value`[1],
  df1 = anova_summary[[1]]$Df[1],
  df2 = anova_summary[[1]]$Df[2],
  P_Value = anova_summary[[1]]$`Pr(>F)`[1],
  Interpretation = ifelse(anova_summary[[1]]$`Pr(>F)`[1] < 0.05, 
                         "Groups differ significantly", "No significant difference")
)

kable(anova_results, digits = 4, caption = "ANOVA comparing change distributions across three groups")

# Format pairwise results
pairwise_matrix <- pairwise_tests$p.value
pairwise_df <- data.frame(
  Comparison = c("Imputed (with intermediate) vs Observed",
                "Imputed (no intermediate) vs Observed",
                "Imputed (no intermediate) vs Imputed (with intermediate)"),
  P_Value = c(pairwise_matrix[1,1], pairwise_matrix[2,1], pairwise_matrix[2,2]),
  Significant = ifelse(c(pairwise_matrix[1,1], pairwise_matrix[2,1], pairwise_matrix[2,2]) < 0.05,
                      "Yes", "No")
)

kable(pairwise_df, digits = 4, caption = "Pairwise comparisons (Bonferroni adjusted)")

# Also do simplified tests for backward compatibility
t_test <- t.test(change_baseline_to_outcome ~ outcome_was_imputed, data = changes_df)
ks_test <- ks.test(
  changes_df$change_baseline_to_outcome[changes_df$outcome_was_imputed],
  changes_df$change_baseline_to_outcome[!changes_df$outcome_was_imputed]
)
```

### Visualization of Change Distributions

```{r change-distributions}
#| fig-height: 8
#| fig-width: 12

# Prepare data for plotting - using detailed groups
plot_data_detailed <- changes_df |>
  filter(!is.na(change_baseline_to_outcome))

# Also prepare simplified version for some plots
plot_data <- changes_df |>
  mutate(
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed"),
    Group = factor(Group, levels = c("Observed", "Imputed"))
  ) |>
  filter(!is.na(change_baseline_to_outcome))

# 1. Density plots - detailed groups
p1_detailed <- ggplot(plot_data_detailed, aes(x = change_baseline_to_outcome, fill = imputation_group)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(data = change_summary_detailed, 
             aes(xintercept = mean_change, color = imputation_group),
             linetype = "solid", size = 1) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", 
                              "Imputed (with intermediate)" = "#A23B72",
                              "Imputed (no intermediate)" = "#F18F01")) +
  scale_color_manual(values = c("Observed" = "#2E86AB", 
                               "Imputed (with intermediate)" = "#A23B72",
                               "Imputed (no intermediate)" = "#F18F01")) +
  labs(
    title = "Distribution of Vision Changes by Detailed Group",
    subtitle = "Vertical lines show group means",
    x = "Change in logMAR (positive = worsening)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 1b. Simplified density plot (for comparison)
p1 <- ggplot(plot_data, aes(x = change_baseline_to_outcome, fill = Group)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  geom_vline(data = change_summary, 
             aes(xintercept = mean_change, color = group),
             linetype = "solid", size = 1) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = "Distribution of Vision Changes (Simplified)",
    subtitle = "Vertical lines show group means",
    x = "Change in logMAR (positive = worsening)",
    y = "Density"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# 2. Boxplots - detailed groups
p2_detailed <- ggplot(plot_data_detailed, aes(x = imputation_group, y = change_baseline_to_outcome, fill = imputation_group)) +
  geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_fill_manual(values = c("Observed" = "#2E86AB", 
                              "Imputed (with intermediate)" = "#A23B72",
                              "Imputed (no intermediate)" = "#F18F01")) +
  labs(
    title = "Vision Changes by Detailed Group",
    subtitle = "Box shows IQR, whiskers extend to 1.5*IQR",
    y = "Change in logMAR (positive = worsening)",
    x = ""
  ) +
  theme_minimal() +
  theme(legend.position = "none",
        axis.text.x = element_text(angle = 45, hjust = 1))

# 3. Q-Q plot
observed_changes <- plot_data$change_baseline_to_outcome[plot_data$Group == "Observed"]
imputed_changes <- plot_data$change_baseline_to_outcome[plot_data$Group == "Imputed"]

# Sample to same size for Q-Q plot
n_min <- min(length(observed_changes), length(imputed_changes))
qq_data <- data.frame(
  observed = sort(sample(observed_changes, n_min)),
  imputed = sort(sample(imputed_changes, n_min))
)

p3 <- ggplot(qq_data, aes(x = observed, y = imputed)) +
  geom_point(alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, color = "red", linetype = "dashed") +
  labs(
    title = "Q-Q Plot: Imputed vs Observed Changes",
    subtitle = "Points should follow red line if distributions are similar",
    x = "Observed Change Quantiles",
    y = "Imputed Change Quantiles"
  ) +
  theme_minimal()

# 4. Cumulative distribution
p4 <- ggplot(plot_data, aes(x = change_baseline_to_outcome, color = Group)) +
  stat_ecdf(size = 1) +
  geom_vline(xintercept = 0, linetype = "dashed", alpha = 0.5) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = "Cumulative Distribution of Changes",
    x = "Change in logMAR",
    y = "Cumulative Probability"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

# Arrange plots - show detailed version first
grid.arrange(p1_detailed, p2_detailed, ncol = 2)

# Then show additional comparisons
grid.arrange(p3, p4, ncol = 2)
```

## Clinical Trajectory by Lapse Status

```{r trajectory-by-lapse}
# Analyze vision changes by lapse status (across all imputation groups)
trajectory_by_lapse <- changes_df |>
  filter(!is.na(change_baseline_to_outcome)) |>
  group_by(ever_lapse_binary, imputation_group) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    pct_improved = round(100 * mean(outcome_direction == "Improved", na.rm = TRUE), 1),
    pct_stable = round(100 * mean(outcome_direction == "Stable", na.rm = TRUE), 1),
    pct_worsened = round(100 * mean(outcome_direction == "Worsened", na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  mutate(
    lapse_status = ifelse(ever_lapse_binary == 1, "Lapsed", "Not Lapsed")
  ) |>
  select(-ever_lapse_binary) |>
  arrange(imputation_group, lapse_status)

kable(trajectory_by_lapse |>
        mutate(across(where(is.numeric) & !n, ~round(., 3))),
      caption = "Vision trajectories by lapse status and imputation group")

# Overall comparison (pooled across groups)
overall_by_lapse <- changes_df |>
  filter(!is.na(change_baseline_to_outcome)) |>
  group_by(ever_lapse_binary) |>
  summarise(
    n = n(),
    mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
    median_change = median(change_baseline_to_outcome, na.rm = TRUE),
    pct_improved = round(100 * mean(outcome_direction == "Improved", na.rm = TRUE), 1),
    pct_worsened = round(100 * mean(outcome_direction == "Worsened", na.rm = TRUE), 1),
    .groups = "drop"
  ) |>
  mutate(
    lapse_status = ifelse(ever_lapse_binary == 1, "Lapsed", "Not Lapsed")
  ) |>
  select(lapse_status, everything(), -ever_lapse_binary)

kable(overall_by_lapse |>
        mutate(across(where(is.numeric) & !n, ~round(., 3))),
      caption = "Overall vision trajectories by lapse status (pooled across all groups)")

# Statistical test
lapse_test <- t.test(change_baseline_to_outcome ~ ever_lapse_binary, data = changes_df)

cat("T-test comparing vision changes between lapsed and not lapsed patients:\n")
cat(sprintf("Mean difference: %.3f logMAR units\n", diff(lapse_test$estimate)))
cat(sprintf("95%% CI: [%.3f, %.3f]\n", lapse_test$conf.int[1], lapse_test$conf.int[2]))
cat(sprintf("p-value: %.4f\n", lapse_test$p.value))

# Visualization
p_lapse <- ggplot(changes_df |> filter(!is.na(change_baseline_to_outcome)),
                  aes(x = factor(ever_lapse_binary, labels = c("Not Lapsed", "Lapsed")),
                      y = change_baseline_to_outcome,
                      fill = factor(ever_lapse_binary))) +
  geom_boxplot(alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
  facet_wrap(~imputation_group) +
  scale_fill_manual(values = c("0" = "#2E86AB", "1" = "#A23B72")) +
  labs(
    title = "Vision Changes by Lapse Status Across Imputation Groups",
    subtitle = "Positive values indicate vision worsening",
    x = "Lapse Status",
    y = "Change in logMAR"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_lapse)

# Check if lapsed patients have paradoxically better outcomes
paradox_check <- trajectory_by_lapse |>
  group_by(imputation_group) |>
  summarise(
    lapsed_mean = mean_change[lapse_status == "Lapsed"],
    not_lapsed_mean = mean_change[lapse_status == "Not Lapsed"],
    difference = lapsed_mean - not_lapsed_mean,
    paradoxical = difference < 0,  # Negative means lapsed have better outcomes
    interpretation = ifelse(
      difference < -0.05, "Lapsed have notably better vision",
      ifelse(difference > 0.05, "Lapsed have worse vision (expected)",
             "Similar outcomes")
    ),
    .groups = "drop"
  )

kable(paradox_check,
      caption = "Clinical plausibility check: Do lapsed patients have worse vision outcomes?")
```

## Directional Change Analysis

```{r directional-changes}
# Compare proportions across detailed groups
direction_summary_detailed <- changes_df |>
  filter(!is.na(outcome_direction)) |>
  group_by(imputation_group, outcome_direction) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(imputation_group) |>
  mutate(
    total = sum(n),
    pct = round(100 * n / total, 1)
  )

# Reshape for display
direction_wide_detailed <- direction_summary_detailed |>
  select(imputation_group, outcome_direction, n, pct) |>
  pivot_wider(
    names_from = outcome_direction,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = 0)
  )

kable(direction_wide_detailed,
      caption = "Direction of vision change by detailed imputation group")

# Also create simplified version for backward compatibility
direction_summary <- changes_df |>
  filter(!is.na(outcome_direction)) |>
  group_by(outcome_was_imputed, outcome_direction) |>
  summarise(n = n(), .groups = "drop") |>
  group_by(outcome_was_imputed) |>
  mutate(
    total = sum(n),
    pct = round(100 * n / total, 1),
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Reshape for display
direction_wide <- direction_summary |>
  select(Group, outcome_direction, n, pct) |>
  pivot_wider(
    names_from = outcome_direction,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = 0)
  )

kable(direction_wide,
      caption = "Direction of vision change by imputation status")

# Chi-square test for independence
direction_table <- table(changes_df$outcome_was_imputed, changes_df$outcome_direction)
chi_test <- chisq.test(direction_table)

chi_result <- data.frame(
  Test = "Chi-square test of independence",
  Chi_Square = chi_test$statistic,
  df = chi_test$parameter,
  P_Value = chi_test$p.value,
  Interpretation = ifelse(chi_test$p.value < 0.05, 
                         "Proportions differ between groups",
                         "Proportions similar between groups")
)

kable(chi_result, digits = 4)

# Visualization
direction_plot_data <- direction_summary |>
  mutate(outcome_direction = factor(outcome_direction, 
                                   levels = c("Improved", "Stable", "Worsened")))

p_direction <- ggplot(direction_plot_data, 
                     aes(x = Group, y = pct, fill = outcome_direction)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_text(aes(label = paste0(pct, "%")), 
           position = position_dodge(width = 0.9),
           vjust = -0.5) +
  scale_fill_manual(values = c("Improved" = "#2ca02c", 
                              "Stable" = "#ff7f0e", 
                              "Worsened" = "#d62728")) +
  labs(
    title = "Direction of Vision Change by Group",
    y = "Percentage",
    x = "",
    fill = "Change Direction"
  ) +
  theme_minimal() +
  theme(legend.position = "bottom")

print(p_direction)
```

## Baseline Predictors of Change

This section examines whether baseline characteristics (like CCI) predict the magnitude of change, which might explain differences between groups.

```{r baseline-predictors}
# Check what variables are available in the original data
available_vars <- names(df_original)
baseline_vars <- intersect(c("row_id", "CCI", "DCSI", "age"), available_vars)

# Merge baseline characteristics with change scores using row_id
if (length(baseline_vars) > 1) {
  analysis_df <- changes_df |>
    left_join(df_original |> select(all_of(baseline_vars)), 
              by = "row_id")
} else {
  # If baseline variables aren't available, create a simpler analysis
  analysis_df <- changes_df
  cat("Note: CCI, DCSI, or age not found in original data. Using simplified analysis.\n")
}

# Only do correlation analysis if CCI exists
if ("CCI" %in% names(analysis_df)) {
  # Compare baseline predictors by imputation group
  predictor_summary <- analysis_df |>
    group_by(imputation_group) |>
    summarise(
      n = n(),
      # Correlation between CCI and change
      cor_CCI_change = cor(CCI, change_baseline_to_outcome, use = "complete.obs"),
      # Correlation between baseline VA and change  
      cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
      # Mean change by CCI tercile
      mean_change_low_CCI = mean(change_baseline_to_outcome[CCI <= quantile(CCI, 0.33, na.rm = TRUE)], na.rm = TRUE),
      mean_change_high_CCI = mean(change_baseline_to_outcome[CCI >= quantile(CCI, 0.67, na.rm = TRUE)], na.rm = TRUE),
      .groups = "drop"
    )

  kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
        caption = "Relationship between baseline characteristics and vision change")
} else {
  # Simplified summary without CCI
  predictor_summary <- analysis_df |>
    group_by(imputation_group) |>
    summarise(
      n = n(),
      mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
      sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
      cor_baseline_change = cor(baseline_VA_logMAR, change_baseline_to_outcome, use = "complete.obs"),
      .groups = "drop"
    )

  kable(predictor_summary |> mutate(across(where(is.numeric), ~round(., 3))),
        caption = "Vision change by imputation group (CCI not available)")
}

# Regression model to check if group differences persist after adjustment
if (all(c("CCI", "DCSI", "age") %in% names(analysis_df))) {
  adj_model <- lm(change_baseline_to_outcome ~ imputation_group + CCI + DCSI + age + baseline_VA_logMAR,
                  data = analysis_df)

  adj_summary <- summary(adj_model)
  adj_coef <- data.frame(
    Variable = rownames(adj_summary$coefficients),
    Estimate = adj_summary$coefficients[,1],
    SE = adj_summary$coefficients[,2],
    P_Value = adj_summary$coefficients[,4]
  ) |>
    filter(grepl("imputation_group", Variable))

  kable(adj_coef |> mutate(across(where(is.numeric), ~round(., 4))),
        caption = "Group differences after adjusting for baseline characteristics")

  # Visualization: Change by CCI tertile and group
  analysis_df <- analysis_df |>
    mutate(
      CCI_tertile = cut(CCI, 
                       breaks = quantile(CCI, c(0, 0.33, 0.67, 1), na.rm = TRUE),
                       labels = c("Low", "Medium", "High"),
                       include.lowest = TRUE)
    )

  p_cci <- ggplot(analysis_df |> filter(!is.na(CCI_tertile)), 
                  aes(x = CCI_tertile, y = change_baseline_to_outcome, fill = imputation_group)) +
    geom_boxplot(alpha = 0.7, outlier.alpha = 0.3) +
    facet_wrap(~ imputation_group, scales = "free_x") +
    geom_hline(yintercept = 0, linetype = "dashed", alpha = 0.5) +
    scale_fill_manual(values = c("Observed" = "#2E86AB", 
                                "Imputed (with intermediate)" = "#A23B72",
                                "Imputed (no intermediate)" = "#F18F01")) +
    labs(
      title = "Vision Change by CCI Tertile and Imputation Group",
      x = "CCI Tertile",
      y = "Change in logMAR"
    ) +
    theme_minimal() +
    theme(legend.position = "none")

  print(p_cci)
} else {
  # Simpler model without baseline covariates
  adj_model <- lm(change_baseline_to_outcome ~ imputation_group + baseline_VA_logMAR,
                  data = analysis_df)

  adj_summary <- summary(adj_model)
  adj_coef <- data.frame(
    Variable = rownames(adj_summary$coefficients),
    Estimate = adj_summary$coefficients[,1],
    SE = adj_summary$coefficients[,2],
    P_Value = adj_summary$coefficients[,4]
  ) |>
    filter(grepl("imputation_group", Variable))

  kable(adj_coef |> mutate(across(where(is.numeric), ~round(., 4))),
        caption = "Group differences after adjusting for baseline VA only")
}
```

## Clinical Plausibility Checks

```{r plausibility}
# Check for extreme changes
extreme_summary <- changes_df |>
  group_by(outcome_was_imputed) |>
  summarise(
    n_extreme_improvement = sum(extreme_improvement, na.rm = TRUE),
    pct_extreme_improvement = round(100 * mean(extreme_improvement, na.rm = TRUE), 2),
    n_extreme_worsening = sum(extreme_worsening, na.rm = TRUE),
    pct_extreme_worsening = round(100 * mean(extreme_worsening, na.rm = TRUE), 2),
    n_any_extreme = sum(extreme_improvement | extreme_worsening, na.rm = TRUE),
    pct_any_extreme = round(100 * mean(extreme_improvement | extreme_worsening, na.rm = TRUE), 2),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(extreme_summary |> select(Group, everything(), -outcome_was_imputed),
      caption = "Frequency of clinically extreme changes (>1.0 logMAR improvement or >1.5 worsening)")

# List specific extreme cases
if (sum(changes_df$extreme_improvement | changes_df$extreme_worsening, na.rm = TRUE) > 0) {
  extreme_cases <- changes_df |>
    filter(extreme_improvement | extreme_worsening) |>
    select(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR, 
           change_baseline_to_outcome, outcome_was_imputed) |>
    mutate(
      Type = ifelse(change_baseline_to_outcome < -1.0, "Extreme Improvement", "Extreme Worsening"),
      Imputed = ifelse(outcome_was_imputed, "Yes", "No")
    ) |>
    arrange(change_baseline_to_outcome)

  if (nrow(extreme_cases) <= 20) {
    kable(extreme_cases |> head(20),
          digits = 3,
          caption = "Examples of extreme changes")
  }
}
```

## Trajectory Analysis

For patients with intermediate values, we check if the imputed outcomes respect the temporal trajectory.

```{r trajectory}
# Focus on those with intermediate values
trajectory_data <- changes_df |>
  filter(had_intermediate) |>
  select(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR,
         change_baseline_to_intermediate, change_intermediate_to_outcome,
         trajectory_reversal, outcome_was_imputed)

# Summary of trajectory consistency
trajectory_summary <- trajectory_data |>
  group_by(outcome_was_imputed) |>
  summarise(
    n = n(),
    n_reversals = sum(trajectory_reversal, na.rm = TRUE),
    pct_reversals = round(100 * mean(trajectory_reversal, na.rm = TRUE), 1),
    mean_slope_1 = mean(change_baseline_to_intermediate, na.rm = TRUE),
    mean_slope_2 = mean(change_intermediate_to_outcome, na.rm = TRUE),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(trajectory_summary |> select(Group, everything(), -outcome_was_imputed),
      digits = 3,
      caption = "Trajectory analysis for patients with intermediate values")

# Visualization of trajectories
# Sample some trajectories for visualization
set.seed(2025)
n_sample <- min(100, nrow(trajectory_data))
trajectory_sample <- trajectory_data |>
  sample_n(n_sample) |>
  mutate(
    patient_id = row_number(),
    Group = ifelse(outcome_was_imputed, "Imputed", "Observed")
  )

# Reshape for plotting
trajectory_long <- trajectory_sample |>
  select(patient_id, Group, baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR) |>
  pivot_longer(
    cols = c(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR),
    names_to = "timepoint",
    values_to = "VA_logMAR"
  ) |>
  mutate(
    timepoint = factor(timepoint, 
                      levels = c("baseline_VA_logMAR", "intermediate_VA_logMAR", "outcome_VA_logMAR"),
                      labels = c("Baseline", "Intermediate", "2 Years"))
  )

p_trajectory <- ggplot(trajectory_long, aes(x = timepoint, y = VA_logMAR, 
                                           group = patient_id, color = Group)) +
  geom_line(alpha = 0.3) +
  geom_point(alpha = 0.3, size = 0.5) +
  facet_wrap(~ Group) +
  scale_color_manual(values = c("Observed" = "#2E86AB", "Imputed" = "#A23B72")) +
  labs(
    title = paste0("Sample Trajectories (n=", n_sample, " patients with intermediate values)"),
    subtitle = "Lines show individual patient trajectories",
    x = "Timepoint",
    y = "Visual Acuity (logMAR)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_trajectory)
```

## Multiple Imputation Variability

Check consistency across multiple imputations.

```{r mi-variability}
# Calculate change scores for all imputations
all_change_summaries <- list()
for (i in 1:n_imputations) {
  change_data <- all_changes[[i]]

  summary_i <- change_data |>
    group_by(outcome_was_imputed) |>
    summarise(
      mean_change = mean(change_baseline_to_outcome, na.rm = TRUE),
      sd_change = sd(change_baseline_to_outcome, na.rm = TRUE),
      .groups = "drop"
    ) |>
    mutate(imputation = i)

  all_change_summaries[[i]] <- summary_i
}

mi_summary <- bind_rows(all_change_summaries)

# Calculate between-imputation variability
mi_variability <- mi_summary |>
  group_by(outcome_was_imputed) |>
  summarise(
    mean_of_means = mean(mean_change),
    sd_of_means = sd(mean_change),
    mean_of_sds = mean(sd_change),
    sd_of_sds = sd(sd_change),
    .groups = "drop"
  ) |>
  mutate(Group = ifelse(outcome_was_imputed, "Imputed", "Observed"))

kable(mi_variability |> select(Group, everything(), -outcome_was_imputed),
      digits = 4,
      caption = "Between-imputation variability in change scores")

# Plot variability across imputations
p_mi_var <- ggplot(mi_summary, aes(x = imputation, y = mean_change, 
                                   color = factor(outcome_was_imputed))) +
  geom_point() +
  geom_line() +
  scale_color_manual(values = c("FALSE" = "#2E86AB", "TRUE" = "#A23B72"),
                    labels = c("Observed", "Imputed")) +
  labs(
    title = "Mean Change Across Imputations",
    x = "Imputation Number",
    y = "Mean Change (logMAR)",
    color = "Group"
  ) +
  theme_minimal()

print(p_mi_var)
```

## Summary and Recommendations

```{r summary}
# Key findings
key_findings <- list()

# 1. Check if distributions are similar
if (ks_test$p.value > 0.05) {
  key_findings$distribution <- "✓ Change distributions are statistically similar between imputed and observed groups"
} else {
  key_findings$distribution <- "⚠ Change distributions differ significantly between groups"
}

# 2. Check mean differences
mean_diff <- abs(change_summary$mean_change[2] - change_summary$mean_change[1])
if (mean_diff < 0.05) {  # Less than 0.5 lines difference
  key_findings$means <- paste0("✓ Mean changes are similar (difference = ", round(mean_diff, 3), " logMAR)")
} else {
  key_findings$means <- paste0("⚠ Mean changes differ by ", round(mean_diff, 3), " logMAR")
}

# 3. Check variability
sd_ratio <- change_summary$sd_change[2] / change_summary$sd_change[1]
if (sd_ratio > 0.8 && sd_ratio < 1.25) {
  key_findings$variability <- paste0("✓ Similar variability (SD ratio = ", round(sd_ratio, 2), ")")
} else {
  key_findings$variability <- paste0("⚠ Different variability (SD ratio = ", round(sd_ratio, 2), ")")
}

# 4. Check extreme values
if (max(extreme_summary$pct_any_extreme) < 5) {
  key_findings$extremes <- "✓ Few extreme values (<5% in both groups)"
} else {
  key_findings$extremes <- paste0("⚠ Notable extreme values (", 
                                  max(extreme_summary$pct_any_extreme), "% in one group)")
}

# Print findings
cat("### Key Findings:\n\n")
for (finding in key_findings) {
  cat("- ", finding, "\n")
}
```

### Key Insights on Lapse Status and Outcomes

```{r lapse-insights}
# Summarize key findings about lapse status
cat("#### Lapse Status Missingness:\n\n")

# Check if lapse status is mostly observed
lapse_missing_overall <- sum(is.na(df_original$person_ever_lapse)) / nrow(df_original) * 100
if (lapse_missing_overall < 5) {
  cat(sprintf("- Lapse status is well-documented (only %.1f%% missing)\n", lapse_missing_overall))
} else {
  cat(sprintf("- ⚠ Lapse status has %.1f%% missing values\n", lapse_missing_overall))
}

# Report on high lapse rates in imputed groups
cat("\n#### Lapse Rates by Group:\n\n")
for (i in 1:nrow(lapse_analysis)) {
  group <- lapse_analysis$imputation_group[i]
  lapse_rate <- lapse_analysis$pct_lapsed_overall[i]
  cat(sprintf("- %s: %.1f%% lapsed\n", group, lapse_rate))
}

cat("\n#### Clinical Trajectory Plausibility:\n\n")
# Check the paradox
for (i in 1:nrow(paradox_check)) {
  group <- paradox_check$imputation_group[i]
  interp <- paradox_check$interpretation[i]
  diff <- paradox_check$difference[i]
  cat(sprintf("- %s: %s (difference = %.3f logMAR)\n", group, interp, diff))
}

cat("\n#### Interpretation:\n\n")
# High lapse rate interpretation
if (any(lapse_analysis$pct_lapsed_overall[lapse_analysis$imputation_group != "Observed"] > 95)) {
  cat("The extremely high lapse rates (>95%) in some imputed groups may actually be accurate:\n")
  cat("- These patients lack 2-year outcomes precisely BECAUSE they lapsed from care\n")
  cat("- The missingness mechanism is directly related to the lapse status\n")
  cat("- This is a form of 'Missing Not At Random' (MNAR) where missingness itself is informative\n\n")
}

# Clinical trajectory interpretation
if (any(paradox_check$paradoxical)) {
  cat("⚠ CONCERN: Lapsed patients show better vision outcomes in some groups:\n")
  cat("- This violates clinical expectations (lapsed patients typically have worse outcomes)\n")
  cat("- Possible explanations:\n")
  cat("  1. Selection bias: Healthier patients may be more likely to skip appointments\n")
  cat("  2. Imputation artifact: The model may not properly account for the lapse-outcome relationship\n")
  cat("  3. Data issue: Lapse definition or timing may not align with clinical deterioration\n\n")
}
```

### Potential Concerns

```{r concerns}
concerns <- character()

# Check for specific issues
if (ks_test$p.value < 0.05) {
  concerns <- c(concerns, 
    "The distribution of changes differs between imputed and observed groups, suggesting the imputation may not be capturing the true distribution.")
}

if (mean_diff > 0.1) {
  concerns <- c(concerns,
    "The mean change differs by more than 1 line on the eye chart, which may be clinically meaningful.")
}

if (sd_ratio > 1.5 || sd_ratio < 0.67) {
  concerns <- c(concerns,
    "The variability in imputed values is substantially different from observed values.")
}

if (mean(trajectory_summary$pct_reversals[trajectory_summary$outcome_was_imputed]) > 30) {
  concerns <- c(concerns,
    "Many imputed trajectories show reversals, which may be clinically implausible.")
}

# Add new concern about lapse-outcome paradox
if (any(paradox_check$paradoxical)) {
  concerns <- c(concerns,
    "Lapsed patients show better vision outcomes than non-lapsed patients in some groups, which is clinically implausible.")
}

# Add concern about extreme lapse rates
if (any(lapse_analysis$pct_lapsed_overall[lapse_analysis$imputation_group != "Observed"] > 95)) {
  concerns <- c(concerns,
    "Extremely high lapse rates (>95%) in imputed groups suggest the imputation may be conflating missingness with lapse status.")
}

if (length(concerns) > 0) {
  cat("\n### Concerns Identified:\n\n")
  for (i in 1:length(concerns)) {
    cat(i, ". ", concerns[i], "\n\n")
  }

  cat("\n### Recommendations:\n\n")
  cat("Consider the following approaches:\n\n")
  cat("1. **Sensitivity Analysis**: Compare results using:\n")
  cat("   - Complete case analysis only\n")
  cat("   - Imputation restricted to those with intermediate values\n")
  cat("   - Different assumptions about lapsed patient outcomes\n\n")
  cat("2. **Alternative Imputation Methods**:\n")
  cat("   - Use pattern mixture models that explicitly model different trajectories for lapsed vs non-lapsed\n")
  cat("   - Implement bounded imputation that restricts vision changes to clinically plausible ranges\n")
  cat("   - Consider worst-case imputation for lapsed patients (assume deterioration)\n\n")
  cat("3. **Reframe the Analysis**:\n")
  cat("   - Treat missing outcomes as a competing risk rather than missing data\n")
  cat("   - Focus on the population that remained in care (per-protocol analysis)\n")
  cat("   - Use inverse probability weighting to account for differential dropout\n\n")
} else {
  cat("\n### Conclusion:\n")
  cat("The imputation procedure appears to be producing clinically plausible values that are similar to observed outcomes.\n")
}

# Finalize logging
log_message("Imputation diagnostics completed")
final_log <- finalize_log(success = TRUE)
```

**Log file:** `r final_log`