### 2025-08-29 reanalysis-2-multiple_imputation_clean.qmd
refactor

---
title: "Reanalysis Step 2: Multiple Imputation (Clean Version)"
subtitle: "Simplified imputation using mice with clustering"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
seed <- 2025
setup_analysis(seed = seed)

library(mice)
library(miceadds)
library(VIM)
library(naniar)
library(dplyr)
library(tidyr)
library(knitr)
library(lme4)

# Initialize logging
log_file <- init_log("multiple_imputation_clean")

# ANALYSIS MODE
analysis_mode <- "test"  # "test" or "final"

if (analysis_mode == "test") {
  n_imputations <- 2
  n_iterations <- 3
} else {
  n_imputations <- 30
  n_iterations <- 20
}

log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Imputations: ", n_imputations, ", Iterations: ", n_iterations))
```

## Load Data

```{r load-data}
# Load the prepared dataset
df_original <- readRDS(file.path(reanalysis_data_dir, "df_for_imputation.rds"))
log_message(paste0("Loaded ", nrow(df_original), " observations"))

# Get unique patient count
n_patients <- n_distinct(df_original$e_mrn_deidentified)
log_message(paste0("Number of unique patients: ", n_patients))
```

## Define Variables for Imputation

```{r define-variables}
# Variables that should NOT be imputed and should NOT be used as predictors
vars_no_impute_no_predict <- c(
  "row_id"  # Internal row identifier
)

# Variables that should NOT be imputed but CAN be used as predictors  
vars_no_impute_yes_predict <- c(
  "ever_lapse_binary",  # Exposure - complete
  "cohort_id"          # Data indicator - complete
)

# Cluster variable (patient ID)
cluster_var <- "e_mrn_deidentified"

# Variables to impute - continuous (use PMM with clustering)
vars_impute_continuous <- c(
  "baseline_VA_logMAR",
  "intermediate_VA_logMAR", 
  "outcome_VA_logMAR",
  "age",
  "CCI",
  "DCSI"
)

# Variables to impute - binary
vars_impute_binary <- c(
  "catsurg_before_hitplus_cat"
)

# Variables to impute - categorical
vars_impute_categorical <- c(
  "gender_cat",
  "race_ethnic_cat",
  "insurance_cat"
)

# Variables to impute - ordinal
vars_impute_ordinal <- c(
  "person_dr"  # 0=No_DR, 1=NPDR, 2=PDR
)

# Treatment variables (binary, likely complete)
vars_treatment <- c(
  "other_inject",
  "anti_VEGF",
  "focal_laser_flag",
  "PRP_flag"
)

# Comorbidity indicators (binary factors from step 1)
vars_comorbidity <- c(
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat"
)

# All variables for imputation dataset
all_vars <- c(
  cluster_var,
  vars_no_impute_no_predict,
  vars_no_impute_yes_predict,
  vars_impute_continuous,
  vars_impute_binary,
  vars_impute_categorical,
  vars_impute_ordinal,
  vars_treatment,
  vars_comorbidity
)
```

## Prepare Data for MICE

```{r prepare-data}
# Select variables and clean up (rely on step 1 outputs)
df_mice <- df_original |>
  select(any_of(all_vars)) |>
  mutate(
    # Ensure person_dr is ordered factor
    person_dr = if (is.numeric(person_dr)) {
      factor(person_dr, levels = c(0L, 1L, 2L), labels = c("No_DR", "NPDR", "PDR"), ordered = TRUE)
    } else {
      factor(person_dr, levels = c("No_DR", "NPDR", "PDR"), ordered = TRUE)
    },
    # Clean factor levels (remove spaces for formula compatibility)
    across(where(is.character), ~ factor(gsub("[^[:alnum:]_]", "_", .x)))
  ) |>
  dplyr::group_by(e_mrn_deidentified) |>
  dplyr::mutate(
    n_eyes = dplyr::n(),
    eye_index = dplyr::row_number(),
    fellow_baseline_VA = dplyr::if_else(n_eyes == 2, baseline_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_intermediate_VA = dplyr::if_else(n_eyes == 2, intermediate_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_outcome_VA = dplyr::if_else(n_eyes == 2, outcome_VA_logMAR[3 - eye_index], as.numeric(NA)),
    fellow_dr = dplyr::if_else(n_eyes == 2,
                               person_dr[3 - eye_index],
                               factor(NA, levels = levels(person_dr), ordered = TRUE))
  ) |>
  dplyr::ungroup() |>
  dplyr::select(-eye_index)

# Check missing data pattern
log_message("Missing data summary:")
missing_summary <- df_mice |>
  summarise(across(everything(), ~ sum(is.na(.)))) |>
  pivot_longer(everything(), names_to = "variable", values_to = "n_missing") |>
  filter(n_missing > 0) |>
  arrange(desc(n_missing))

print(missing_summary)
```

## Setup Imputation Methods

```{r setup-methods}
# Initialize mice to get default setup
ini <- mice(df_mice, maxit = 0, printFlag = FALSE)

# Set up methods vector
methods <- ini$method

# No imputation for these
methods[intersect(vars_no_impute_no_predict, names(methods))] <- ""
methods[intersect(vars_no_impute_yes_predict, names(methods))] <- ""
if (cluster_var %in% names(methods)) {
  methods[cluster_var] <- ""
}

# Continuous variables - use 2l.pmm for clustering
for (var in vars_impute_continuous) {
  if (var %in% names(methods)) {
    methods[var] <- "2l.pmm"
  }
}

# Binary variables
for (var in vars_impute_binary) {
  if (var %in% names(methods)) {
    methods[var] <- "logreg"
  }
}

# Categorical variables  
for (var in vars_impute_categorical) {
  if (var %in% names(methods)) {
    methods[var] <- "polyreg"
  }
}

# Ordinal variables
for (var in vars_impute_ordinal) {
  if (var %in% names(methods)) {
    methods[var] <- "polr"
  }
}

# Treatment and comorbidity variables (likely complete, but set methods just in case)
for (var in c(vars_treatment, vars_comorbidity)) {
  if (var %in% names(methods)) {
    methods[var] <- "logreg"
  }
}

# Fellow-eye variables: predictors only (do not impute)
vars_fellow <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr", "n_eyes")
for (var in vars_fellow) {
  if (var %in% names(methods)) {
    methods[var] <- ""
  }
}

log_message("Imputation methods configured")
```

## Setup Predictor Matrix

```{r setup-predictors}
# Use quickpred to get smart predictor selection
# This selects predictors with correlation >= 0.10 (R^2 >= 0.01)
# As recommended in Van Buuren's MICE documentation

# Variables that should always be included as predictors (key variables)
vars_always_predict <- c(
  "ever_lapse_binary",      # Exposure
  "baseline_VA_logMAR",     # Key outcome
  "outcome_VA_logMAR",      # Key outcome
  "age",                    # Demographics
  "gender_cat",
  "CCI", "DCSI",           # Comorbidities
  "anti_VEGF"              # Key treatment
)

# Variables that should never be predictors
vars_never_predict <- c(
  vars_no_impute_no_predict
)

# Generate predictor matrix using quickpred
pred_matrix <- quickpred(
  df_mice, 
  exclude = vars_never_predict,
  include = vars_always_predict,
  mincor = 0.10,  # Minimum correlation for inclusion
  method = "pearson"
)

# Mark cluster variable for 2-level imputation
# -2 indicates this is the cluster ID
pred_matrix[cluster_var, ] <- 0  # Don't predict the cluster ID

# For 2l.pmm variables, ensure cluster is marked
vars_2l <- names(methods)[methods == "2l.pmm"]
for (var in vars_2l) {
  pred_matrix[var, cluster_var] <- -2
}

# Ensure temporal ordering for VA variables
# Baseline can predict intermediate and outcome
pred_matrix["intermediate_VA_logMAR", "baseline_VA_logMAR"] <- 1
pred_matrix["outcome_VA_logMAR", "baseline_VA_logMAR"] <- 1

# Intermediate can predict outcome  
pred_matrix["outcome_VA_logMAR", "intermediate_VA_logMAR"] <- 1

# Prevent reverse-time prediction
pred_matrix["baseline_VA_logMAR", "outcome_VA_logMAR"] <- 0
pred_matrix["baseline_VA_logMAR", "intermediate_VA_logMAR"] <- 0
pred_matrix["intermediate_VA_logMAR", "outcome_VA_logMAR"] <- 0

# Fellow-eye predictors
fellow_vars <- c("fellow_baseline_VA", "fellow_intermediate_VA", "fellow_outcome_VA", "fellow_dr")
for (v in fellow_vars) {
  if (v %in% rownames(pred_matrix)) {
    pred_matrix[v, ] <- 0
  }
}
if ("fellow_baseline_VA" %in% colnames(pred_matrix)) {
  pred_matrix["baseline_VA_logMAR", "fellow_baseline_VA"] <- 1
}
if ("fellow_intermediate_VA" %in% colnames(pred_matrix)) {
  pred_matrix["intermediate_VA_logMAR", "fellow_intermediate_VA"] <- 1
}
if ("fellow_outcome_VA" %in% colnames(pred_matrix)) {
  pred_matrix["outcome_VA_logMAR", "fellow_outcome_VA"] <- 1
}
if ("fellow_dr" %in% colnames(pred_matrix) && "person_dr" %in% rownames(pred_matrix)) {
  pred_matrix["person_dr", "fellow_dr"] <- 1
}

log_message("Predictor matrix configured with temporal constraints")
```

## Run Imputation

```{r run-imputation}
log_message("Starting MICE imputation...")

# Run the imputation
imp <- mice(
  df_mice, 
  m = n_imputations,
  method = methods,
  predictorMatrix = pred_matrix,
  maxit = n_iterations,
  seed = seed,
  printFlag = TRUE
)

log_message("MICE imputation completed")

# Check for logged events
if (!is.null(imp$loggedEvents)) {
  log_message("Logged events during imputation:")
  print(imp$loggedEvents)
}

# Save convergence plot
pdf(file.path(reanalysis_data_dir, "mice_convergence_clean.pdf"))
plot(imp)
dev.off()
```

## Extract and Save Results

```{r save-results}
# Extract all imputed datasets
imputed_list <- complete(imp, action = "all")

# Apply any derived variables needed
imputed_list <- lapply(imputed_list, function(data) {
  data |>
    mutate(
      # Create binary outcome from logMAR
      outcome_va_vi_binary = ifelse(outcome_VA_logMAR >= 1.0, 1, 0)
    )
})

# Save the mice object
saveRDS(imp, file.path(reanalysis_data_dir, "mice_results_clean.rds"))
log_message(paste0("Saved MICE object to: mice_results_clean.rds"))

# Save the list of complete datasets
saveRDS(imputed_list, file.path(reanalysis_data_dir, "imputed_datasets_clean.rds"))
log_message(paste0("Saved ", length(imputed_list), " imputed datasets"))

# Create summary
summary_stats <- data.frame(
  Metric = c(
    "Original observations",
    "Unique patients",
    "Variables in imputation", 
    "Number of imputations",
    "Iterations per imputation",
    "Variables with missing data",
    "Total missing values"
  ),
  Value = c(
    nrow(df_mice),
    n_patients,
    ncol(df_mice),
    n_imputations,
    n_iterations,
    sum(colSums(is.na(df_mice)) > 0),
    sum(is.na(df_mice))
  )
)

kable(summary_stats, caption = "Imputation Summary")

# Finalize logging
final_log <- finalize_log(success = TRUE)
```

**Analysis complete. Log file:** `r final_log`

## Next Steps

Run `reanalysis-3-imputation_diagnostics.qmd` to assess imputation quality.