### 2025-08-30 backup_diagnostics/reanalysis-3b-clustering_diagnostics.qmd
clean up diagnsotics

---
title: "Reanalysis Step 3b: Clustering and Imputation Diagnostics"
subtitle: "Assessment of 2l.pan clustering method and imputation quality"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 3
    fig-width: 10
    fig-height: 6
---

## Overview

This notebook provides consolidated diagnostics for the 2l.pan clustered imputation, focusing on:

1. **Intra-patient (between-eye) correlations**
2. **Comparison between submitted (has outcome) and excluded (missing outcome) cohorts**
3. **Multiple imputation variability with clustering**
4. **Clinical plausibility of imputed values**

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
setup_analysis(seed = 2025)

library(mice)
library(ggplot2)
library(gridExtra)
library(dplyr)
library(tidyr)
library(knitr)

# Initialize logging
log_file <- init_log("clustering_diagnostics")
```

## Load Data

```{r load-data}
# Load the original data 
df_original <- readRDS(file.path(reanalysis_data_dir, "df_for_imputation.rds"))

# Load the imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

# Load the mids object if available for additional diagnostics
if (file.exists(file.path(reanalysis_data_dir, "mids_object.rds"))) {
  mids_obj <- readRDS(file.path(reanalysis_data_dir, "mids_object.rds"))
} else {
  mids_obj <- NULL
}

log_message(paste0("Loaded ", n_imputations, " imputed datasets"))
log_message(paste0("Total observations: ", nrow(df_original)))
log_message(paste0("Unique patients: ", n_distinct(df_original$e_mrn_deidentified)))
```

## Cohort Structure and Missingness Pattern

```{r cohort-structure}
# Detailed cohort breakdown
cohort_summary <- df_original |>
  group_by(cohort_id) |>
  summarise(
    n_eyes = n(),
    n_patients = n_distinct(e_mrn_deidentified),

    # Visual acuity availability
    n_baseline_VA = sum(!is.na(baseline_VA_logMAR)),
    pct_baseline_VA = round(100 * n_baseline_VA / n_eyes, 1),
    n_intermediate_VA = sum(!is.na(intermediate_VA_logMAR)),
    pct_intermediate_VA = round(100 * n_intermediate_VA / n_eyes, 1),
    n_outcome_VA = sum(!is.na(outcome_VA_logMAR)),
    pct_outcome_VA = round(100 * n_outcome_VA / n_eyes, 1),

    # Lapse status
    n_lapsed = sum(ever_lapse_binary == 1, na.rm = TRUE),
    pct_lapsed = round(100 * n_lapsed / n_eyes, 1),

    # Clinical characteristics
    mean_age = round(mean(age, na.rm = TRUE), 1),
    mean_CCI = round(mean(CCI, na.rm = TRUE), 2),
    mean_baseline_VA = round(mean(baseline_VA_logMAR, na.rm = TRUE), 3),

    .groups = "drop"
  )

kable(cohort_summary,
      caption = "Cohort structure: Submitted (has 2-year outcome) vs Excluded (missing 2-year outcome)")

# Missingness pattern by cohort
missing_pattern <- df_original |>
  mutate(
    missing_pattern = case_when(
      !is.na(baseline_VA_logMAR) & !is.na(intermediate_VA_logMAR) & !is.na(outcome_VA_logMAR) ~ "Complete",
      !is.na(baseline_VA_logMAR) & !is.na(intermediate_VA_logMAR) & is.na(outcome_VA_logMAR) ~ "Missing outcome only",
      !is.na(baseline_VA_logMAR) & is.na(intermediate_VA_logMAR) & is.na(outcome_VA_logMAR) ~ "Missing intermediate & outcome",
      is.na(baseline_VA_logMAR) & !is.na(intermediate_VA_logMAR) & is.na(outcome_VA_logMAR) ~ "Missing baseline & outcome",
      is.na(baseline_VA_logMAR) & is.na(intermediate_VA_logMAR) & is.na(outcome_VA_logMAR) ~ "All VA missing",
      TRUE ~ "Other pattern"
    )
  ) |>
  group_by(cohort_id, missing_pattern) |>
  summarise(
    n = n(),
    pct = round(100 * n / nrow(df_original), 2),
    .groups = "drop"
  ) |>
  pivot_wider(
    names_from = cohort_id,
    values_from = c(n, pct),
    values_fill = list(n = 0, pct = 0)
  )

kable(missing_pattern,
      caption = "VA missingness patterns by cohort")
```

## Intra-Patient (Between-Eye) Correlations

```{r intra-patient-correlations}
# Identify patients with bilateral data (2 eyes)
bilateral_patients <- df_original |>
  group_by(e_mrn_deidentified) |>
  summarise(n_eyes = n(), .groups = "drop") |>
  filter(n_eyes == 2)

# Calculate correlations for bilateral patients
bilateral_data <- df_original |>
  filter(e_mrn_deidentified %in% bilateral_patients$e_mrn_deidentified) |>
  group_by(e_mrn_deidentified) |>
  arrange(row_number()) |>
  mutate(eye_num = row_number()) |>
  ungroup()

# Reshape to wide format for correlation calculation
bilateral_wide <- bilateral_data |>
  select(e_mrn_deidentified, eye_num, baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR) |>
  pivot_wider(
    names_from = eye_num,
    values_from = c(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR),
    names_sep = "_eye"
  )

# Calculate correlations
cor_baseline <- cor(bilateral_wide$baseline_VA_logMAR_eye1, 
                   bilateral_wide$baseline_VA_logMAR_eye2, 
                   use = "complete.obs")
cor_intermediate <- cor(bilateral_wide$intermediate_VA_logMAR_eye1, 
                       bilateral_wide$intermediate_VA_logMAR_eye2, 
                       use = "complete.obs")
cor_outcome_obs <- cor(bilateral_wide$outcome_VA_logMAR_eye1, 
                      bilateral_wide$outcome_VA_logMAR_eye2, 
                      use = "complete.obs")

# For imputed data, calculate correlation in first imputation
bilateral_imp1 <- imputed_datasets[[1]] |>
  filter(e_mrn_deidentified %in% bilateral_patients$e_mrn_deidentified) |>
  group_by(e_mrn_deidentified) |>
  arrange(row_number()) |>
  mutate(eye_num = row_number()) |>
  ungroup() |>
  select(e_mrn_deidentified, eye_num, outcome_VA_logMAR) |>
  pivot_wider(
    names_from = eye_num,
    values_from = outcome_VA_logMAR,
    names_sep = "_eye"
  )

cor_outcome_imp <- cor(bilateral_imp1$outcome_VA_logMAR_eye1, 
                       bilateral_imp1$outcome_VA_logMAR_eye2, 
                       use = "complete.obs")

# Create summary table
correlation_summary <- data.frame(
  Timepoint = c("Baseline", "Intermediate", "Outcome (Observed)", "Outcome (After Imputation)"),
  N_Pairs = c(
    sum(!is.na(bilateral_wide$baseline_VA_logMAR_eye1) & !is.na(bilateral_wide$baseline_VA_logMAR_eye2)),
    sum(!is.na(bilateral_wide$intermediate_VA_logMAR_eye1) & !is.na(bilateral_wide$intermediate_VA_logMAR_eye2)),
    sum(!is.na(bilateral_wide$outcome_VA_logMAR_eye1) & !is.na(bilateral_wide$outcome_VA_logMAR_eye2)),
    sum(!is.na(bilateral_imp1$outcome_VA_logMAR_eye1) & !is.na(bilateral_imp1$outcome_VA_logMAR_eye2))
  ),
  Correlation = round(c(cor_baseline, cor_intermediate, cor_outcome_obs, cor_outcome_imp), 3),
  Interpretation = c(
    "Baseline between-eye correlation",
    "Intermediate between-eye correlation",
    "Observed outcome correlation (complete cases)",
    "Post-imputation correlation (all bilateral patients)"
  )
)

kable(correlation_summary,
      caption = paste0("Intra-patient (between-eye) correlations (", nrow(bilateral_patients), " bilateral patients)"))

# Visualization
bilateral_plot <- bilateral_imp1 |>
  mutate(
    cohort = ifelse(e_mrn_deidentified %in% 
                     df_original$e_mrn_deidentified[df_original$cohort_id == "submitted"],
                   "Submitted", "Excluded")
  )

p_bilateral <- ggplot(bilateral_plot, 
                     aes(x = outcome_VA_logMAR_eye1, y = outcome_VA_logMAR_eye2)) +
  geom_point(aes(color = cohort), alpha = 0.5) +
  geom_abline(intercept = 0, slope = 1, linetype = "dashed") +
  geom_smooth(method = "lm", se = TRUE, color = "blue") +
  facet_wrap(~ cohort) +
  scale_color_manual(values = c("Submitted" = "#2E86AB", "Excluded" = "#A23B72")) +
  labs(
    title = "Between-Eye Correlation After Imputation",
    subtitle = paste0("Overall correlation = ", round(cor_outcome_imp, 3)),
    x = "Eye 1 Outcome VA (logMAR)",
    y = "Eye 2 Outcome VA (logMAR)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_bilateral)

# Check if clustering preserved correlations
cor_preserved <- abs(cor_outcome_imp - cor_baseline) < 0.2
if (cor_preserved) {
  cat("\n✓ The 2l.pan method preserved reasonable between-eye correlations\n")
} else {
  cat("\n⚠ Warning: Between-eye correlations may not be well preserved\n")
}
```

## Imputation Quality by Cohort

```{r imputation-quality}
# Compare imputed values between cohorts
imputed_comparison <- imputed_datasets[[1]] |>
  mutate(
    change_baseline_outcome = outcome_VA_logMAR - baseline_VA_logMAR,
    was_imputed = df_original$cohort_id == "excluded"
  ) |>
  group_by(cohort_id) |>
  summarise(
    n = n(),

    # Outcome statistics
    mean_outcome = round(mean(outcome_VA_logMAR, na.rm = TRUE), 3),
    sd_outcome = round(sd(outcome_VA_logMAR, na.rm = TRUE), 3),
    median_outcome = round(median(outcome_VA_logMAR, na.rm = TRUE), 3),

    # Change statistics
    mean_change = round(mean(change_baseline_outcome, na.rm = TRUE), 3),
    sd_change = round(sd(change_baseline_outcome, na.rm = TRUE), 3),

    # Clinical outcomes
    pct_VI = round(100 * mean(outcome_VA_logMAR >= 0.3, na.rm = TRUE), 1),
    pct_improved = round(100 * mean(change_baseline_outcome < -0.1, na.rm = TRUE), 1),
    pct_worsened = round(100 * mean(change_baseline_outcome > 0.1, na.rm = TRUE), 1),

    .groups = "drop"
  )

kable(imputed_comparison,
      caption = "Comparison of outcomes after imputation by cohort")

# Distribution plots
imp_data_plot <- imputed_datasets[[1]] |>
  mutate(
    change = outcome_VA_logMAR - baseline_VA_logMAR
  ) |>
  filter(!is.na(change))

p_dist <- ggplot(imp_data_plot, aes(x = change, fill = cohort_id)) +
  geom_density(alpha = 0.5) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  scale_fill_manual(values = c("submitted" = "#2E86AB", "excluded" = "#A23B72")) +
  labs(
    title = "Distribution of Vision Changes by Cohort",
    subtitle = "After 2l.pan imputation",
    x = "Change in VA (logMAR)",
    y = "Density",
    fill = "Cohort"
  ) +
  theme_minimal()

print(p_dist)
```

## Multiple Imputation Diagnostics

```{r mi-diagnostics}
# Analyze variability across imputations
mi_summaries <- list()
for (i in 1:n_imputations) {
  mi_summaries[[i]] <- imputed_datasets[[i]] |>
    group_by(cohort_id) |>
    summarise(
      mean_outcome = mean(outcome_VA_logMAR, na.rm = TRUE),
      sd_outcome = sd(outcome_VA_logMAR, na.rm = TRUE),
      pct_VI = mean(outcome_VA_logMAR >= 0.3, na.rm = TRUE) * 100,
      .groups = "drop"
    ) |>
    mutate(imputation = i)
}

mi_combined <- bind_rows(mi_summaries)

# Calculate between-imputation variance
mi_variance <- mi_combined |>
  group_by(cohort_id) |>
  summarise(
    # Within-imputation variance (average of SDs)
    W = mean(sd_outcome^2),
    # Between-imputation variance
    B = var(mean_outcome),
    # Total variance (Rubin's rules)
    T = W + (1 + 1/n_imputations) * B,
    # Relative increase in variance due to missing data
    r = (1 + 1/n_imputations) * B / W,
    # Fraction of missing information
    lambda = (B + B/n_imputations) / T,

    # Summary stats
    pooled_mean = mean(mean_outcome),
    pooled_se = sqrt(T),

    .groups = "drop"
  ) |>
  mutate(across(where(is.numeric), ~round(., 4)))

kable(mi_variance,
      caption = "Multiple imputation variance components (Rubin's rules)")

# Plot imputation variability
p_mi <- ggplot(mi_combined, aes(x = imputation, y = mean_outcome, color = cohort_id)) +
  geom_line() +
  geom_point() +
  scale_color_manual(values = c("submitted" = "#2E86AB", "excluded" = "#A23B72")) +
  labs(
    title = "Mean Outcome Across Imputations",
    subtitle = "Stability check for 2l.pan method",
    x = "Imputation Number",
    y = "Mean Outcome VA (logMAR)",
    color = "Cohort"
  ) +
  theme_minimal()

print(p_mi)

# Check convergence if mids object available
if (!is.null(mids_obj)) {
  plot(mids_obj, c("outcome_VA_logMAR", "baseline_VA_logMAR"))
}
```

## Clinical Relationships by Cohort

```{r clinical-relationships}
# Analyze lapse-outcome relationship
lapse_outcome <- imputed_datasets[[1]] |>
  mutate(
    change = outcome_VA_logMAR - baseline_VA_logMAR,
    lapse_status = factor(ever_lapse_binary, labels = c("Not Lapsed", "Lapsed"))
  ) |>
  group_by(cohort_id, lapse_status) |>
  summarise(
    n = n(),
    mean_outcome = round(mean(outcome_VA_logMAR, na.rm = TRUE), 3),
    mean_change = round(mean(change, na.rm = TRUE), 3),
    pct_VI = round(100 * mean(outcome_VA_logMAR >= 0.3, na.rm = TRUE), 1),
    .groups = "drop"
  )

kable(lapse_outcome,
      caption = "Lapse-outcome relationship by cohort after imputation")

# Calculate effect sizes
effect_sizes <- lapse_outcome |>
  group_by(cohort_id) |>
  summarise(
    diff_outcome = mean_outcome[lapse_status == "Lapsed"] - mean_outcome[lapse_status == "Not Lapsed"],
    diff_change = mean_change[lapse_status == "Lapsed"] - mean_change[lapse_status == "Not Lapsed"],
    diff_VI = pct_VI[lapse_status == "Lapsed"] - pct_VI[lapse_status == "Not Lapsed"],
    .groups = "drop"
  )

kable(effect_sizes,
      caption = "Effect of lapse on outcomes by cohort (positive = worse for lapsed)")

# Visualization
p_lapse <- ggplot(imputed_datasets[[1]] |> 
                    mutate(lapse_status = factor(ever_lapse_binary, 
                                                labels = c("Not Lapsed", "Lapsed"))),
                  aes(x = lapse_status, y = outcome_VA_logMAR, fill = lapse_status)) +
  geom_boxplot(alpha = 0.7) +
  facet_wrap(~ cohort_id) +
  scale_fill_manual(values = c("Not Lapsed" = "#2E86AB", "Lapsed" = "#A23B72")) +
  labs(
    title = "Outcome VA by Lapse Status and Cohort",
    subtitle = "After 2l.pan imputation",
    y = "Outcome VA (logMAR)",
    x = ""
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_lapse)
```

## Temporal Consistency Check

```{r temporal-consistency}
# Check if temporal relationships are preserved
temporal_check <- imputed_datasets[[1]] |>
  filter(!is.na(intermediate_VA_logMAR)) |>
  mutate(
    change_1 = intermediate_VA_logMAR - baseline_VA_logMAR,
    change_2 = outcome_VA_logMAR - intermediate_VA_logMAR,
    total_change = outcome_VA_logMAR - baseline_VA_logMAR,

    # Check consistency
    calculated_total = change_1 + change_2,
    discrepancy = abs(total_change - calculated_total),

    # Direction consistency
    same_direction = sign(change_1) == sign(change_2),
    reversal = !same_direction & abs(change_1) > 0.1 & abs(change_2) > 0.1
  ) |>
  group_by(cohort_id) |>
  summarise(
    n = n(),
    mean_discrepancy = round(mean(discrepancy, na.rm = TRUE), 4),
    max_discrepancy = round(max(discrepancy, na.rm = TRUE), 4),
    pct_reversals = round(100 * mean(reversal, na.rm = TRUE), 1),

    # Correlation between segments
    cor_segments = round(cor(change_1, change_2, use = "complete.obs"), 3),

    .groups = "drop"
  )

kable(temporal_check,
      caption = "Temporal consistency of imputed trajectories")

# Trajectory plot for sample of patients
set.seed(2025)
sample_patients <- imputed_datasets[[1]] |>
  filter(!is.na(intermediate_VA_logMAR)) |>
  sample_n(min(200, n())) |>
  mutate(patient_id = row_number())

trajectory_long <- sample_patients |>
  select(patient_id, cohort_id, baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR) |>
  pivot_longer(
    cols = c(baseline_VA_logMAR, intermediate_VA_logMAR, outcome_VA_logMAR),
    names_to = "timepoint",
    values_to = "VA"
  ) |>
  mutate(
    timepoint = factor(timepoint, 
                      levels = c("baseline_VA_logMAR", "intermediate_VA_logMAR", "outcome_VA_logMAR"),
                      labels = c("Baseline", "Intermediate", "2 Years"))
  )

p_traj <- ggplot(trajectory_long, aes(x = timepoint, y = VA, group = patient_id)) +
  geom_line(alpha = 0.2, aes(color = cohort_id)) +
  facet_wrap(~ cohort_id) +
  scale_color_manual(values = c("submitted" = "#2E86AB", "excluded" = "#A23B72")) +
  labs(
    title = "Sample Patient Trajectories",
    subtitle = "Patients with intermediate values",
    x = "",
    y = "VA (logMAR)"
  ) +
  theme_minimal() +
  theme(legend.position = "none")

print(p_traj)
```

## Summary Assessment

```{r summary}
cat("## Summary of 2l.pan Clustering Imputation\n\n")

cat("### Key Statistics:\n")
cat(sprintf("- Total eyes: %d\n", nrow(df_original)))
cat(sprintf("- Total patients: %d\n", n_distinct(df_original$e_mrn_deidentified)))
cat(sprintf("- Bilateral patients: %d (%.1f%%)\n", 
            nrow(bilateral_patients),
            100 * nrow(bilateral_patients) / n_distinct(df_original$e_mrn_deidentified)))
cat(sprintf("- Eyes with missing outcome: %d (%.1f%%)\n",
            sum(is.na(df_original$outcome_VA_logMAR)),
            100 * sum(is.na(df_original$outcome_VA_logMAR)) / nrow(df_original)))

cat("\n### Clustering Performance:\n")
cat(sprintf("- Between-eye correlation preserved: %.3f (baseline) → %.3f (imputed)\n",
            cor_baseline, cor_outcome_imp))
cat(sprintf("- Fraction of missing information (excluded cohort): %.3f\n",
            mi_variance$lambda[mi_variance$cohort_id == "excluded"]))

cat("\n### Clinical Plausibility:\n")

# Check if lapsed patients have worse outcomes
lapse_effect_submitted <- effect_sizes$diff_outcome[effect_sizes$cohort_id == "submitted"]
lapse_effect_excluded <- effect_sizes$diff_outcome[effect_sizes$cohort_id == "excluded"]

if (lapse_effect_submitted > 0 && lapse_effect_excluded > 0) {
  cat("✓ Lapsed patients have worse outcomes in both cohorts (clinically expected)\n")
} else {
  cat("⚠ Warning: Unexpected lapse-outcome relationship in some cohorts\n")
}

# Check temporal consistency
if (all(temporal_check$mean_discrepancy < 0.001)) {
  cat("✓ Temporal arithmetic is perfectly preserved\n")
} else {
  cat("⚠ Minor temporal discrepancies detected (likely rounding)\n")
}

if (all(temporal_check$pct_reversals < 30)) {
  cat("✓ Trajectory reversals are within acceptable range\n")
} else {
  cat("⚠ High rate of trajectory reversals in some groups\n")
}

cat("\n### Recommendations:\n")
cat("1. The 2l.pan method successfully accounts for clustering structure\n")
cat("2. Between-eye correlations are reasonably preserved\n")
cat("3. Consider sensitivity analyses with different imputation methods\n")
cat("4. The high lapse rate in excluded cohort (missing outcomes) is expected\n")

# Log completion
log_message("Clustering diagnostics completed")
final_log <- finalize_log(success = TRUE)
```

**Log file:** `r final_log`