### 2025-09-01 reanalysis-5-outcome_analysis-interaction.qmd
dont cache

---
title: "Reanalysis Step 5: Outcome Analysis with Three-Way Interaction"
subtitle: "IPTW-weighted analysis of lapse × DR severity × treatment interactions"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formulas and helper functions
setup_analysis(seed = 2025)

# Additional packages
library(emmeans)
library(margins)
library(survey)
library(ggplot2)
library(gridExtra)
library(knitr)
library(tidyr)
library(dplyr)
library(mice)  # For pool() function with Barnard-Rubin df adjustment
library(mitools)  # For additional MI utilities if needed

# ANALYSIS MODE: Use same as imputation/PS
analysis_mode <- "final"  # Should match imputation and PS modes

# CACHE SETTINGS - Set these to TRUE to force recalculation
force_refit_models <- FALSE  # Set to TRUE to re-fit all models
force_recalc_emmeans <- FALSE  # Set to TRUE to recalculate marginal means

# Set parameters based on mode
if (analysis_mode == "test") {
  max_imputations_to_process <- 2  # Process only 2 imputations in test mode
} else {
  max_imputations_to_process <- NULL  # Process all available imputations
}
```

**Analysis mode:** `r analysis_mode`  
**Imputations to process:** `r if(!is.null(max_imputations_to_process)) max_imputations_to_process else "all available"`

## Analysis Overview

This analysis examines the three-way interaction between:
- **Lapse status** (primary exposure)
- **DR severity** (effect modifier)
- **Treatment status** (effect modifier)

The model uses:
- Multiple imputation to handle missing data
- IPTW weights from propensity scores
- Survey methods for clustered data (multiple eyes per patient)
- Pooling via Rubin's rules across imputations

## Load Data and Weights

```{r load-data}
# Load the imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_available <- length(imputed_datasets)

# Load the twang PS results
ps_results <- readRDS(file.path(reanalysis_data_dir, "ps_results_twang.rds"))
n_ps_available <- length(ps_results$twang_results)

# Determine how many imputations to actually process
if (!is.null(max_imputations_to_process)) {
  n_imputations <- min(max_imputations_to_process, n_available, n_ps_available)
  # Subset the datasets if needed
  if (n_imputations < n_available) {
    imputed_datasets <- imputed_datasets[1:n_imputations]
  }
} else {
  n_imputations <- min(n_available, n_ps_available)
}

# Extract weights for each imputation
# Ensure we don't exceed available PS results
n_ps_weights <- length(ps_results$twang_results)
if (n_ps_weights < n_imputations) {
  warning("Only ", n_ps_weights, " PS weight sets available for ", n_imputations, " imputations")
  stop("Mismatch between imputations and PS results. Ensure PS analysis was run with same number of imputations.")
}

for (i in 1:n_imputations) {
  imputed_datasets[[i]]$iptw_weight <- ps_results$twang_results[[i]]$weights
}
```

**Available imputed datasets:** `r n_available`  
**Available PS results:** `r n_ps_available`  
**Processing:** `r n_imputations` **datasets** `r if(analysis_mode == "test") "(test mode limit)" else "(all available)"`

```{r continue-load}
# Summary of first dataset for verification
first_data <- imputed_datasets[[1]]

dataset_summary <- data.frame(
  Metric = c("Imputed datasets loaded", "N observations", "N unique patients", 
             "Mean IPTW weight", "Weight range"),
  Value = c(n_imputations,
            nrow(first_data),
            length(unique(first_data$e_mrn_deidentified)),
            round(mean(first_data$iptw_weight), 3),
            paste0("[", round(min(first_data$iptw_weight), 3), ", ",
                   round(max(first_data$iptw_weight), 3), "]"))
)

kable(dataset_summary, caption = "Dataset Summary")
```

## Create Analysis Variables

```{r create-variables}
# Create consistent analysis variables across all imputed datasets
for (i in 1:n_imputations) {
  imputed_datasets[[i]] <- imputed_datasets[[i]] |>
    mutate(
      # Binary treatment indicator (any treatment vs none)
      any_treatment = as.numeric(anti_VEGF == 1 | PRP_flag == 1 | 
                                 other_inject == 1 | focal_laser_flag == 1),
      any_treatment = factor(any_treatment, levels = c(0, 1), 
                            labels = c("No_Treatment", "Any_Treatment")),

      # Ensure DR severity is properly factored
      dr_severity = factor(dr_severity,
                          levels = c("No_DR", "NPDR", "PDR")),

      # Ensure lapse is properly coded
      ever_lapse_binary = factor(ever_lapse_binary, levels = c(0, 1),
                                 labels = c("No_Lapse", "Lapsed"))
    )
}

```

### Variable Distributions in First Imputed Dataset

**DR Severity:**
```{r dr-severity-dist}
dr_table <- table(imputed_datasets[[1]]$dr_severity)
dr_df <- data.frame(
  Category = names(dr_table),
  Count = as.numeric(dr_table),
  Percentage = paste0(round(100 * as.numeric(dr_table) / sum(dr_table), 1), "%")
)
kable(dr_df, align = c('l', 'r', 'r'))
```

**Treatment Status:**
```{r treatment-dist}
treatment_table <- table(imputed_datasets[[1]]$any_treatment)
treatment_df <- data.frame(
  Treatment = c("No Treatment", "Treatment"),
  Count = as.numeric(treatment_table),
  Percentage = paste0(round(100 * as.numeric(treatment_table) / sum(treatment_table), 1), "%")
)
kable(treatment_df, align = c('l', 'r', 'r'))
```

**Lapse Status:**
```{r lapse-dist}
lapse_table <- table(imputed_datasets[[1]]$ever_lapse_binary)
lapse_df <- data.frame(
  Lapse_Status = c("No Lapse", "Lapsed"),
  Count = as.numeric(lapse_table),
  Percentage = paste0(round(100 * as.numeric(lapse_table) / sum(lapse_table), 1), "%")
)
kable(lapse_df, align = c('l', 'r', 'r'))
```

**Three-way Cross-tabulation:**
```{r crosstab}
# Create cross-tabulation and convert to data frame for proper display
crosstab <- with(imputed_datasets[[1]], 
                 table(dr_severity, any_treatment, ever_lapse_binary))

# Convert to data frame format using as.data.frame.table
crosstab_df <- as.data.frame.table(crosstab)
names(crosstab_df) <- c("DR_Severity", "Treatment", "Lapse_Status", "Freq")

# Rename Freq to Count and ensure it's numeric
crosstab_df$Count <- as.numeric(crosstab_df$Freq)
crosstab_df$Freq <- NULL

# Calculate total for percentages
total_n <- sum(crosstab_df$Count)

# Pivot to wide format for better readability with percentages
crosstab_wide <- crosstab_df |>
  pivot_wider(names_from = Lapse_Status, 
              values_from = Count,
              values_fill = 0) |>
  arrange(DR_Severity, Treatment) |>
  mutate(
    Total = No_Lapse + Lapsed,
    No_Lapse_Pct = paste0(round(100 * No_Lapse / Total, 1), "%"),
    Lapsed_Pct = paste0(round(100 * Lapsed / Total, 1), "%"),
    Row_Pct = paste0(round(100 * Total / total_n, 1), "%")
  ) |>
  select(DR_Severity, Treatment, 
         No_Lapse, No_Lapse_Pct,
         Lapsed, Lapsed_Pct,
         Total, Row_Pct)

kable(crosstab_wide, 
      caption = "Distribution of patients by DR severity, treatment, and lapse status",
      col.names = c("DR Severity", "Treatment", 
                    "No Lapse (n)", "No Lapse (%)",
                    "Lapsed (n)", "Lapsed (%)",
                    "Total (n)", "Row %"),
      align = c('l', 'l', 'r', 'r', 'r', 'r', 'r', 'r'))
```

## Define Model Formula

```{r define-formula}
# Get the interaction formulas from functions.R
interaction_formulas <- get_analysis_formulas_reanalysis_interaction()

# Use the full formula which includes three-way interaction
interaction_formula <- interaction_formulas$full

```

### Model Specification

**Formula:**
```{r show-formula}
print(interaction_formula)
```

**Interaction Structure:**
- Main effects: 3 (lapse, DR severity [2 df], treatment)
- Two-way interactions: 3 (lapse×DR [2 df], lapse×treatment [1 df], DR×treatment [2 df])
- Three-way interaction: 1 (lapse×DR×treatment [2 df])
- Total interaction df: 7

## Fit Models to Each Imputed Dataset

```{r fit-models}
# Check if models are already cached
models_cache_file <- file.path(reanalysis_data_dir, "fitted_models_interaction.rds")

if (file.exists(models_cache_file) && !force_refit_models) {
  # Load cached models
  cached_models <- readRDS(models_cache_file)
  models <- cached_models$models
  model_summaries <- cached_models$model_summaries
  model_progress <- cached_models$model_progress
  models_loaded_from_cache <- TRUE
  cache_location <- models_cache_file
  message("Models loaded from cache: ", cache_location)
} else {
  # Fit new models
  models_loaded_from_cache <- FALSE
  models <- list()
  model_summaries <- list()

  model_progress <- data.frame(
    Imputation = integer(),
    Converged = logical(),
    ThreeWay_Terms = integer()
  )

  for (i in 1:n_imputations) {

  design_weighted <- svydesign(
    ids = ~e_mrn_deidentified,
    data = imputed_datasets[[i]],
    weights = ~iptw_weight
  )

  model <- svyglm(
    formula = interaction_formula,
    design = design_weighted,
    family = quasibinomial()
  )

  models[[i]] <- model

  coef_summary <- summary(model)$coefficients

  # Check if three-way interaction terms exist
  threeway_terms <- grep("ever_lapse_binary.*:dr_severity.*:any_treatment", 
                         rownames(coef_summary), value = TRUE)

  # Track progress
  model_progress <- rbind(model_progress, data.frame(
    Imputation = i,
    Converged = model$converged,
    ThreeWay_Terms = length(threeway_terms)
  ))

  model_summaries[[i]] <- coef_summary
  }

  # Save to cache
  cached_models <- list(
    models = models,
    model_summaries = model_summaries,
    model_progress = model_progress,
    formula_used = interaction_formula,
    n_imputations = n_imputations,
    timestamp = Sys.time()
  )

  saveRDS(cached_models, models_cache_file)
  cache_location <- models_cache_file
  message("Models fitted and saved to cache: ", cache_location)
}

# Verify we have the expected number of models
if (length(models) != n_imputations) {
  stop("Model list has ", length(models), " models but expected ", n_imputations)
}

# Remove any NULL entries if they exist
models <- models[!sapply(models, is.null)]
if (length(models) == 0) {
  stop("No valid models to pool")
}
```

### Model Fitting Summary

```{r model-summary}
kable(model_progress, caption = "Model fitting progress across imputations")
```

## Pool Results Using Rubin's Rules

```{r pool-results}
# Use mice::pool for robust pooling with Barnard-Rubin df adjustment
# mice::pool automatically applies Barnard-Rubin (1999) correction

# Ensure models is a proper list
if (!is.list(models)) {
  stop("Models must be a list")
}

# Check that all models have the same structure
if (length(models) > 1) {
  coef_names_first <- names(coef(models[[1]]))
  for (i in 2:length(models)) {
    if (!identical(names(coef(models[[i]])), coef_names_first)) {
      warning("Model ", i, " has different coefficients than model 1")
    }
  }
}

# Calculate complete-data degrees of freedom for Barnard-Rubin adjustment
# dfcom = residual degrees of freedom from complete-data model
n_obs <- nrow(models[[1]]$data)
n_params <- length(coef(models[[1]]))
dfcom <- n_obs - n_params  # This is the residual df

# Pool using mice with explicit dfcom for proper Barnard-Rubin df calculation
# mice::pool automatically applies Barnard-Rubin (1999) small-sample correction
pooled_model <- pool(models, dfcom = dfcom)
pooled_summary <- summary(pooled_model)

# Extract results - mice::pool provides df column with Barnard-Rubin adjustment
# Each parameter gets its own df based on its fraction of missing information
pooled_results <- pooled_summary |>
  mutate(
    Term = term,
    Estimate = estimate,
    SE = std.error,
    t_stat = statistic,
    df = df,  # Barnard-Rubin (1999) adjusted degrees of freedom
    P_Value = p.value,
    # Recalculate CIs using t-distribution with proper df
    CI_Lower = estimate - qt(0.975, df) * std.error,
    CI_Upper = estimate + qt(0.975, df) * std.error,
    OR = exp(estimate),
    OR_CI_Lower = exp(CI_Lower),
    OR_CI_Upper = exp(CI_Upper)
  ) |>
  select(Term, Estimate, SE, t_stat, df, P_Value, CI_Lower, CI_Upper, OR, OR_CI_Lower, OR_CI_Upper)

# Ensure proper formatting
rownames(pooled_results) <- NULL

# Construct pooled variance-covariance matrix from the pooled variances
# mice::pool returns a mipo object which stores variances in pooled$t
# We need to create a diagonal matrix since pool() doesn't provide covariances
pooled_variances <- pooled_model$pooled$t
pooled_vcov <- diag(pooled_variances)
rownames(pooled_vcov) <- pooled_model$pooled$term
colnames(pooled_vcov) <- pooled_model$pooled$term

```

### Pooling Diagnostics

```{r pooling-diagnostics}
# Create diagnostics table - now with df from mice::pool
pool_diag <- data.frame(
  Metric = c("Number of imputations", 
             "Number of parameters",
             "Average degrees of freedom (Barnard-Rubin adjusted)",
             "Minimum degrees of freedom",
             "Maximum degrees of freedom"),
  Value = c(n_imputations,
            nrow(pooled_results),
            round(mean(pooled_results$df), 1),
            round(min(pooled_results$df), 1),
            round(max(pooled_results$df), 1))
)

kable(pool_diag, caption = "Pooling diagnostics")

# Show key terms with their df and p-values
key_terms <- grep("ever_lapse", pooled_results$Term, value = TRUE)[1:min(3, length(grep("ever_lapse", pooled_results$Term, value = TRUE)))]
key_results <- pooled_results[pooled_results$Term %in% key_terms, c("Term", "Estimate", "SE", "df", "P_Value")]
key_results$Term <- substr(key_results$Term, 1, 40)
key_results$Estimate <- round(key_results$Estimate, 3)
key_results$SE <- round(key_results$SE, 3)
key_results$df <- round(key_results$df, 1)
key_results$P_Value <- round(key_results$P_Value, 4)

kable(key_results, caption = "Key lapse-related terms with degrees of freedom")

# Three-way interaction terms are included in the model
# but not displayed individually as they are not meaningful on their own
# See subgroup effects and marginal effects sections for interpretation
```

## Joint Tests of Interaction Terms

```{r joint-tests}
pool_chi_square_tests <- function(models, hypothesis_strings, test_name) {

  # Perform test on each imputation
  test_stats <- numeric(length(models))
  df_tests <- numeric(length(models))

  for (i in 1:length(models)) {
    test_result <- linearHypothesis(
      models[[i]],
      hypothesis_strings,
      vcov = vcov(models[[i]])
    )
    test_stats[i] <- test_result$Chisq[2]
    df_tests[i] <- test_result$Df[2]
  }

  # Pool using Li, Raghunathan, and Rubin (1991) method for chi-square tests
  # D-bar: average test statistic
  D_bar <- mean(test_stats)
  df <- df_tests[1]  # Should be same across all imputations

  # For large samples, can use D_bar directly with original df
  # For more accuracy, would need to compute the adjusted statistic
  p_value <- pchisq(D_bar, df = df, lower.tail = FALSE)

  return(list(
    test_name = test_name,
    D_bar = D_bar,
    df = df,
    p_value = p_value
  ))
}

first_model <- models[[1]]

# Test 1: Joint test of three-way interaction
threeway_terms_test <- grep("ever_lapse.*:dr_severity.*:any_treatment", 
                            names(coef(first_model)), value = TRUE)

# Test 1: Three-way interaction
if (length(threeway_terms_test) > 0) {
  hypothesis_3way <- paste0(threeway_terms_test, " = 0")
  pooled_test_3way <- pool_chi_square_tests(models, hypothesis_3way, "Three-way interaction")
}

# Test 2: Joint test of lapse × DR interaction (averaging over treatment)
lapse_dr_terms <- grep("ever_lapse.*:dr_severity(?!.*:any_treatment)", 
                      names(coef(first_model)), value = TRUE, perl = TRUE)

if (length(lapse_dr_terms) > 0) {
  hypothesis_lapse_dr <- paste0(lapse_dr_terms, " = 0")
  pooled_test_lapse_dr <- pool_chi_square_tests(models, hypothesis_lapse_dr, "Lapse × DR interaction")
}

# Test 3: Joint test of lapse × treatment interaction (averaging over DR)
lapse_tx_terms <- grep("ever_lapse.*:any_treatment(?!.*:dr_severity)", 
                      names(coef(first_model)), value = TRUE, perl = TRUE)

if (length(lapse_tx_terms) > 0) {
  hypothesis_lapse_tx <- paste0(lapse_tx_terms, " = 0")
  pooled_test_lapse_tx <- pool_chi_square_tests(models, hypothesis_lapse_tx, "Lapse × treatment interaction")
}
```

### Joint Hypothesis Test Results

```{r joint-test-results}
# Compile all test results
joint_tests <- data.frame(
  Test = character(),
  Chi_Square = numeric(),
  df = numeric(),
  P_Value = character(),
  Significant = character(),
  stringsAsFactors = FALSE
)

if (exists("pooled_test_3way")) {
  joint_tests <- rbind(joint_tests, data.frame(
    Test = "Lapse × DR × Treatment (3-way)",
    Chi_Square = round(pooled_test_3way$D_bar, 3),
    df = pooled_test_3way$df,
    P_Value = format.pval(pooled_test_3way$p_value, digits = 4, eps = 0.0001),
    Significant = ifelse(pooled_test_3way$p_value < 0.05, "Yes", "No")
  ))
}

if (exists("pooled_test_lapse_dr")) {
  joint_tests <- rbind(joint_tests, data.frame(
    Test = "Lapse × DR Severity",
    Chi_Square = round(pooled_test_lapse_dr$D_bar, 3),
    df = pooled_test_lapse_dr$df,
    P_Value = format.pval(pooled_test_lapse_dr$p_value, digits = 4, eps = 0.0001),
    Significant = ifelse(pooled_test_lapse_dr$p_value < 0.05, "Yes", "No")
  ))
}

if (exists("pooled_test_lapse_tx")) {
  joint_tests <- rbind(joint_tests, data.frame(
    Test = "Lapse × Treatment",
    Chi_Square = round(pooled_test_lapse_tx$D_bar, 3),
    df = pooled_test_lapse_tx$df,
    P_Value = format.pval(pooled_test_lapse_tx$p_value, digits = 4, eps = 0.0001),
    Significant = ifelse(pooled_test_lapse_tx$p_value < 0.05, "Yes", "No")
  ))
}

kable(joint_tests, caption = "Pooled joint hypothesis tests (Li-Raghunathan-Rubin method)")
```

## Calculate Effects Within Each Subgroup Using emmeans

```{r emmeans-analysis}
# Function to pool emmeans results across imputations (matching main effects approach)
pool_emmeans_subgroup_effects <- function(models, n_imputations) {

  # Set emmeans options to handle larger reference grids (matching main effects)
  emm_options(rg.limit = 130000)

  # Store results from each imputation
  all_emmeans <- list()
  all_contrasts <- list()
  all_risk_diffs <- list()

  for (i in 1:n_imputations) {

    # Get marginal means on both log odds and response scales
    emm_logit <- emmeans(models[[i]], 
                         ~ ever_lapse_binary | dr_severity * any_treatment,
                         weights = "proportional")

    emm_response <- emmeans(models[[i]], 
                           ~ ever_lapse_binary | dr_severity * any_treatment,
                           weights = "proportional",
                           type = "response")

    # Get contrasts on log odds scale (for odds ratios)
    lapse_contrasts <- contrast(emm_logit, method = "revpairwise", 
                               by = c("dr_severity", "any_treatment"))

    # Get pairwise comparisons on response scale (for risk differences)
    risk_diffs <- pairs(emm_response, reverse = TRUE,
                       by = c("dr_severity", "any_treatment"))

    all_emmeans[[i]] <- as.data.frame(emm_response)
    all_contrasts[[i]] <- as.data.frame(lapse_contrasts)
    all_risk_diffs[[i]] <- as.data.frame(risk_diffs)
  }

  # Get structure from first imputation
  first_contrast <- all_contrasts[[1]]
  first_rd <- all_risk_diffs[[1]]
  n_subgroups <- nrow(first_contrast)

  # Initialize pooled results data frames
  pooled_or_results <- data.frame(
    dr_severity = first_contrast$dr_severity,
    any_treatment = first_contrast$any_treatment,
    contrast = first_contrast$contrast,
    log_odds = numeric(n_subgroups),
    SE_log_odds = numeric(n_subgroups),
    df = numeric(n_subgroups),
    stringsAsFactors = FALSE
  )

  pooled_rd_results <- data.frame(
    dr_severity = first_rd$dr_severity,
    any_treatment = first_rd$any_treatment,
    contrast = first_rd$contrast,
    risk_diff = numeric(n_subgroups),
    SE_risk_diff = numeric(n_subgroups),
    df = numeric(n_subgroups),
    stringsAsFactors = FALSE
  )

  # Check column names in first imputation for debugging
  if (length(all_contrasts) > 0) {
    message("Contrast columns: ", paste(names(all_contrasts[[1]]), collapse = ", "))
  }
  if (length(all_risk_diffs) > 0) {
    message("Risk diff columns: ", paste(names(all_risk_diffs[[1]]), collapse = ", "))
  }

  # Pool each subgroup using Rubin's rules
  for (j in 1:n_subgroups) {
    # Pool odds ratio contrasts (log scale)
    log_odds_estimates <- sapply(all_contrasts, function(x) {
      if ("estimate" %in% names(x)) {
        x$estimate[j]
      } else {
        # Try other possible column names
        x[j, which(names(x) %in% c("estimate", "emmean", "contrast"))[1]]
      }
    })

    log_odds_SEs <- sapply(all_contrasts, function(x) {
      if ("SE" %in% names(x)) {
        x$SE[j]
      } else if ("std.error" %in% names(x)) {
        x$std.error[j]
      } else {
        x[j, which(names(x) %in% c("SE", "std.error", "se"))[1]]
      }
    })

    # Convert to numeric and check for validity
    log_odds_estimates <- as.numeric(log_odds_estimates)
    log_odds_SEs <- as.numeric(log_odds_SEs)

    # Apply Rubin's rules for log odds
    Q_bar_log <- mean(log_odds_estimates, na.rm = TRUE)
    U_bar_log <- mean(log_odds_SEs^2, na.rm = TRUE)
    B_log <- var(log_odds_estimates, na.rm = TRUE)
    T_log <- U_bar_log + (1 + 1/n_imputations) * B_log

    # Degrees of freedom (Barnard-Rubin)
    lambda_log <- (B_log + B_log/n_imputations) / T_log
    df_log <- (n_imputations - 1) / (lambda_log^2)

    pooled_or_results$log_odds[j] <- Q_bar_log
    pooled_or_results$SE_log_odds[j] <- sqrt(T_log)
    pooled_or_results$df[j] <- df_log

    # Pool risk differences (probability scale)
    # Check for column names - emmeans may use 'estimate' or other names
    rd_estimates <- sapply(all_risk_diffs, function(x) {
      if ("estimate" %in% names(x)) {
        x$estimate[j]
      } else if ("emmean" %in% names(x)) {
        x$emmean[j]
      } else if ("prob" %in% names(x)) {
        x$prob[j]
      } else {
        # Try to get the estimate column (might be named differently)
        x[j, which(names(x) %in% c("estimate", "emmean", "prob", "prediction"))[1]]
      }
    })

    rd_SEs <- sapply(all_risk_diffs, function(x) {
      if ("SE" %in% names(x)) {
        x$SE[j]
      } else if ("std.error" %in% names(x)) {
        x$std.error[j]
      } else {
        # Try to get the SE column
        x[j, which(names(x) %in% c("SE", "std.error", "se"))[1]]
      }
    })

    # Convert to numeric and check for validity
    rd_estimates <- as.numeric(rd_estimates)
    rd_SEs <- as.numeric(rd_SEs)

    # Check if we got valid numeric values
    if (any(is.na(rd_estimates)) || any(is.na(rd_SEs))) {
      warning(paste("Missing values in risk difference pooling for subgroup", j))
    }

    # Apply Rubin's rules for risk differences
    Q_bar_rd <- mean(rd_estimates, na.rm = TRUE)
    U_bar_rd <- mean(rd_SEs^2, na.rm = TRUE)
    B_rd <- var(rd_estimates, na.rm = TRUE)
    T_rd <- U_bar_rd + (1 + 1/n_imputations) * B_rd

    # Degrees of freedom
    lambda_rd <- (B_rd + B_rd/n_imputations) / T_rd
    df_rd <- (n_imputations - 1) / (lambda_rd^2)

    pooled_rd_results$risk_diff[j] <- Q_bar_rd
    pooled_rd_results$SE_risk_diff[j] <- sqrt(T_rd)
    pooled_rd_results$df[j] <- df_rd
  }

  return(list(
    odds_ratios = pooled_or_results,
    risk_differences = pooled_rd_results,
    emmeans_list = all_emmeans
  ))
}

# Check if emmeans results are cached
emmeans_cache_file <- file.path(reanalysis_data_dir, "emmeans_results_interaction.rds")

if (file.exists(emmeans_cache_file) && !force_recalc_emmeans) {
  # Load cached results
  pooled_results <- readRDS(emmeans_cache_file)
  emmeans_loaded_from_cache <- TRUE
  emmeans_cache_location <- emmeans_cache_file
  message("Emmeans results loaded from cache: ", emmeans_cache_location)
} else {
  # Calculate new emmeans results
  emmeans_loaded_from_cache <- FALSE

  # Pool contrasts across all imputations
  pooled_results <- pool_emmeans_subgroup_effects(models, n_imputations)

  # Save to cache
  saveRDS(pooled_results, emmeans_cache_file)
  emmeans_cache_location <- emmeans_cache_file
  message("Emmeans results calculated and saved to cache: ", emmeans_cache_location)
}

# Process odds ratio results
or_effects <- pooled_results$odds_ratios |>
  mutate(
    t_stat = log_odds / SE_log_odds,
    p.value = 2 * pt(-abs(t_stat), df = df),
    CI_Lower_log = log_odds - qt(0.975, df = df) * SE_log_odds,
    CI_Upper_log = log_odds + qt(0.975, df = df) * SE_log_odds,
    OR = exp(log_odds),
    OR_CI_Lower = exp(CI_Lower_log),
    OR_CI_Upper = exp(CI_Upper_log)
  )

# Process risk difference results
rd_effects <- pooled_results$risk_differences |>
  mutate(
    t_stat = risk_diff / SE_risk_diff,
    p.value = 2 * pt(-abs(t_stat), df = df),
    CI_Lower = risk_diff - qt(0.975, df = df) * SE_risk_diff,
    CI_Upper = risk_diff + qt(0.975, df = df) * SE_risk_diff
  )

# Create odds ratio table
or_table <- data.frame(
  DR_Severity = or_effects$dr_severity,
  Treatment = or_effects$any_treatment,
  OR = round(or_effects$OR, 3),
  `OR_95CI` = paste0("(", round(or_effects$OR_CI_Lower, 3), ", ", 
                     round(or_effects$OR_CI_Upper, 3), ")"),
  P_Value = format.pval(or_effects$p.value, digits = 4, eps = 0.0001)
)

# Create risk difference table
rd_table <- data.frame(
  DR_Severity = rd_effects$dr_severity,
  Treatment = rd_effects$any_treatment,
  Risk_Diff = round(rd_effects$risk_diff, 4),
  `RD_95CI` = paste0("(", round(rd_effects$CI_Lower, 4), ", ", 
                     round(rd_effects$CI_Upper, 4), ")"),
  P_Value = format.pval(rd_effects$p.value, digits = 4, eps = 0.0001)
)

kable(or_table,
      caption = "Odds Ratios for Lapsing Effect Within Each Subgroup",
      col.names = c("DR Severity", "Treatment", "Odds Ratio", "95% CI", "P-value"),
      row.names = FALSE)

kable(rd_table,
      caption = "Risk Differences for Lapsing Effect Within Each Subgroup",
      col.names = c("DR Severity", "Treatment", "Risk Diff", "95% CI", "P-value"),
      row.names = FALSE)

# Store for visualization (use OR results for compatibility)
subgroup_effects_stored <- data.frame(
  DR_Severity = or_effects$dr_severity,
  Treatment = or_effects$any_treatment,
  Log_Odds = or_effects$log_odds,
  SE = or_effects$SE_log_odds,
  CI_Lower = or_effects$CI_Lower_log,
  CI_Upper = or_effects$CI_Upper_log,
  OR = or_effects$OR,
  OR_CI_Lower = or_effects$OR_CI_Lower,
  OR_CI_Upper = or_effects$OR_CI_Upper,
  P_Value = or_effects$p.value
)
```

### Understanding the Different Effect Measures

::: {.callout-note}
## Why are the confidence intervals so different between measures?

The three effect measures presented above capture different aspects of the lapsing effect:

1. **Odds Ratios (OR)**: 
   - Multiplicative scale comparing odds of outcome
   - Wider CIs because they incorporate uncertainty from multiple interaction terms
   - Non-collapsible measure that depends on covariate distribution

2. **Risk Differences (RD)**:
   - Additive scale showing absolute change in probability
   - Narrower CIs because probabilities are bounded [0,1]
   - More intuitive interpretation for clinical decision-making

3. **Average Marginal Effects (AME)** (shown later):
   - Average of individual-level effects across actual patients
   - Smallest CIs because averaging reduces variance
   - Accounts for actual covariate distribution in each subgroup

The OR confidence intervals are wider because:
- They must account for uncertainty in the main effect plus all relevant interaction terms
- The log scale amplifies uncertainty when exponentiating
- Small subgroups have fewer effective degrees of freedom
:::

## Manual Calculation of Subgroup Effects (Validation)

::: {.callout-note collapse="true"}
## Click to expand manual calculations

This section shows the manual calculation of subgroup effects for validation purposes.

```{r manual-effects}
# Manually calculate effects for each of the 6 subgroups
# This validates the emmeans results and provides a fallback

coef_vals <- pooled_results$Estimate
names(coef_vals) <- pooled_results$Term
vcov_pooled <- diag(pooled_results$SE^2)
rownames(vcov_pooled) <- colnames(vcov_pooled) <- pooled_results$Term

# Initialize results
manual_effects <- expand.grid(
  DR_Severity = c("No_DR", "NPDR", "PDR"),
  Treatment = c("No_Treatment", "Any_Treatment"),
  stringsAsFactors = FALSE
)
manual_effects$Log_Odds <- NA
manual_effects$SE <- NA

# Calculate effect for each combination
# Base effect: ever_lapse_binaryLapsed

# No_DR, No_Treatment (reference group)
manual_effects[1, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"]
manual_effects[1, "SE"] <- sqrt(vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"])

# NPDR, No_Treatment
term_npdr <- "ever_lapse_binaryLapsed:dr_severityNPDR"
if (term_npdr %in% names(coef_vals)) {
  manual_effects[2, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"] + coef_vals[term_npdr]
  # Note: For simplicity, ignoring covariance in this demonstration
  manual_effects[2, "SE"] <- sqrt(vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"] + 
                                  vcov_pooled[term_npdr, term_npdr])
}

# PDR, No_Treatment
term_pdr <- "ever_lapse_binaryLapsed:dr_severityPDR"
if (term_pdr %in% names(coef_vals)) {
  manual_effects[3, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"] + coef_vals[term_pdr]
  manual_effects[3, "SE"] <- sqrt(vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"] + 
                                  vcov_pooled[term_pdr, term_pdr])
}

# No_DR, Any_Treatment
term_tx <- "ever_lapse_binaryLapsed:any_treatmentAny_Treatment"
if (term_tx %in% names(coef_vals)) {
  manual_effects[4, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"] + coef_vals[term_tx]
  manual_effects[4, "SE"] <- sqrt(vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"] + 
                                  vcov_pooled[term_tx, term_tx])
}

# NPDR, Any_Treatment
term_npdr_tx <- "ever_lapse_binaryLapsed:dr_severityNPDR:any_treatmentAny_Treatment"
if (all(c(term_npdr, term_tx, term_npdr_tx) %in% names(coef_vals))) {
  manual_effects[5, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"] + 
                                   coef_vals[term_npdr] + 
                                   coef_vals[term_tx] + 
                                   coef_vals[term_npdr_tx]
  # Simplified SE (ignoring covariances for demonstration)
  manual_effects[5, "SE"] <- sqrt(sum(c(
    vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"],
    vcov_pooled[term_npdr, term_npdr],
    vcov_pooled[term_tx, term_tx],
    vcov_pooled[term_npdr_tx, term_npdr_tx]
  )))
}

# PDR, Any_Treatment
term_pdr_tx <- "ever_lapse_binaryLapsed:dr_severityPDR:any_treatmentAny_Treatment"
if (all(c(term_pdr, term_tx, term_pdr_tx) %in% names(coef_vals))) {
  manual_effects[6, "Log_Odds"] <- coef_vals["ever_lapse_binaryLapsed"] + 
                                   coef_vals[term_pdr] + 
                                   coef_vals[term_tx] + 
                                   coef_vals[term_pdr_tx]
  manual_effects[6, "SE"] <- sqrt(sum(c(
    vcov_pooled["ever_lapse_binaryLapsed", "ever_lapse_binaryLapsed"],
    vcov_pooled[term_pdr, term_pdr],
    vcov_pooled[term_tx, term_tx],
    vcov_pooled[term_pdr_tx, term_pdr_tx]
  )))
}

# Calculate derived statistics for non-NA values
manual_effects <- manual_effects |>
  mutate(
    CI_Lower = ifelse(!is.na(Log_Odds) & !is.na(SE), Log_Odds - 1.96 * SE, NA),
    CI_Upper = ifelse(!is.na(Log_Odds) & !is.na(SE), Log_Odds + 1.96 * SE, NA),
    OR = ifelse(!is.na(Log_Odds), exp(Log_Odds), NA),
    P_Value = ifelse(!is.na(Log_Odds) & !is.na(SE) & SE > 0, 
                    2 * pnorm(-abs(Log_Odds / SE)), NA)
  )

# Format the results
manual_effects <- manual_effects |>
  mutate(across(c(Log_Odds, SE, CI_Lower, CI_Upper, OR), ~round(., 4))) |>
  mutate(P_Value = ifelse(!is.na(P_Value), 
                          format.pval(P_Value, digits = 4, eps = 0.0001), 
                          NA))

# Check if we have valid results
n_valid <- sum(!is.na(manual_effects$Log_Odds))
if (n_valid == 0) {
  message("Warning: No valid manual effect calculations")
} else if (n_valid < nrow(manual_effects)) {
  message("Warning: Some manual effect calculations contain NA values (", 
          nrow(manual_effects) - n_valid, " out of ", nrow(manual_effects), ")")
}

kable(manual_effects,
      caption = "Manual Calculation of Subgroup Effects (Validation)",
      row.names = FALSE)
```

:::

## Visualization of Three-Way Interaction

```{r visualization}
#| fig-height: 8
#| fig-width: 12

# Use the subgroup effects for visualization
plot_data <- subgroup_effects_stored

# Add OR confidence intervals (converting from log odds scale)
plot_data <- plot_data |>
  mutate(
    OR_CI_Lower = exp(CI_Lower),
    OR_CI_Upper = exp(CI_Upper)
  )

# Check for extreme values and cap if necessary
max_or <- max(c(plot_data$OR, plot_data$OR_CI_Upper), na.rm = TRUE)
if (max_or > 100) {
  warning(paste("Extreme OR values detected (max =", round(max_or, 1), "). Consider checking the model."))
  # Cap at reasonable values for visualization
  plot_data <- plot_data |>
    mutate(
      OR_display = pmin(OR, 20),
      OR_CI_Lower_display = pmax(OR_CI_Lower, 0.05),
      OR_CI_Upper_display = pmin(OR_CI_Upper, 20),
      capped = OR > 20 | OR_CI_Upper > 20
    )
} else {
  plot_data <- plot_data |>
    mutate(
      OR_display = OR,
      OR_CI_Lower_display = OR_CI_Lower,
      OR_CI_Upper_display = OR_CI_Upper,
      capped = FALSE
    )
}

# Create interaction plot with Odds Ratios (converted from log odds)
p1 <- ggplot(plot_data, aes(x = DR_Severity, y = OR_display, 
                            color = Treatment, group = Treatment)) +
  geom_point(size = 3) +
  geom_line(size = 1) +
  geom_errorbar(aes(ymin = OR_CI_Lower_display, ymax = OR_CI_Upper_display), 
                width = 0.1, size = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", alpha = 0.5) +
  labs(title = "Three-Way Interaction: Lapse × DR Severity × Treatment",
       subtitle = ifelse(any(plot_data$capped), 
                        "Effect of lapsing (OR scale, capped at 20 for display)",
                        "Effect of lapsing on vision impairment (odds ratio scale)"),
       x = "DR Severity",
       y = "Odds Ratio",
       color = "Treatment Status") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        legend.position = "bottom") +
  scale_color_manual(values = c("No_Treatment" = "#2ca02c", 
                               "Any_Treatment" = "#d62728")) +
  coord_cartesian(ylim = c(0, max(plot_data$OR_CI_Upper_display) * 1.1))

# Create OR bar plot with error bars
p2 <- ggplot(plot_data, aes(x = DR_Severity, y = OR_display, 
                            fill = Treatment)) +
  geom_col(position = position_dodge(width = 0.8), alpha = 0.7) +
  geom_errorbar(aes(ymin = OR_CI_Lower_display, ymax = OR_CI_Upper_display),
                position = position_dodge(width = 0.8),
                width = 0.2, size = 0.8) +
  geom_hline(yintercept = 1, linetype = "dashed", alpha = 0.5) +
  labs(title = "Odds Ratios by Subgroup",
       subtitle = ifelse(any(plot_data$capped),
                        "Effect of lapsing (with 95% CI, capped at 20 for display)",
                        "Effect of lapsing on vision impairment (with 95% CI)"),
       x = "DR Severity",
       y = "Odds Ratio",
       fill = "Treatment Status") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"),
        axis.text = element_text(size = 11),
        axis.title = element_text(size = 12),
        legend.position = "bottom") +
  scale_fill_manual(values = c("No_Treatment" = "#2ca02c", 
                               "Any_Treatment" = "#d62728")) +
  geom_text(aes(label = ifelse(capped, 
                               sprintf(">%.0f", OR_display),
                               sprintf("%.2f", OR))),
            position = position_dodge(width = 0.8),
            vjust = -1.5, size = 3.5) +
  coord_cartesian(ylim = c(0, max(plot_data$OR_CI_Upper_display) * 1.1))

# Arrange plots
grid.arrange(p1, p2, ncol = 2)
```

## Convergence Diagnostics

```{r convergence-diagnostics}
# Extract estimates across imputations for key subgroups
convergence_data <- data.frame()

for (i in 1:n_imputations) {
  # Get log odds estimates from each imputation for each subgroup
  contrast_df <- pooled_results$odds_ratios

  # Track No_DR + No_Treatment as reference
  ref_estimates <- sapply(1:n_imputations, function(imp) {
    df <- as.data.frame(pooled_results$emmeans_list[[imp]])
    subset_df <- df[df$dr_severity == "No_DR" & 
                   df$any_treatment == "No_Treatment" & 
                   df$ever_lapse_binary == "Lapsed", ]
    if (nrow(subset_df) > 0) {
      return(qlogis(subset_df$prob[1]))  # Convert prob to log odds
    } else {
      return(NA)
    }
  })

  # Track PDR + Any_Treatment as most complex interaction
  complex_estimates <- sapply(1:n_imputations, function(imp) {
    df <- as.data.frame(pooled_results$emmeans_list[[imp]])
    subset_df <- df[df$dr_severity == "PDR" & 
                   df$any_treatment == "Any_Treatment" & 
                   df$ever_lapse_binary == "Lapsed", ]
    if (nrow(subset_df) > 0) {
      return(qlogis(subset_df$prob[1]))
    } else {
      return(NA)
    }
  })
}

# Calculate running averages
running_avg_ref <- cumsum(ref_estimates) / seq_along(ref_estimates)
running_avg_complex <- cumsum(complex_estimates) / seq_along(complex_estimates)

# Create convergence plot
convergence_df <- data.frame(
  Imputation = rep(1:n_imputations, 2),
  Estimate = c(ref_estimates, complex_estimates),
  Running_Average = c(running_avg_ref, running_avg_complex),
  Subgroup = factor(rep(c("No_DR + No_Treatment", "PDR + Any_Treatment"), 
                        each = n_imputations))
)

# Plot convergence
library(ggplot2)
convergence_plot <- ggplot(convergence_df, aes(x = Imputation)) +
  geom_point(aes(y = Estimate, color = Subgroup), alpha = 0.5) +
  geom_line(aes(y = Running_Average, color = Subgroup), size = 1.2) +
  labs(title = "Convergence of Subgroup Effect Estimates",
       subtitle = "Points show individual estimates, lines show running averages",
       x = "Imputation Number",
       y = "Log Odds of Outcome (Lapsed)",
       color = "Subgroup") +
  theme_minimal() +
  theme(legend.position = "bottom")

print(convergence_plot)

# Summary statistics
convergence_stats <- data.frame(
  Subgroup = c("No_DR + No_Treatment", "PDR + Any_Treatment"),
  Mean = c(mean(ref_estimates), mean(complex_estimates)),
  SD = c(sd(ref_estimates), sd(complex_estimates)),
  CV = c(sd(ref_estimates)/abs(mean(ref_estimates)) * 100,
         sd(complex_estimates)/abs(mean(complex_estimates)) * 100),
  Final_Running_Avg = c(tail(running_avg_ref, 1), tail(running_avg_complex, 1))
) |>
  mutate(across(where(is.numeric), ~round(., 4)))

kable(convergence_stats,
      caption = "Convergence Statistics for Key Subgroups",
      col.names = c("Subgroup", "Mean", "SD", "CV (%)", "Final Running Avg"))
```

## Export Results

```{r export}
# Note: Using emmeans-based risk differences as the primary effect measure
# These are calculated at the reference grid rather than averaged over observed data

# Save key results for reporting
results_to_save <- list(
  pooled_coefficients = pooled_results,
  subgroup_effects = subgroup_effects_stored,
  emmeans_results = pooled_results,  # The emmeans-based effects
  probability_table = if(exists("prob_table")) prob_table else NULL,
  model_summaries = model_summaries,
  n_imputations = n_imputations,
  analysis_mode = analysis_mode,
  timestamp = Sys.time()
)

saveRDS(results_to_save, 
        file.path(reanalysis_data_dir, "outcome_analysis_results.rds"))

message("Results saved to: ", 
        file.path(reanalysis_data_dir, "outcome_analysis_results.rds"))
```