### 2025-09-01 reanalysis-7-missingness_patterns.qmd
fix ai idiocy

---
title: "Reanalysis Step 7: Missingness Pattern Analysis"
subtitle: "Comparing characteristics of excluded vs included patients"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
# Setup analysis environment
source("dependencies.R")
source("functions.R")
setup_analysis(seed = 2025)

# Additional packages for this analysis
library(knitr)
library(tableone)
library(ggplot2)
library(gridExtra)

# Initialize logging
log_file <- init_log("missingness_pattern_analysis")

# Load the original data before exclusions
# Use the load_and_prepare_data function to get processed variables
original_data <- load_and_prepare_data(reanalysis_data_file_path)

# Load imputed datasets (this is a list of 20 imputed dataframes)
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))

log_message("Data loaded successfully")
```

## Motivation

When we performed multiple imputation, the observed treatment effect increased. This suggests that those with missing data (excluded from complete case analysis) may differ systematically from those included. Specifically, those with missing outcomes may be more likely to lapse and have worse health and behavioral characteristics.

This analysis compares baseline characteristics between:
- **Excluded**: Patients with missing outcome data (excluded from complete case analysis)
- **Included**: Patients with complete outcome data (included in analysis)

## Identify Excluded vs Included Patients

```{r identify-groups}
# Use cohort_id if it exists to identify excluded vs included
if ("cohort_id" %in% names(original_data)) {
  # cohort_id has values "excluded" and "submitted"
  # "excluded" = patients with missing outcomes who were excluded from complete case analysis
  # "submitted" = patients with complete outcomes who were included
  original_data$excluded <- original_data$cohort_id == "excluded"
  log_message("Using cohort_id to identify excluded patients")
  log_message(paste("Cohort distribution:", 
                   sum(original_data$cohort_id == "excluded"), "excluded,",
                   sum(original_data$cohort_id == "submitted"), "submitted"))
} else if ("outcome_va_vi_binary" %in% names(original_data)) {
  # Use missing outcome as the exclusion criterion
  original_data$excluded <- is.na(original_data$outcome_va_vi_binary)
  log_message("Using missing outcome_va_vi_binary to identify excluded patients")
} else if ("outcome_va_logMAR" %in% names(original_data)) {
  # Use missing outcome VA
  original_data$excluded <- is.na(original_data$outcome_va_logMAR)
  log_message("Using missing outcome_va_logMAR to identify excluded patients")
} else {
  stop("Cannot identify excluded patients - no cohort_id or outcome variables found")
}

# Verify the excluded variable has two levels
if (length(unique(original_data$excluded)) < 2) {
  log_message(paste("Warning: All patients are", 
                   ifelse(all(original_data$excluded), "excluded", "included")))
  log_message("Check data processing - there should be both excluded and included patients")
}

# Summary counts
n_total <- nrow(original_data)
n_included <- sum(!original_data$excluded)  # FALSE = included
n_excluded <- sum(original_data$excluded)   # TRUE = excluded
pct_excluded <- round(100 * n_excluded / n_total, 1)

log_message(sprintf("Total patients (eyes): %d", n_total))
log_message(sprintf("Included (submitted/complete outcome): %d (%.1f%%)", n_included, 100 - pct_excluded))
log_message(sprintf("Excluded (missing outcome): %d (%.1f%%)", n_excluded, pct_excluded))

# Create summary table
summary_df <- data.frame(
  Group = c("Included (submitted)", "Excluded", "Total"),
  N = c(n_included, n_excluded, n_total),
  Percentage = c(round(100 - pct_excluded, 1), pct_excluded, 100)
)
```

```{r summary-table}
#| echo: false
kable(summary_df,
      caption = "Sample Size Summary",
      col.names = c("Group", "N (eyes)", "Percentage (%)"),
      align = c("l", "r", "r"))
```

## Variable Selection

```{r variable-selection}
# Get variables used in outcome and propensity score models
# Log what columns exist in the processed data
log_message(paste("Total columns in processed data:", length(names(original_data))))
log_message(paste("Sample columns:", paste(names(original_data)[1:min(10, length(names(original_data)))], collapse = ", ")))

# Variables from the actual matching and outcome formulas in functions.R
# Demographics created by load_and_prepare_data
demographic_vars <- c("age", "age_cat", "gender_cat", "race_ethnic_cat", "insurance_cat")

# Comorbidity scores (these ARE in the raw data)
comorbidity_vars <- c("CCI", "DCSI")

# Vision and treatment variables
vision_vars <- c("baseline_VA_logMAR")

# Treatment flags (these ARE in the raw data)
treatment_vars <- c("other_inject", "anti_VEGF", "focal_laser_flag", "PRP_flag")

# Other eye conditions
eye_condition_vars <- c("glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat", 
                        "catsurg_before_hitplus_cat")

# DR severity (created by the function)
dr_vars <- c("dr_severity")

# Treatment/exposure (created by the function)
exposure_var <- "ever_lapse_binary"

# Outcome variables (created by the function)
outcome_vars <- c("outcome_va_logMAR", "outcome_va_vi_binary")

# Combine all variables
all_vars <- c(demographic_vars, comorbidity_vars, vision_vars, treatment_vars,
              eye_condition_vars, dr_vars, exposure_var, outcome_vars)

# Check which variables exist in the data
available_vars <- intersect(all_vars, names(original_data))
missing_vars <- setdiff(all_vars, names(original_data))

if (length(missing_vars) > 0) {
  log_message(paste("Variables not found in data:", paste(missing_vars, collapse = ", ")))
}

log_message(paste("Analyzing", length(available_vars), "variables"))
```

## Descriptive Statistics by Exclusion Status

```{r descriptive-stats}
# Create Table 1 comparing excluded vs included
# Specify which variables should be treated as factors
factor_vars <- c("age_cat", "gender_cat", "race_ethnic_cat", "insurance_cat",
                 "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat", 
                 "catsurg_before_hitplus_cat", "dr_severity", "ever_lapse_binary")

# Ensure factor variables are factors
for (var in factor_vars) {
  if (var %in% names(original_data)) {
    if (!is.factor(original_data[[var]])) {
      original_data[[var]] <- as.factor(original_data[[var]])
    }
  }
}

# Create Table 1
table1 <- CreateTableOne(
  vars = available_vars,
  strata = "excluded",
  data = original_data,
  factorVars = intersect(factor_vars, available_vars),
  includeNA = TRUE,
  test = TRUE,
  smd = TRUE
)

# Print table with standardized mean differences
# Label the strata clearly
table1_print <- print(table1, 
                      smd = TRUE, 
                      showAllLevels = TRUE,
                      formatOptions = list(big.mark = ","),
                      printToggle = FALSE)

# Update column names for clarity
colnames(table1_print)[colnames(table1_print) == "FALSE"] <- "Included (n)"
colnames(table1_print)[colnames(table1_print) == "TRUE"] <- "Excluded (n)"

kable(table1_print, 
      caption = "Comparison of Baseline Characteristics: Excluded vs Included Patients")
```

## Key Differences Visualization

```{r key-differences}
#| fig-height: 10
#| fig-width: 12

# Extract SMDs with more detail for factor variables
# Get the table itself to access all SMD information
table1_matrix <- print(table1, smd = TRUE, printToggle = FALSE)

# Extract SMDs manually from the table for each row
# This will give us SMDs for each level of factor variables
smd_data <- data.frame()

# Process each variable in the table
for (var in available_vars) {
  if (var %in% names(original_data)) {
    if (var %in% factor_vars && is.factor(original_data[[var]])) {
      # For factor variables, calculate SMD for each level
      var_levels <- levels(original_data[[var]])
      for (level in var_levels) {
        # Calculate proportion in each group
        prop_excluded <- mean(original_data[[var]][original_data$excluded] == level, na.rm = TRUE)
        prop_included <- mean(original_data[[var]][!original_data$excluded] == level, na.rm = TRUE)

        # Calculate SMD for this level
        # SMD for proportions = (p1 - p2) / sqrt((p1*(1-p1) + p2*(1-p2))/2)
        pooled_var <- (prop_excluded * (1 - prop_excluded) + prop_included * (1 - prop_included)) / 2
        smd_val <- if (pooled_var > 0) {
          abs(prop_excluded - prop_included) / sqrt(pooled_var)
        } else {
          0
        }

        smd_data <- rbind(smd_data, data.frame(
          Variable = paste0(var, ": ", level),
          SMD = smd_val,
          VarType = "Factor",
          BaseVar = var,
          stringsAsFactors = FALSE
        ))
      }
    } else {
      # For continuous variables, calculate single SMD
      if (sum(!is.na(original_data[[var]])) > 0) {
        mean_excluded <- mean(original_data[[var]][original_data$excluded], na.rm = TRUE)
        mean_included <- mean(original_data[[var]][!original_data$excluded], na.rm = TRUE)
        sd_pooled <- sqrt((var(original_data[[var]][original_data$excluded], na.rm = TRUE) + 
                          var(original_data[[var]][!original_data$excluded], na.rm = TRUE)) / 2)

        smd_val <- if (!is.na(sd_pooled) && sd_pooled > 0) {
          abs(mean_excluded - mean_included) / sd_pooled
        } else {
          0
        }

        smd_data <- rbind(smd_data, data.frame(
          Variable = var,
          SMD = smd_val,
          VarType = "Continuous",
          BaseVar = var,
          stringsAsFactors = FALSE
        ))
      }
    }
  }
}

# Remove NA values and order by absolute SMD
smd_data <- smd_data[!is.na(smd_data$SMD), ]
smd_data <- smd_data[order(abs(smd_data$SMD), decreasing = TRUE), ]

# Get top differences (for display purposes, limit to top 20)
smd_data_top <- head(smd_data, 20)

# Identify variables with substantial differences (SMD > 0.1)
substantial_diff <- smd_data[abs(smd_data$SMD) > 0.1, ]

# Plot 1: SMD plot (showing top 20 differences)
p1 <- ggplot(smd_data_top, aes(x = reorder(Variable, abs(SMD)), y = SMD)) +
  geom_point(aes(color = VarType), size = 3) +
  geom_point(data = smd_data_top[abs(smd_data_top$SMD) > 0.1, ],
             color = "red", size = 3.5) +
  geom_hline(yintercept = c(-0.1, 0, 0.1), linetype = c(2, 1, 2), alpha = 0.5) +
  coord_flip() +
  labs(title = "Standardized Mean Differences: Excluded vs Included (Top 20)",
       subtitle = "Red overlay indicates SMD > 0.1; Factor levels shown separately",
       x = "Variable",
       y = "Standardized Mean Difference",
       color = "Variable Type") +
  theme_minimal() +
  theme(plot.title = element_text(size = 14, face = "bold"))

print(p1)
```

## Focus on Key Variables

```{r key-variables}
# Analyze key variables of interest
key_results <- data.frame(
  Variable = character(),
  Included = character(),
  Excluded = character(),
  P_value = numeric(),
  SMD = numeric(),
  stringsAsFactors = FALSE
)

# Helper function to format results
format_continuous <- function(data, var, excluded_flag) {
  subset_data <- data[data$excluded == excluded_flag & !is.na(data[[var]]), var]
  sprintf("%.2f (%.2f)", mean(subset_data, na.rm = TRUE), sd(subset_data, na.rm = TRUE))
}

format_binary <- function(data, var, excluded_flag, level = "1") {
  subset_data <- data[data$excluded == excluded_flag, ]
  n <- sum(!is.na(subset_data[[var]]))
  n_level <- sum(subset_data[[var]] == level, na.rm = TRUE)
  sprintf("%d (%.1f%%)", n_level, 100 * n_level / n)
}

# Analyze lapsing rate
if ("ever_lapse_binary" %in% available_vars) {
  lapse_included <- mean(original_data$ever_lapse_binary[!original_data$excluded] == 1, na.rm = TRUE)
  lapse_excluded <- mean(original_data$ever_lapse_binary[original_data$excluded] == 1, na.rm = TRUE)
  lapse_p <- chisq.test(table(original_data$excluded, original_data$ever_lapse_binary))$p.value

  key_results <- rbind(key_results, data.frame(
    Variable = "Lapsed from care",
    Included = sprintf("%.1f%%", 100 * lapse_included),
    Excluded = sprintf("%.1f%%", 100 * lapse_excluded),
    P_value = lapse_p,
    SMD = abs(lapse_excluded - lapse_included) / sqrt((lapse_included * (1 - lapse_included) + 
                                                        lapse_excluded * (1 - lapse_excluded)) / 2)
  ))
}

# Analyze baseline VA
if ("baseline_VA_logMAR" %in% available_vars) {
  va_included <- mean(original_data$baseline_VA_logMAR[!original_data$excluded], na.rm = TRUE)
  va_excluded <- mean(original_data$baseline_VA_logMAR[original_data$excluded], na.rm = TRUE)
  va_p <- t.test(original_data$baseline_VA_logMAR ~ original_data$excluded)$p.value

  key_results <- rbind(key_results, data.frame(
    Variable = "Baseline VA (logMAR)",
    Included = sprintf("%.3f", va_included),
    Excluded = sprintf("%.3f", va_excluded),
    P_value = va_p,
    SMD = abs(va_excluded - va_included) / sd(original_data$baseline_VA_logMAR, na.rm = TRUE)
  ))
}

# Analyze CCI
if ("CCI" %in% available_vars) {
  cci_included <- mean(original_data$CCI[!original_data$excluded], na.rm = TRUE)
  cci_excluded <- mean(original_data$CCI[original_data$excluded], na.rm = TRUE)
  cci_p <- t.test(original_data$CCI ~ original_data$excluded)$p.value

  key_results <- rbind(key_results, data.frame(
    Variable = "Charlson Comorbidity Index",
    Included = sprintf("%.2f", cci_included),
    Excluded = sprintf("%.2f", cci_excluded),
    P_value = cci_p,
    SMD = abs(cci_excluded - cci_included) / sd(original_data$CCI, na.rm = TRUE)
  ))
}

# Format and display
key_results$P_value <- format.pval(key_results$P_value, digits = 3)
key_results$SMD <- round(key_results$SMD, 3)

kable(key_results,
      caption = "Key Variable Comparisons",
      col.names = c("Variable", "Included", "Excluded", "P-value", "SMD"))
```

## Missingness Patterns by DR Severity

```{r missingness-by-dr}
#| fig-height: 6
#| fig-width: 10

if ("dr_severity" %in% available_vars) {
  # Calculate missingness rate by DR severity
  miss_by_dr <- original_data %>%
    group_by(dr_severity) %>%
    summarise(
      n_total = n(),
      n_excluded = sum(excluded),
      pct_excluded = 100 * n_excluded / n_total,
      .groups = 'drop'
    )

  # Plot missingness by DR severity
  p2 <- ggplot(miss_by_dr, aes(x = dr_severity, y = pct_excluded)) +
    geom_col(fill = "steelblue", alpha = 0.7) +
    geom_text(aes(label = sprintf("%.1f%%", pct_excluded)), 
              vjust = -0.5, size = 4) +
    labs(title = "Outcome Missingness by DR Severity",
         subtitle = "Percentage of patients excluded due to missing outcomes",
         x = "DR Severity",
         y = "% Excluded") +
    theme_minimal() +
    theme(plot.title = element_text(size = 14, face = "bold"))

  print(p2)

  # Table of missingness by DR severity
  kable(miss_by_dr,
        caption = "Missingness Patterns by DR Severity",
        col.names = c("DR Severity", "Total N", "N Excluded", "% Excluded"),
        digits = 1)
}
```

## Missingness Patterns by Treatment and Lapse Status

```{r missingness-by-treatment}
#| fig-height: 8
#| fig-width: 12

# Create 2x2 plots for treatment and lapse patterns
plots <- list()

# Plot 1: Missingness by treatment status
if ("any_treatment" %in% available_vars) {
  miss_by_treatment <- original_data %>%
    group_by(any_treatment) %>%
    summarise(
      n_total = n(),
      n_excluded = sum(excluded),
      pct_excluded = 100 * n_excluded / n_total,
      .groups = 'drop'
    )

  plots[[1]] <- ggplot(miss_by_treatment, aes(x = any_treatment, y = pct_excluded)) +
    geom_col(fill = "coral", alpha = 0.7) +
    geom_text(aes(label = sprintf("%.1f%%", pct_excluded)), 
              vjust = -0.5, size = 4) +
    labs(title = "Missingness by Treatment Status",
         x = "Any Treatment", y = "% Excluded") +
    theme_minimal()
}

# Plot 2: Missingness by lapse status
if ("ever_lapse_binary" %in% available_vars) {
  miss_by_lapse <- original_data %>%
    mutate(lapse_status = factor(ever_lapse_binary, 
                                 levels = c(0, 1),
                                 labels = c("Not Lapsed", "Lapsed"))) %>%
    group_by(lapse_status) %>%
    summarise(
      n_total = n(),
      n_excluded = sum(excluded),
      pct_excluded = 100 * n_excluded / n_total,
      .groups = 'drop'
    )

  plots[[2]] <- ggplot(miss_by_lapse, aes(x = lapse_status, y = pct_excluded)) +
    geom_col(fill = "darkgreen", alpha = 0.7) +
    geom_text(aes(label = sprintf("%.1f%%", pct_excluded)), 
              vjust = -0.5, size = 4) +
    labs(title = "Missingness by Lapse Status",
         x = "Lapse Status", y = "% Excluded") +
    theme_minimal()
}

# Plot 3: Joint distribution
if ("any_treatment" %in% available_vars && "ever_lapse_binary" %in% available_vars) {
  miss_joint <- original_data %>%
    mutate(lapse_status = factor(ever_lapse_binary, 
                                 levels = c(0, 1),
                                 labels = c("Not Lapsed", "Lapsed"))) %>%
    group_by(any_treatment, lapse_status) %>%
    summarise(
      n_total = n(),
      n_excluded = sum(excluded),
      pct_excluded = 100 * n_excluded / n_total,
      .groups = 'drop'
    )

  plots[[3]] <- ggplot(miss_joint, aes(x = any_treatment, y = pct_excluded, 
                                       fill = lapse_status)) +
    geom_col(position = position_dodge(width = 0.8), alpha = 0.7) +
    labs(title = "Missingness by Treatment and Lapse Status",
         x = "Treatment Status", y = "% Excluded",
         fill = "Lapse Status") +
    theme_minimal() +
    scale_fill_manual(values = c("Not Lapsed" = "blue", "Lapsed" = "red"))
}

# Arrange plots
if (length(plots) > 0) {
  # Remove NULL elements from plots list
  plots <- plots[!sapply(plots, is.null)]

  if (length(plots) == 1) {
    print(plots[[1]])
  } else if (length(plots) > 1) {
    do.call(grid.arrange, c(plots, list(ncol = 2)))
  }
}
```

## Comparison Across Imputed Datasets

```{r imputation-comparison}
# For key variables, compare means across imputed datasets for originally missing vs complete
# This shows how imputation filled in the missing values

imputation_summary <- data.frame()

# Get IDs of included patients (those not excluded)
id_var <- if("cohort_id" %in% names(original_data)) "cohort_id" else "e_mrn_deidentified"
included_data <- original_data[!original_data$excluded, ]
included_ids <- unique(included_data[[id_var]])

for (var in c("baseline_VA_logMAR", "CCI", "DCSI")) {
  if (var %in% names(original_data)) {  # Check if variable exists in original data
    # Get which observations were originally missing
    originally_missing <- is.na(original_data[[var]])

    # Calculate mean across imputations for originally missing
    if (sum(originally_missing, na.rm = TRUE) > 0) {
      means_missing <- sapply(imputed_datasets, function(df) {
        # Match IDs to find corresponding rows in imputed data
        imputed_ids <- df[[id_var]]
        matching_rows <- imputed_ids %in% included_ids
        missing_and_included <- originally_missing[match(imputed_ids, original_data[[id_var]])] & matching_rows

        if (sum(missing_and_included, na.rm = TRUE) > 0) {
          mean(df[[var]][missing_and_included], na.rm = TRUE)
        } else {
          NA
        }
      })

      # Calculate mean for originally complete
      means_complete <- sapply(imputed_datasets, function(df) {
        imputed_ids <- df[[id_var]]
        matching_rows <- imputed_ids %in% included_ids
        complete_and_included <- !originally_missing[match(imputed_ids, original_data[[id_var]])] & matching_rows

        if (sum(complete_and_included, na.rm = TRUE) > 0) {
          mean(df[[var]][complete_and_included], na.rm = TRUE)
        } else {
          NA
        }
      })

      # Remove NAs before calculating statistics
      means_missing <- means_missing[!is.na(means_missing)]
      means_complete <- means_complete[!is.na(means_complete)]

      if (length(means_missing) > 0 && length(means_complete) > 0) {
        imputation_summary <- rbind(imputation_summary, data.frame(
          Variable = var,
          Originally_Complete = sprintf("%.3f (%.3f)", 
                                       mean(means_complete), sd(means_complete)),
          Imputed_for_Missing = sprintf("%.3f (%.3f)", 
                                        mean(means_missing), sd(means_missing)),
          Difference = sprintf("%.3f", mean(means_missing) - mean(means_complete))
        ))
      }
    }
  }
}

if (nrow(imputation_summary) > 0) {
  kable(imputation_summary,
        caption = "Comparison of Imputed Values: Originally Missing vs Complete",
        col.names = c("Variable", "Originally Complete", "Imputed (for Missing)", "Difference"))
}
```

## Summary and Implications

```{r summary}
# Summarize key findings
significant_differences <- substantial_diff$Variable

# Calculate effect size measures
lapse_or_excluded_vs_included <- NA
if ("ever_lapse_binary" %in% available_vars) {
  # Calculate odds ratio of being excluded for lapsed vs not lapsed
  lapse_table <- table(Lapse = original_data$ever_lapse_binary, 
                      Excluded = original_data$excluded)
  if (ncol(lapse_table) == 2 && nrow(lapse_table) == 2) {
    # OR = (lapsed & excluded * not lapsed & not excluded) / 
    #      (lapsed & not excluded * not lapsed & excluded)
    lapse_or_excluded_vs_included <- (lapse_table["1", "TRUE"] * lapse_table["0", "FALSE"]) / 
                                     (lapse_table["1", "FALSE"] * lapse_table["0", "TRUE"])
  }
}

log_message("Analysis completed successfully")
```

### Key Findings

**Variables with Substantial Differences (SMD > 0.1):**
```{r}
#| echo: false
if (length(significant_differences) > 0) {
  for (var in significant_differences) {
    cat("- ", var, "\n")
  }
} else {
  cat("No variables showed substantial differences (SMD > 0.1)\n")
}
```

```{r lapse-missingness-summary}
#| echo: false
# Create summary of lapse status and missingness relationship
if (!is.na(lapse_or_excluded_vs_included)) {
  lapse_summary <- data.frame(
    Measure = c("Odds ratio of exclusion (lapsed vs not lapsed)",
                "Interpretation"),
    Value = c(sprintf("%.2f", lapse_or_excluded_vs_included),
              ifelse(lapse_or_excluded_vs_included > 1.5,
                     "Lapsed patients were substantially more likely to have missing outcomes",
                     "No substantial difference in missingness by lapse status"))
  )
  kable(lapse_summary, 
        caption = "Lapse Status and Missingness Relationship",
        col.names = c("Measure", "Value"))
} else {
  cat("Lapse status and missingness relationship could not be calculated.\n")
}
```

### Implications for Analysis

1. **Selection Bias**: The differences between excluded and included patients suggest potential selection bias in complete case analysis.

2. **Direction of Bias**: If excluded patients are more likely to lapse AND have worse health characteristics, complete case analysis may underestimate the treatment effect.

3. **Multiple Imputation Benefit**: The increase in treatment effect after imputation aligns with recovering information from a sicker, more vulnerable population that was excluded.

4. **Generalizability**: Results from imputed data may better represent the full patient population, including those with incomplete follow-up.

```{r finalize-log}
#| include: false
# Finalize logging
log_message("Missingness pattern analysis completed successfully")
final_log <- finalize_log(success = TRUE)
```