### 2025-09-02 reanalysis-2-multiple_imputation.qmd
imputation fix

---
title: "Reanalysis Step 2: Multiple Imputation"
subtitle: "Imputing missing values using mice package"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
seed <- 2025
setup_analysis(seed = seed)

# Additional packages for imputation
library(mice)
library(VIM)
library(naniar)

# Initialize logging
log_file <- init_log("multiple_imputation")

# ANALYSIS MODE: "test" or "final"
# test = 2 imputations, 2 iterations (fast)
# final = 50 imputations, 20 iterations (slow but thorough)
analysis_mode <- "final"  # Change to "final" for production run
log_message(paste0("Analysis mode: ", analysis_mode))

# CACHING CONFIGURATION FOR MICE
# Set to TRUE to use cached MICE results (skip imputation, only run post-processing)
use_cached_mice <- FALSE
mice_cache_file <- file.path(reanalysis_data_dir, "mice_results.rds")
log_message(paste0("MICE caching: ", ifelse(use_cached_mice, "ENABLED", "DISABLED")))

# Set parameters based on mode
if (analysis_mode == "test") {
  n_imputations <- 2
  n_iterations <- 2
} else if (analysis_mode == "final") {
  n_imputations <- 50
  n_iterations <- 20
} else {
  stop("Invalid analysis_mode. Must be 'test' or 'final'")
}

log_message(paste0("Imputations: ", n_imputations, ", Iterations: ", n_iterations))
```

**Running in `r toupper(analysis_mode)` mode:**  
- **Imputations:** `r n_imputations`  
- **Iterations:** `r n_iterations`

## Load Prepared Data

```{r load-data}
# Load the prepared dataset from Step 1
df_for_imputation <- readRDS(file.path(reanalysis_data_dir, "df_for_imputation.rds"))
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))

ps_model_vars <- variable_lists$ps_model_vars
outcome_model_vars <- variable_lists$outcome_model_vars

log_message(paste0("Loaded ", nrow(df_for_imputation), " observations"))
log_message(paste0("Unique patients: ", n_distinct(df_for_imputation$e_mrn_deidentified)))
```

## Define Variables for Imputation

```{r define-imputation-vars}
# Variables to include in imputation model
# Include: outcome, exposure, all covariates source variables only
# Derived variables will be recreated post-imputation

# Core analysis variables - SOURCE ONLY
imputation_vars <- c(
  "e_mrn_deidentified",

  # Temporal VA variables
  "cohort_id",  # Data indicator only - not used as predictor
  "baseline_VA_logMAR",  # Baseline VA
  "intermediate_VA_logMAR",  # Intermediate follow-up VA (for excluded cohort)
  "outcome_VA_logMAR",  # 2-year outcome (true for submitted, missing for excluded)
  # NOT outcome_va_vi_binary - derived from logMAR >= 0.3

  # Exposure
  "ever_lapse_binary",

  # Demographics/Clinical
  "gender_cat",
  "race_ethnic_cat",
  "insurance_cat",
  "age",  # Continuous age (source)
  # NOT age_cat - derived from cut(age)
  "CCI",
  "DCSI",

  # Treatment indicators - ALL SOURCES
  "other_inject",
  "anti_VEGF",
  "focal_laser_flag",
  "PRP_flag",
  # NOT treatment_type or any_treatment - derived from these 4

  # Comorbidities (already recoded to 0/1)
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat",
  "catsurg_before_hitplus_cat",

  # DR severity - use person_dr (created from No_DR/NPDR/PDR)
  "person_dr"  # Numeric 0/1/2
  # NOT dr_severity or person_dr_severity - derived from person_dr
)

# Prepare data for MICE - NO person_dr_max, use eye-level person_dr
df_mice <- df_for_imputation |>
  mutate(
    # Create a unique row identifier to preserve exact row matching
    row_id = row_number(),
    # Create numeric cluster ID for 2l.pan (required for multilevel imputation)
    cluster_id = as.numeric(factor(e_mrn_deidentified))
  ) |>
  dplyr::select(row_id, cluster_id, dplyr::all_of(imputation_vars)) |>
  mutate(
    # Convert person_dr to ordered factor for polr imputation method
    # person_dr is eye-level DR severity (0=No_DR, 1=NPDR, 2=PDR)
    person_dr = factor(person_dr, levels = c(0, 1, 2), 
                      labels = c("No_DR", "NPDR", "PDR"),
                      ordered = TRUE),
    # Lock factor levels for categorical variables to ensure consistency
    cohort_id = factor(cohort_id),  # Ensure cohort_id is a factor
    gender_cat = factor(gender_cat),
    race_ethnic_cat = factor(race_ethnic_cat),
    insurance_cat = factor(insurance_cat),

    # Convert binary variables to factors for logreg method
    # These need to be factors with 2 levels for mice logreg to work
    ever_lapse_binary = factor(ever_lapse_binary, levels = c(0, 1)),
    catsurg_before_hitplus_cat = factor(catsurg_before_hitplus_cat, levels = c(0, 1)),
    glaucoma_bef_hitplus_cat = factor(glaucoma_bef_hitplus_cat, levels = c(0, 1)),
    otherretina_bef_hitplus_cat = factor(otherretina_bef_hitplus_cat, levels = c(0, 1)),
    other_inject = factor(other_inject, levels = c(0, 1)),
    anti_VEGF = factor(anti_VEGF, levels = c(0, 1)),
    focal_laser_flag = factor(focal_laser_flag, levels = c(0, 1)),
    PRP_flag = factor(PRP_flag, levels = c(0, 1))
  )

log_message(paste0("person_dr (eye-level): ", 
                  sum(!is.na(df_mice$person_dr)), " non-missing, ",
                  sum(is.na(df_mice$person_dr)), " missing"))
```

## Setup Imputation Model

```{r setup-mice}
# Initialize mice
# First, do a dry run to check the imputation setup
mice_dry <- mice(df_mice, maxit = 0, printFlag = FALSE)

# Imputation model summary
miss_vars <- data.frame(
  Variable = names(mice_dry$nmis),
  N_Missing = mice_dry$nmis,
  Pct_Missing = round(mice_dry$nmis/nrow(df_mice) * 100, 1)
) |>
  filter(N_Missing > 0) |>
  arrange(desc(N_Missing))
```

### Initial Imputation Model Summary

**Total variables in dataset:** `r ncol(df_mice)`  
**Variables to be imputed:** `r sum(mice_dry$nmis > 0)`  
**Total missing values:** `r sum(mice_dry$nmis)`

**Missing data by variable:**
```{r missing-vars-table}
kable(miss_vars)

method_table <- data.frame(
  Variable = names(mice_dry$method),
  Method = mice_dry$method,
  stringsAsFactors = FALSE
) |>
  filter(Method != "")
```

**Initial imputation methods by variable:**
```{r method-table}
kable(method_table)

# Customize imputation methods as needed
custom_methods <- mice_dry$method

# Temporal VA variables - use 2l.pan for clustered imputation
custom_methods["baseline_VA_logMAR"] <- "2l.pan"  # Clustered imputation
custom_methods["intermediate_VA_logMAR"] <- "2l.pan"  # Clustered imputation
custom_methods["outcome_VA_logMAR"] <- "2l.pan"  # Clustered imputation
custom_methods["cohort_id"] <- ""  # Never impute - data management indicator only

# Other continuous variables
custom_methods["CCI"] <- "pmm"  # Predictive mean matching for count data
custom_methods["DCSI"] <- "pmm"  # Predictive mean matching for count data
custom_methods["age"] <- "pmm"

# Use proportional odds model for ordinal person_dr (0/1/2)
# person_dr is the eye-level DR severity
custom_methods["person_dr"] <- "polr"  # Ordinal logistic regression

# Use logistic regression for binary variables
custom_methods["catsurg_before_hitplus_cat"] <- "logreg"

# Use polytomous regression for categorical variables with >2 levels
custom_methods["race_ethnic_cat"] <- "polyreg"
custom_methods["insurance_cat"] <- "polyreg"

# Don't impute these variables
# Method = "" means: don't impute (but still use as predictor EXCEPT for IDs)
custom_methods["row_id"] <- ""  # Row ID - never impute, never use as predictor; see below.
custom_methods["cluster_id"] <- ""  # Cluster ID for 2l methods - never impute
custom_methods["e_mrn_deidentified"] <- ""  # Patient ID - never impute, never use as predictor; see below.

# Variables that are complete (no missing data) - leave method empty
# mice will automatically skip variables with no missing data
# Only set methods for variables that actually have missing data to avoid type mismatches

# Note: Variables with complete data won't be imputed even if they have a method assigned
# The method specification ensures proper measurement level if imputation were needed

# Check predictor matrix (which variables predict which)
pred_matrix <- mice_dry$predictorMatrix

# Don't use IDs as predictors AND don't impute them
# EXCEPT: cluster_id is set to -2 as cluster indicator for 2l methods
pred_matrix[, "cluster_id"] <- -2  # Cluster indicator for 2l.pan
pred_matrix["cluster_id", ] <- 0  # Don't predict cluster_id itself
pred_matrix[, "e_mrn_deidentified"] <- 0  # Don't use patient ID as predictor
pred_matrix["e_mrn_deidentified", ] <- 0  # Don't predict patient ID
pred_matrix[, "row_id"] <- 0
pred_matrix["row_id", ] <- 0

# TEMPORAL STRUCTURE RULES
# 1. Baseline can predict intermediate and outcome
pred_matrix["intermediate_VA_logMAR", "baseline_VA_logMAR"] <- 1
pred_matrix["outcome_VA_logMAR", "baseline_VA_logMAR"] <- 1

# 2. Intermediate can predict outcome
pred_matrix["outcome_VA_logMAR", "intermediate_VA_logMAR"] <- 1

# 3. CRITICAL: Prevent reverse-time prediction
pred_matrix["intermediate_VA_logMAR", "outcome_VA_logMAR"] <- 0  # Outcome cannot predict intermediate
pred_matrix["baseline_VA_logMAR", "outcome_VA_logMAR"] <- 0  # Outcome cannot predict baseline
pred_matrix["baseline_VA_logMAR", "intermediate_VA_logMAR"] <- 0  # Intermediate cannot predict baseline

# Cohort_id should not be predicted by anything AND should not predict anything
pred_matrix["cohort_id", ] <- 0  # Not predicted
pred_matrix[, "cohort_id"] <- 0  # Not a predictor

# Don't use variables as their own predictors
diag(pred_matrix) <- 0

# Ensure exposure is used as predictor for outcomes
pred_matrix[, "ever_lapse_binary"] <- 1
pred_matrix["ever_lapse_binary", "ever_lapse_binary"] <- 0

# Handle person_dr (eye-level DR severity)
# person_dr should be used as predictor and can be predicted
pred_matrix[, "person_dr"] <- 1  # Use as predictor for most variables
pred_matrix["person_dr", "person_dr"] <- 0  # Not its own predictor
pred_matrix["baseline_VA_logMAR", "person_dr"] <- 0  # Don't use to predict baseline VA
pred_matrix["row_id", "person_dr"] <- 0  # Don't use for IDs
pred_matrix["e_mrn_deidentified", "person_dr"] <- 0  # Don't use for IDs
pred_matrix["cohort_id", "person_dr"] <- 0  # Don't use for cohort_id

# Save the predictor matrix to CSV for inspection
pred_matrix_df <- as.data.frame(pred_matrix)
pred_matrix_df <- cbind(Variable = rownames(pred_matrix_df), pred_matrix_df)
write.csv(pred_matrix_df, file.path(reanalysis_data_dir, "predictor_matrix.csv"), row.names = FALSE)
log_message(paste0("Saved predictor matrix to: ", file.path(reanalysis_data_dir, "predictor_matrix.csv")))

# Check for perfect collinearity between intermediate and outcome for submitted cohort
collinearity_check <- df_mice |>
  filter(cohort_id == "submitted") |>
  summarise(
    n = n(),
    n_with_both = sum(!is.na(intermediate_VA_logMAR) & !is.na(outcome_VA_logMAR)),
    cor_intermediate_outcome = cor(intermediate_VA_logMAR, outcome_VA_logMAR, use = "complete.obs"),
    are_identical = all(intermediate_VA_logMAR == outcome_VA_logMAR, na.rm = TRUE)
  )

log_message(paste0("Submitted cohort - Correlation between intermediate and outcome: ",
                  round(collinearity_check$cor_intermediate_outcome, 3)))
log_message(paste0("Submitted cohort - Are intermediate and outcome identical? ",
                  collinearity_check$are_identical))

# Log which predictors are set for outcome_VA_logMAR
outcome_predictors <- names(which(pred_matrix["outcome_VA_logMAR", ] == 1))
log_message(paste0("Predictors for outcome_VA_logMAR (", length(outcome_predictors), "): ",
                  paste(outcome_predictors, collapse = ", ")))

```

### Final Imputation Model Configuration

```{r final-model-config}

# Get actual missing data information
missing_info <- data.frame(
  Variable = names(mice_dry$nmis),
  N_Missing = mice_dry$nmis,
  stringsAsFactors = FALSE
)

# Summary of methods
final_methods <- data.frame(
  Variable = names(custom_methods),
  Method = custom_methods,
  Type = dplyr::case_when(
    custom_methods == "" ~ "Excluded (IDs - not used as predictor)",
    custom_methods == "pmm" ~ "Predictive mean matching",
    custom_methods == "polr" ~ "Proportional odds regression",
    custom_methods == "logreg" ~ "Logistic regression",
    custom_methods == "polyreg" ~ "Polytomous regression",
    TRUE ~ custom_methods
  ),
  stringsAsFactors = FALSE
) |>
  left_join(missing_info, by = "Variable") |>
  mutate(
    Will_Be_Imputed = N_Missing > 0 & Method != "",
    Role = case_when(
      Variable %in% c("row_id", "e_mrn_deidentified") ~ "ID (excluded)",
      N_Missing == 0 & Method != "" ~ "Complete predictor",
      N_Missing > 0 & Method != "" ~ "To be imputed",
      TRUE ~ "Other"
    )
  )

# Sort to show meaningful groupings
final_methods <- final_methods |>
  dplyr::arrange(Role == "ID (excluded)", N_Missing == 0, Variable)

# Variables that will ACTUALLY be imputed (have missing data AND a method)
vars_to_impute <- final_methods |>
  filter(Will_Be_Imputed) |>
  pull(Variable)
```

**Imputation methods and missing data by variable:**
```{r final-methods-table}
final_methods |>
  select(Variable, Type, N_Missing, Role) |>
  kable(col.names = c("Variable", "Method Type", "Missing Values", "Role in Imputation"))
```

**Variables that will ACTUALLY be imputed** (have missing data, `r length(vars_to_impute)` total):
```{r vars-to-impute}
if (length(vars_to_impute) > 0) {
  data.frame(
    Variable = vars_to_impute,
    Missing_Count = final_methods |> 
      filter(Variable %in% vars_to_impute) |> 
      pull(N_Missing)
  ) |>
  kable()
} else {
  no_imputation_msg <- "No variables require imputation (all data complete)"
  print(no_imputation_msg)
}

# Show complete variables with methods assigned
complete_with_methods <- final_methods |>
  filter(N_Missing == 0 & Method != "") |>
  select(Variable, Type)
```

**Complete variables with imputation methods assigned** (`r nrow(complete_with_methods)` total):
These variables have no missing data and won't be imputed, but have methods specified to ensure proper measurement level:
```{r complete-vars-table}
if (nrow(complete_with_methods) > 0) {
  kable(complete_with_methods, col.names = c("Variable", "Method Type"))
} else {
  no_complete_msg <- "No complete variables with methods assigned"
  print(no_complete_msg)
}

# Show predictor matrix summary
# Get the number of predictors for each variable
predictors_per_var <- rowSums(pred_matrix)

# Separate by variable type
id_vars_pred <- predictors_per_var[c("row_id", "e_mrn_deidentified")]
imputed_vars_pred <- predictors_per_var[vars_to_impute]
other_vars_pred <- predictors_per_var[!names(predictors_per_var) %in% c("row_id", "e_mrn_deidentified", vars_to_impute)]

# Calculate meaningful statistics
n_predictors_for_imputed <- if(length(imputed_vars_pred) > 0) unique(imputed_vars_pred)[1] else NA
predictors <- names(which(colSums(pred_matrix) > 0))

# Verify IDs are excluded
ids_excluded <- all(pred_matrix[, "row_id"] == 0) & all(pred_matrix[, "e_mrn_deidentified"] == 0)
```

### Predictor Matrix Summary

**Number of predictors used:**
- For variables being imputed: **`r if(!is.na(n_predictors_for_imputed)) n_predictors_for_imputed else 0`** predictors each
- For ID variables: **0** predictors (excluded from imputation)

**ID variables properly excluded from predictors:** `r if(ids_excluded) "Yes" else "No"`

**Variables used as predictors** (`r length(predictors)` total):
```{r predictor-list}
if ("row_id" %in% predictors || "e_mrn_deidentified" %in% predictors) {
  warning("WARNING: ID variables are being used as predictors! This should not happen.")
}
print(predictors)
```

### Visit Sequence

```{r visit-sequence}
# Define the order of imputation to respect temporal flow
# Impute earlier timepoints before later ones
visit_sequence <- c()

# First impute baseline if needed
if ("baseline_VA_logMAR" %in% vars_to_impute) {
  visit_sequence <- c(visit_sequence, "baseline_VA_logMAR")
}

# Then impute intermediate
if ("intermediate_VA_logMAR" %in% vars_to_impute) {
  visit_sequence <- c(visit_sequence, "intermediate_VA_logMAR")
}

# Then impute outcome
if ("outcome_VA_logMAR" %in% vars_to_impute) {
  visit_sequence <- c(visit_sequence, "outcome_VA_logMAR")
}

# Add other variables that need imputation
other_vars_to_impute <- vars_to_impute[!vars_to_impute %in% 
                                       c("baseline_VA_logMAR", 
                                         "intermediate_VA_logMAR", 
                                         "outcome_VA_logMAR")]
visit_sequence <- c(visit_sequence, other_vars_to_impute)

# Store for display
imputation_order <- paste(visit_sequence, collapse = ", ")
```

**Imputation order (respecting temporal structure):** `r imputation_order`

### Imputation Parameters

```{r imputation-params}
imp_params <- data.frame(
  Parameter = c("Mode", "Number of imputations (m)", "Max iterations (maxit)", "Random seed", "Visit sequence"),
  Value = c(analysis_mode, n_imputations, n_iterations, seed, "Temporal order enforced")
)
kable(imp_params)
```

## Run Multiple Imputation

```{r run-imputation}
# Run mice imputation
# Note: Parameters are set based on analysis_mode in setup
time_estimate <- if(analysis_mode == "test") "a few minutes" else "15-30 minutes or more"
```

**Starting multiple imputation...**

- Mode: `r analysis_mode`
- Imputations: `r n_imputations`
- Iterations: `r n_iterations`
- Estimated time: `r time_estimate`
- MICE caching: `r ifelse(use_cached_mice, "ENABLED", "DISABLED")`
- Cache file: `r if(use_cached_mice && file.exists(mice_cache_file)) "EXISTS (will load)" else "NOT FOUND (will run fresh)"`

```{r mice-run}
# Check for cached MICE results
if (use_cached_mice && file.exists(mice_cache_file)) {
  log_message("Loading cached MICE results...")
  log_message(paste0("Cache file: ", mice_cache_file))

  mice_results <- readRDS(mice_cache_file)

  # Verify cached results match expected parameters
  cached_m <- mice_results$m
  cached_maxit <- mice_results$iteration

  log_message(paste0("Loaded ", cached_m, " imputations from cache"))
  log_message(paste0("Original run used ", cached_maxit, " iterations"))

  if (cached_m != n_imputations) {
    warning(paste0("Cached imputations (", cached_m, ") differ from requested (", n_imputations, ")"))
  }

  imp_time <- 0  # No time spent on imputation
  log_message("Using cached MICE results - skipping imputation step")

} else {
  # Run fresh MICE imputation
  if (use_cached_mice && !file.exists(mice_cache_file)) {
    log_message("Cache file not found - running fresh imputation")
  } else {
    log_message("Cache disabled - running fresh imputation")
  }

  log_message("Starting MICE imputation...")
  imp_start <- Sys.time()

  mice_results <- mice(
    df_mice,
    m = n_imputations,
    method = custom_methods,
    predictorMatrix = pred_matrix,
    visitSequence = visit_sequence,  # Enforce temporal order
    maxit = n_iterations,  # Uses n_iterations from setup
    seed = 2025,
    printFlag = TRUE
  )

  imp_time <- round(as.numeric(difftime(Sys.time(), imp_start, units = "mins")), 2)
  log_message(paste0("MICE imputation completed in ", imp_time, " minutes"))

  # Save to cache for future use
  saveRDS(mice_results, mice_cache_file)
  log_message(paste0("Saved MICE results to cache: ", mice_cache_file))
}

# Log person_dr imputation status
if ("person_dr" %in% names(mice_results$imp)) {
  n_imputed_dr <- sum(mice_results$nmis["person_dr"])
  log_message(paste0("person_dr imputed for ", n_imputed_dr, " observations"))
}

# Check convergence
plot(mice_results)
```

## Create Complete Datasets and Derive Variables

```{r complete-datasets}
# Extract all imputed datasets
imputed_datasets <- complete(mice_results, action = "all")

# Apply transformation functions to recreate derived variables

imputed_datasets <- lapply(1:n_imputations, function(i) {
  imp_data <- imputed_datasets[[i]]

  # Convert person_dr from factor back to numeric if needed
  if (is.factor(imp_data$person_dr)) {
    # Convert factor levels (No_DR=1, NPDR=2, PDR=3) to numeric (0, 1, 2)
    imp_data$person_dr <- as.numeric(imp_data$person_dr) - 1
  }

  # Apply transformation functions to recreate all derived variables
  imp_data <- imp_data |>
    create_outcome_variables() |>  # Creates outcome_va_vi_binary from logMAR
    create_age_cat() |>  # Creates age_cat from continuous age  
    create_treatment_type() |>  # Creates treatment_type and any_treatment
    create_dr_severity_variables()  # Creates dr_severity factor from person_dr

  # Verify dr_severity was created properly
  if (!"dr_severity" %in% names(imp_data)) {
    stop("dr_severity not created in imputation ", i)
  }
  if (!is.factor(imp_data$dr_severity)) {
    stop("dr_severity is not a factor in imputation ", i)
  }

  return(imp_data)
})

# Add back any variables not included in imputation or transformation
# Get the full original dataset with row IDs
df_original_full <- df_for_imputation |>
  mutate(row_id = row_number())

# Variables to add back (those not in imputation model or created by transformations)
vars_in_imputed <- names(imputed_datasets[[1]])
# Remove row_id from comparison since it's our join key
vars_to_add <- setdiff(names(df_original_full), c(vars_in_imputed, "row_id"))
```

### Adding Back Non-Imputed Variables

**Variables to add back:** `r if(length(vars_to_add) > 0) length(vars_to_add) else 0` variables not in imputation model

```{r add-back-vars}
if(length(vars_to_add) > 0) {
  # IMPORTANT: Exclude No_DR, NPDR, PDR from vars_to_add to prevent overwriting person_dr
  # These would cause create_dr_severity_variables to recalculate person_dr from eye-level data
  original_vars_to_add <- vars_to_add
  vars_to_add <- setdiff(vars_to_add, c("No_DR", "NPDR", "PDR"))

  if (length(original_vars_to_add) != length(vars_to_add)) {
    log_message("Excluded No_DR/NPDR/PDR from variables to add back to preserve person_dr consistency")
  }

  # Add these variables back to each imputed dataset using row_id for exact matching
  imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
    # Store original row count
    orig_rows <- nrow(imp_data)

    # Store the consistent person_dr we just created
    person_dr_consistent <- imp_data$person_dr

    # Join using row_id for exact 1:1 matching
    imp_data <- imp_data |>
      left_join(
        df_original_full |> 
          dplyr::select(row_id, dplyr::all_of(vars_to_add)),
        by = "row_id"
      ) |>
      # Ensure person_dr remains consistent (in case it somehow got overwritten)
      mutate(person_dr = person_dr_consistent)

    # Verify no row duplication
    if (nrow(imp_data) != orig_rows) {
      stop("Join changed row count from ", orig_rows, " to ", nrow(imp_data), 
           ". This should not happen with row_id join.")
    }

    # Remove row_id as it's no longer needed
    imp_data <- imp_data |> select(-row_id)

    return(imp_data)
  })
}

# FINAL STEP: Enforce patient-level variable consistency across all datasets
# person_dr, CCI, and DCSI should be consistent within patients
imputed_datasets <- lapply(imputed_datasets, function(imp_data) {
  # Calculate consistent values per patient (using max for all)
  patient_level_vars <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      # For CCI and DCSI, take minimum (better value) per patient
      # Lower scores indicate better health
      CCI_consistent = ifelse(all(is.na(CCI)), 
                             NA_real_, 
                             min(CCI, na.rm = TRUE)),
      DCSI_consistent = ifelse(all(is.na(DCSI)), 
                              NA_real_, 
                              min(DCSI, na.rm = TRUE)),
      # Note: person_dr is now eye-level, so we don't enforce consistency
      .groups = 'drop'
    )

  # Replace with consistent values (CCI and DCSI only, not person_dr)
  imp_data <- imp_data |>
    left_join(patient_level_vars, by = "e_mrn_deidentified") |>
    mutate(
      # Only replace CCI and DCSI with consistent (minimum) values
      CCI = coalesce(CCI_consistent, CCI),
      DCSI = coalesce(DCSI_consistent, DCSI)
      # person_dr remains eye-level, not enforced to be consistent
    ) |>
    select(-CCI_consistent, -DCSI_consistent)

  return(imp_data)
})

log_message("Enforced patient-level consistency for CCI and DCSI by taking minimum (better) value per patient")

completeness_check <- lapply(imputed_datasets, function(d) {
  sum(!complete.cases(d[, ps_model_vars]))
})

```

### Comprehensive Completeness Check

```{r comprehensive-completeness}
# Define key variables that should have been imputed
key_imputed_vars <- c(
  # Visual acuity variables
  "baseline_VA_logMAR",
  "intermediate_VA_logMAR", 
  "outcome_VA_logMAR",

  # DR severity (derived from person_dr)
  "person_dr",
  "dr_severity",

  # Comorbidity indices
  "CCI",
  "DCSI",

  # Treatment variables
  "catsurg_before_hitplus_cat",
  "glaucoma_bef_hitplus_cat",
  "otherretina_bef_hitplus_cat"
)

# Check completeness for each imputation
completeness_summary <- lapply(1:min(5, n_imputations), function(i) {
  imp_data <- imputed_datasets[[i]]

  # Check each key variable
  var_status <- lapply(key_imputed_vars, function(v) {
    if (v %in% names(imp_data)) {
      n_missing <- sum(is.na(imp_data[[v]]))
      pct_missing <- round(100 * n_missing / nrow(imp_data), 2)
      data.frame(
        Variable = v,
        Imputation = i,
        N_Missing = n_missing,
        Pct_Missing = pct_missing,
        Status = ifelse(n_missing == 0, "Complete", paste0(n_missing, " missing"))
      )
    } else {
      data.frame(
        Variable = v,
        Imputation = i,
        N_Missing = NA,
        Pct_Missing = NA,
        Status = "Variable not found"
      )
    }
  })

  do.call(rbind, var_status)
})

# Combine all results
completeness_df <- do.call(rbind, completeness_summary)

# Create summary table by variable across imputations
completeness_wide <- completeness_df |>
  select(Variable, Imputation, N_Missing) |>
  pivot_wider(names_from = Imputation, 
              values_from = N_Missing,
              names_prefix = "Imp_")

kable(completeness_wide, 
      caption = "Missing values by variable across first 5 imputations (should all be 0)",
      align = c('l', rep('r', min(5, n_imputations))))

# Flag any variables with remaining missing data
problem_vars <- completeness_df |>
  filter(N_Missing > 0) |>
  select(Variable, Imputation, N_Missing) |>
  unique()

if (nrow(problem_vars) > 0) {
  cat("\n⚠️ WARNING: The following variables still have missing data after imputation:\n")
  kable(problem_vars)
} else {
  cat("\n✓ All key variables are complete across all checked imputations\n")
}

# Compare before and after imputation
cat("\n### Before vs After Imputation Summary\n")

original_missing <- mice_results$nmis[key_imputed_vars[key_imputed_vars %in% names(mice_results$nmis)]]
post_imp_missing <- completeness_df |>
  filter(Imputation == 1) |>
  group_by(Variable) |>
  summarise(Post_Missing = sum(N_Missing, na.rm = TRUE)) |>
  filter(Variable %in% names(original_missing))

comparison_df <- data.frame(
  Variable = names(original_missing),
  Original_Missing = original_missing,
  Post_Imputation_Missing = 0  # Will be updated if any found
)

# Update with actual post-imputation values if available
for (v in comparison_df$Variable) {
  if (v %in% post_imp_missing$Variable) {
    comparison_df[comparison_df$Variable == v, "Post_Imputation_Missing"] <- 
      post_imp_missing[post_imp_missing$Variable == v, "Post_Missing"]
  }
}

comparison_df$Status <- ifelse(comparison_df$Post_Imputation_Missing == 0, 
                               "✓ Complete", 
                               "⚠️ Still missing")

kable(comparison_df,
      caption = "Imputation effectiveness: Original vs Post-imputation missing counts",
      col.names = c("Variable", "Original Missing", "After Imputation", "Status"),
      row.names = FALSE)
```

**Number of incomplete cases in PS model variables per imputed dataset:**
```{r completeness-results}
print(unlist(completeness_check))
```

### DR Severity Variable Check

```{r dr-severity-check}
# Verify dr_severity was created properly in all datasets
dr_severity_summary <- lapply(1:min(3, n_imputations), function(i) {
  imp_data <- imputed_datasets[[i]]

  # Check dr_severity exists and is a factor
  if (!"dr_severity" %in% names(imp_data)) {
    return(data.frame(imputation = i, status = "MISSING"))
  }

  # Get summary
  dr_table <- table(imp_data$dr_severity, useNA = "ifany")

  data.frame(
    imputation = i,
    No_DR = dr_table["No_DR"],
    NPDR = dr_table["NPDR"],
    PDR = dr_table["PDR"],
    Missing = sum(is.na(imp_data$dr_severity)),
    is_factor = is.factor(imp_data$dr_severity),
    n_levels = nlevels(imp_data$dr_severity)
  )
})

dr_summary_df <- do.call(rbind, dr_severity_summary)
kable(dr_summary_df, caption = "DR severity distribution in first 3 imputed datasets")

# Confirm factor levels are correct
if (is.factor(imputed_datasets[[1]]$dr_severity)) {
  cat("DR severity levels:", paste(levels(imputed_datasets[[1]]$dr_severity), collapse = ", "), "\n")
}
```

### Patient-Level Variable Consistency Check

```{r patient-consistency-check}
# Check that patient-level variables are consistent within patients after imputation
consistency_check <- lapply(1:n_imputations, function(i) {
  imp_data <- imputed_datasets[[i]]

  # Check consistency of person_dr, CCI, and DCSI within patients
  patient_check <- imp_data |>
    group_by(e_mrn_deidentified) |>
    summarise(
      n_eyes = n(),
      unique_dr = n_distinct(person_dr, na.rm = TRUE),
      unique_cci = n_distinct(CCI, na.rm = TRUE),
      unique_dcsi = n_distinct(DCSI, na.rm = TRUE),
      dr_consistent = unique_dr <= 1,
      cci_consistent = unique_cci <= 1,
      dcsi_consistent = unique_dcsi <= 1,
      all_consistent = dr_consistent & cci_consistent & dcsi_consistent
    )

  return(list(
    dataset = i,
    n_patients = nrow(patient_check),
    dr_pct_consistent = round(mean(patient_check$dr_consistent) * 100, 2),
    cci_pct_consistent = round(mean(patient_check$cci_consistent) * 100, 2),
    dcsi_pct_consistent = round(mean(patient_check$dcsi_consistent) * 100, 2),
    all_pct_consistent = round(mean(patient_check$all_consistent) * 100, 2)
  ))
})

# Display results
consistency_df <- bind_rows(consistency_check)
log_message(paste0("Patient-level consistency - person_dr: ", 
                  min(consistency_df$dr_pct_consistent), "%, ",
                  "CCI: ", min(consistency_df$cci_pct_consistent), "%, ",
                  "DCSI: ", min(consistency_df$dcsi_pct_consistent), "%"))
```

**Patient-level variable consistency after imputation:**
```{r consistency-table}
kable(consistency_df, 
      col.names = c("Dataset", "Patients", "DR %", "CCI %", "DCSI %", "All %"),
      caption = "Percentage of patients with consistent values across eyes")
```

## Save Imputed Datasets

```{r save-imputed}
# Only save mice_results if it wasn't loaded from cache (to avoid overwriting good data)
if (!use_cached_mice || !file.exists(mice_cache_file)) {
  saveRDS(mice_results, file.path(reanalysis_data_dir, "mice_results.rds"))
  log_message("Saved fresh MICE results to mice_results.rds")
} else {
  log_message("Skipping mice_results.rds save (using cached version)")
}

# Always save the post-processed imputed datasets (these have our fixes applied)
saveRDS(imputed_datasets, file.path(reanalysis_data_dir, "imputed_datasets.rds"))
saveRDS(imputed_datasets[[1]], file.path(reanalysis_data_dir, "imputed_dataset_1.rds"))
log_message("Saved post-processed imputed datasets")
```

**Saved `r n_imputations` imputed datasets to:** `r reanalysis_data_dir`

```{r empty-chunk}
```

## Summary

```{r summary}
# Count variables more precisely
# Total columns in df_mice includes row_id which we added
total_vars <- ncol(df_mice)
id_vars <- 2  # row_id and e_mrn_deidentified
substantive_vars <- total_vars - id_vars  # Variables actually used in the imputation model

summary_table <- tibble(
  Metric = c(
    "Original observations",
    "Total variables in dataset",
    "ID variables (excluded from model)",
    "Substantive variables in imputation model",
    "Variables with missing data",
    "Variables actually imputed",
    "Number of imputations",
    "Maximum missing percentage",
    "Imputation iterations"
  ),
  Value = c(
    nrow(df_mice),
    total_vars,
    id_vars,
    substantive_vars,
    sum(mice_results$nmis > 0),
    length(vars_to_impute),
    n_imputations,
    paste0(round(max(mice_results$nmis/nrow(df_mice)) * 100, 1), "%"),
    mice_results$iteration
  )
)

summary_table |>
  kable(caption = "Multiple imputation summary")

imputed_vars <- names(mice_results$nmis)[mice_results$nmis > 0]
```

**Variables that were imputed:**
```{r imputed-vars-list}
print(imputed_vars)
```

### Temporal VA Imputation Summary

- **Baseline VA:** Starting point for temporal chain
- **Intermediate VA:** Bridge between baseline and 2-year outcome
  - For "submitted" cohort with observed outcome: set to outcome value
  - For "excluded" cohort: contains observed intermediate follow-up values
  - Reduces missing data substantially compared to leaving as NA
- **Outcome VA:** Final 2-year outcome (missing for 'excluded' cohort)
- **Temporal order enforced:** baseline → intermediate → outcome
- **No reverse-time prediction allowed**