### 2025-09-04 reanalysis-7-missingness_patterns.qmd

---
title: "Reanalysis Step 7: Sample Comparison Table"
subtitle: "Comparing Original, Complete Cases, and Imputed Data"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: html
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
source("settings.R")
setup_analysis(seed = 2025)

library(tableone)
library(knitr)
library(dplyr)

# Initialize logging
log_file <- init_log("sample_comparison")
```

## Load Data from Three Sources

```{r load-data}
# 1. Load original data (from analysis.Rmd)
data_path <- file.path(s_root, "Gina's Project/df_Final_240909.csv")
original_data <- load_and_prepare_data(data_path)
original_complete <- apply_exclusions(original_data)  # Remove missing CCI/DCSI

log_message(paste0("Original data after exclusions: ", nrow(original_complete), " observations"))

# 2. Load complete cases from reanalysis (reanalysis-1)
complete_cases_file <- file.path(reanalysis_data_dir, "df_complete_original.rds")
if (!file.exists(complete_cases_file)) {
  stop("Complete cases data not found. Run reanalysis-1-data_preparation.qmd first.")
}
complete_cases_data <- readRDS(complete_cases_file)

log_message(paste0("Complete cases data: ", nrow(complete_cases_data), " observations"))

# 3. Load imputed datasets (reanalysis-2)
imputed_file <- file.path(reanalysis_data_dir, "imputed_datasets.rds")
if (!file.exists(imputed_file)) {
  stop("Imputed datasets not found. Run reanalysis-2-multiple_imputation.qmd first.")
}
imputed_datasets <- readRDS(imputed_file)
n_imputations <- length(imputed_datasets)

log_message(paste0("Loaded ", n_imputations, " imputed datasets"))
```

## Create Comparison Table

```{r prepare-variables}
# Define variables for Table 1
vars_continuous <- c("age", "baseline_VA_logMAR", "CCI", "DCSI")
vars_categorical <- c("gender_cat", "race_ethnic_cat", "insurance_cat", "age_cat",
                      "dr_severity", "ever_lapse_binary", 
                      "glaucoma_bef_hitplus_cat", "otherretina_bef_hitplus_cat",
                      "catsurg_before_hitplus_cat",
                      "anti_VEGF", "PRP_flag", "focal_laser_flag", "other_inject",
                      "outcome_va_vi_binary")

all_vars <- c(vars_continuous, vars_categorical)

# Function to compute summary statistics for continuous variables
summarize_continuous <- function(data, var) {
  values <- data[[var]]
  values <- values[!is.na(values)]
  if (length(values) == 0) {
    return(data.frame(
      variable = var,
      mean = NA,
      sd = NA,
      median = NA,
      q25 = NA,
      q75 = NA,
      n_missing = nrow(data),
      pct_missing = 100
    ))
  }

  data.frame(
    variable = var,
    mean = mean(values, na.rm = TRUE),
    sd = sd(values, na.rm = TRUE),
    median = median(values, na.rm = TRUE),
    q25 = quantile(values, 0.25, na.rm = TRUE),
    q75 = quantile(values, 0.75, na.rm = TRUE),
    n_missing = sum(is.na(data[[var]])),
    pct_missing = 100 * sum(is.na(data[[var]])) / nrow(data)
  )
}

# Function to compute summary statistics for categorical variables
summarize_categorical <- function(data, var) {
  tbl <- table(data[[var]], useNA = "ifany")
  props <- prop.table(table(data[[var]], useNA = "no"))

  levels_df <- data.frame(
    variable = paste0(var, "_", names(props)),
    n = as.vector(table(data[[var]], useNA = "no")),
    pct = 100 * as.vector(props)
  )

  # Add missing info
  n_missing <- sum(is.na(data[[var]]))
  levels_df$n_missing <- n_missing
  levels_df$pct_missing <- 100 * n_missing / nrow(data)

  levels_df
}

# Process each dataset
process_dataset <- function(data, name) {
  # Continuous variables
  cont_summaries <- do.call(rbind, lapply(vars_continuous, function(v) {
    if (v %in% names(data)) {
      summarize_continuous(data, v)
    } else {
      data.frame(variable = v, mean = NA, sd = NA, median = NA, 
                 q25 = NA, q75 = NA, n_missing = NA, pct_missing = NA)
    }
  }))

  # Categorical variables
  cat_summaries <- do.call(rbind, lapply(vars_categorical, function(v) {
    if (v %in% names(data)) {
      summarize_categorical(data, v)
    } else {
      data.frame(variable = v, n = NA, pct = NA, n_missing = NA, pct_missing = NA)
    }
  }))

  list(continuous = cont_summaries, categorical = cat_summaries, dataset = name)
}
```

```{r create-table}
# Process original data
original_summary <- process_dataset(original_complete, "Original")

# Process complete cases
complete_summary <- process_dataset(complete_cases_data, "Complete Cases")

# Process imputed data (average across imputations)
# For imputed data, we need to average across all imputations
imputed_summaries <- lapply(1:n_imputations, function(i) {
  process_dataset(imputed_datasets[[i]], paste0("Imputation_", i))
})

# Average the imputed results
average_imputed_continuous <- do.call(rbind, lapply(vars_continuous, function(v) {
  # Extract values for this variable across all imputations
  imp_values <- lapply(imputed_summaries, function(s) {
    s$continuous[s$continuous$variable == v, ]
  })

  # Average the statistics
  if (nrow(imp_values[[1]]) > 0) {
    data.frame(
      variable = v,
      mean = mean(sapply(imp_values, function(x) x$mean), na.rm = TRUE),
      sd = mean(sapply(imp_values, function(x) x$sd), na.rm = TRUE),
      median = mean(sapply(imp_values, function(x) x$median), na.rm = TRUE),
      q25 = mean(sapply(imp_values, function(x) x$q25), na.rm = TRUE),
      q75 = mean(sapply(imp_values, function(x) x$q75), na.rm = TRUE),
      n_missing = 0,  # No missing in imputed data
      pct_missing = 0
    )
  } else {
    data.frame(variable = v, mean = NA, sd = NA, median = NA, 
               q25 = NA, q75 = NA, n_missing = NA, pct_missing = NA)
  }
}))

# For categorical variables, average the percentages
all_cat_vars <- unique(unlist(lapply(imputed_summaries, function(s) s$categorical$variable)))
average_imputed_categorical <- do.call(rbind, lapply(all_cat_vars, function(v) {
  # Extract values for this variable level across all imputations
  imp_values <- lapply(imputed_summaries, function(s) {
    s$categorical[s$categorical$variable == v, ]
  })

  # Remove empty results
  imp_values <- imp_values[sapply(imp_values, nrow) > 0]

  if (length(imp_values) > 0) {
    data.frame(
      variable = v,
      n = mean(sapply(imp_values, function(x) x$n), na.rm = TRUE),
      pct = mean(sapply(imp_values, function(x) x$pct), na.rm = TRUE),
      n_missing = 0,
      pct_missing = 0
    )
  } else {
    data.frame(variable = v, n = NA, pct = NA, n_missing = NA, pct_missing = NA)
  }
}))

imputed_summary <- list(
  continuous = average_imputed_continuous,
  categorical = average_imputed_categorical,
  dataset = "Imputed (Average)"
)
```

```{r display-table}
# Create formatted table for continuous variables
format_continuous_table <- function(original, complete, imputed) {
  vars <- unique(c(original$variable, complete$variable, imputed$variable))

  table_df <- data.frame(
    Variable = vars,
    Original = NA,
    Complete_Cases = NA,
    Imputed = NA,
    stringsAsFactors = FALSE
  )

  for (i in 1:length(vars)) {
    v <- vars[i]

    # Original
    orig_row <- original[original$variable == v, ]
    if (nrow(orig_row) > 0 && !is.na(orig_row$mean)) {
      table_df$Original[i] <- sprintf("%.2f (%.2f)", orig_row$mean, orig_row$sd)
    }

    # Complete cases
    comp_row <- complete[complete$variable == v, ]
    if (nrow(comp_row) > 0 && !is.na(comp_row$mean)) {
      table_df$Complete_Cases[i] <- sprintf("%.2f (%.2f)", comp_row$mean, comp_row$sd)
    }

    # Imputed
    imp_row <- imputed[imputed$variable == v, ]
    if (nrow(imp_row) > 0 && !is.na(imp_row$mean)) {
      table_df$Imputed[i] <- sprintf("%.2f (%.2f)", imp_row$mean, imp_row$sd)
    }
  }

  table_df
}

# Create formatted table for categorical variables
format_categorical_table <- function(original, complete, imputed) {
  vars <- unique(c(original$variable, complete$variable, imputed$variable))

  table_df <- data.frame(
    Variable = vars,
    Original = NA,
    Complete_Cases = NA, 
    Imputed = NA,
    stringsAsFactors = FALSE
  )

  for (i in 1:length(vars)) {
    v <- vars[i]

    # Original
    orig_row <- original[original$variable == v, ]
    if (nrow(orig_row) > 0 && !is.na(orig_row$n)) {
      table_df$Original[i] <- sprintf("%.0f (%.1f%%)", orig_row$n, orig_row$pct)
    }

    # Complete cases
    comp_row <- complete[complete$variable == v, ]
    if (nrow(comp_row) > 0 && !is.na(comp_row$n)) {
      table_df$Complete_Cases[i] <- sprintf("%.0f (%.1f%%)", comp_row$n, comp_row$pct)
    }

    # Imputed
    imp_row <- imputed[imputed$variable == v, ]
    if (nrow(imp_row) > 0 && !is.na(imp_row$n)) {
      table_df$Imputed[i] <- sprintf("%.1f (%.1f%%)", imp_row$n, imp_row$pct)
    }
  }

  table_df
}

# Create the tables
continuous_table <- format_continuous_table(
  original_summary$continuous,
  complete_summary$continuous,
  imputed_summary$continuous
)

categorical_table <- format_categorical_table(
  original_summary$categorical,
  complete_summary$categorical,
  imputed_summary$categorical
)

# Display continuous variables
kable(continuous_table, 
      caption = "Continuous Variables: Mean (SD)",
      col.names = c("Variable", "Original Data", "Complete Cases", "Imputed (Average)"))

# Display categorical variables
kable(categorical_table,
      caption = "Categorical Variables: N (%)",
      col.names = c("Variable", "Original Data", "Complete Cases", "Imputed (Average)"))

# Sample sizes
sample_sizes <- data.frame(
  Dataset = c("Original (after exclusions)", "Complete Cases", "Imputed (each)"),
  N = c(nrow(original_complete), nrow(complete_cases_data), nrow(imputed_datasets[[1]]))
)

kable(sample_sizes, caption = "Sample Sizes")
```

```{r finalize}
# Finalize log
finalize_log()
```