### 2025-09-08 reanalysis-10-pooling_results.qmd
refactor pooling into separate fit and summarize steps

---
title: "Step 10: Pool Results Across PS Methods Using Rubin's Rules"
author: "Analysis Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    fig-width: 8
    fig-height: 6
    fig-dpi: 150
    embed-resources: true
    self-contained: true
---

## Overview

This step pools the outcome model results across all propensity score methods and multiple imputations using Rubin's rules.

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
source("alt_specs.R")
setup_analysis(seed = 2025)

library(ggplot2)
library(knitr)
library(kableExtra)
library(mice)      # For pooling with Rubin's rules
library(miceadds)  # Additional pooling utilities
library(mitools)   # More MI tools

# Initialize logging
log_file <- init_log("pooling_results")
log_message("Starting pooling of outcome model results")

# Configuration
outcome_cache_dir <- file.path(reanalysis_data_dir, "outcome_cache")
pooled_cache_dir <- file.path(reanalysis_data_dir, "ps_cache", "pooled_results")

if (!dir.exists(pooled_cache_dir)) {
  dir.create(pooled_cache_dir, recursive = TRUE)
}

# Summary file for tracking
summary_file <- file.path(pooled_cache_dir, "pooled_summary.csv")
```

## Load Cached Outcome Models

```{r load-outcomes}
# Define all methods to pool
all_methods <- c(
  "nearest_glm",
  "nearest_gam", 
  "nearest_gbm",
  "nearest_lasso",
  "nearest_rpart",
  "nearest_mahalanobis",
  "subclass_glm",
  "cbps",
  "entropy",
  "bart",
  "twang_gbm"
)

log_message(paste0("Will pool results for ", length(all_methods), " methods"))

# Function to load outcome results for a method
load_outcome_results <- function(method_name, outcome_cache_dir, n_imp = 50) {
  results <- list()
  n_loaded <- 0

  for (i in 1:n_imp) {
    cache_file <- file.path(outcome_cache_dir, 
                           paste0("outcome_", method_name, "_imp", i, ".rds"))
    if (file.exists(cache_file)) {
      results[[i]] <- readRDS(cache_file)
      n_loaded <- n_loaded + 1
    }
  }

  log_message(paste0("  ", method_name, ": Loaded ", n_loaded, " outcome models"))
  return(results)
}

# Load all outcome results
all_outcome_results <- list()
cat("\nLoading cached outcome models:\n")
cat(paste(rep("-", 50), collapse = ""), "\n")

for (method in all_methods) {
  all_outcome_results[[method]] <- load_outcome_results(method, outcome_cache_dir)
}

cat(paste(rep("-", 50), collapse = ""), "\n")
```

## Pool Results Using Rubin's Rules

```{r pool-results}
# Initialize storage for pooled results
pooled_results <- list()
summary_data <- data.frame()

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("POOLING RESULTS ACROSS IMPUTATIONS USING MICE\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Function to pool using mice package
pool_with_mice <- function(outcome_results, method_name) {
  # Extract coefficients and variance-covariance matrices
  coef_list <- list()
  vcov_list <- list()
  n_valid <- 0

  for (i in seq_along(outcome_results)) {
    if (!is.null(outcome_results[[i]]) && outcome_results[[i]]$success) {
      coef_list[[length(coef_list) + 1]] <- outcome_results[[i]]$coefficients
      vcov_list[[length(vcov_list) + 1]] <- outcome_results[[i]]$vcov
      n_valid <- n_valid + 1
    }
  }

  if (n_valid < 2) {
    return(list(success = FALSE, error = "Less than 2 valid imputations"))
  }

  # Pool using miceadds for more robust handling
  tryCatch({
    # Create mids-like object for pooling
    pooled <- miceadds::pool_mi(
      qhat = coef_list,  # List of coefficient vectors
      u = vcov_list,     # List of variance-covariance matrices
      method = "rubin"   # Use Rubin's rules
    )

    # Extract ever_lapse coefficient
    lapse_idx <- grep("ever_lapse", names(coef_list[[1]]))[1]
    if (is.na(lapse_idx)) {
      return(list(success = FALSE, error = "Could not find ever_lapse coefficient"))
    }

    # Get pooled estimate for lapse
    estimate <- pooled$qbar[lapse_idx]
    se <- sqrt(pooled$t[lapse_idx, lapse_idx])
    df <- pooled$df[lapse_idx]

    # Calculate CI and p-value
    t_crit <- qt(0.975, df)
    ci_lower <- estimate - t_crit * se
    ci_upper <- estimate + t_crit * se
    t_stat <- estimate / se
    p_value <- 2 * pt(abs(t_stat), df, lower.tail = FALSE)

    # Convert to OR scale
    or <- exp(estimate)
    or_ci_lower <- exp(ci_lower)
    or_ci_upper <- exp(ci_upper)

    return(list(
      success = TRUE,
      method = method_name,
      estimate = estimate,
      se = se,
      df = df,
      ci_lower = ci_lower,
      ci_upper = ci_upper,
      t_stat = t_stat,
      p_value = p_value,
      or = or,
      or_ci_lower = or_ci_lower,
      or_ci_upper = or_ci_upper,
      n_imputations = n_valid,
      fmi = pooled$fmi[lapse_idx] # Fraction of missing information
    ))
  }, error = function(e) {
    # Fallback to manual Rubin's rules if miceadds fails
    return(pool_single_method(outcome_results, method_name))
  })
}

for (method in all_methods) {
  cat(paste0("\nProcessing ", method, "...\n"))

  outcome_results <- all_outcome_results[[method]]

  if (length(outcome_results) == 0) {
    cat(paste0("  WARNING: No outcome results found for ", method, "\n"))
    log_message(paste0("WARNING: No results for ", method))
    next
  }

  # Pool using mice/miceadds
  tryCatch({
    pooled <- pool_with_mice(outcome_results, method)

    if (pooled$success) {
      pooled_results[[method]] <- pooled

      # Save individual pooled result
      pooled_file <- file.path(pooled_cache_dir, paste0("pooled_", method, ".rds"))
      saveRDS(pooled, pooled_file)

      # Log the pooled estimate
      cat(paste0("  ✓ Pooled ", pooled$n_imputations, " imputations\n"))
      cat(paste0("    OR = ", sprintf("%.3f", pooled$or), 
                " (95% CI: ", sprintf("%.3f", pooled$or_ci_lower),
                " - ", sprintf("%.3f", pooled$or_ci_upper), ")\n"))
      cat(paste0("    p-value = ", sprintf("%.4f", pooled$p_value), "\n"))
      if (!is.null(pooled$fmi)) {
        cat(paste0("    FMI = ", sprintf("%.3f", pooled$fmi), "\n"))
      }

      log_message(paste0(method, ": OR = ", sprintf("%.3f", pooled$or),
                        " (", sprintf("%.3f", pooled$or_ci_lower),
                        " - ", sprintf("%.3f", pooled$or_ci_upper),
                        "), p = ", sprintf("%.4f", pooled$p_value)))

      # Add to summary data
      new_row <- data.frame(
        method = method,
        or = pooled$or,
        ci_lower = pooled$or_ci_lower,
        ci_upper = pooled$or_ci_upper,
        p_value = pooled$p_value,
        n_imputations = pooled$n_imputations,
        fmi = ifelse(!is.null(pooled$fmi), pooled$fmi, NA),
        timestamp = Sys.time(),
        stringsAsFactors = FALSE
      )
      summary_data <- rbind(summary_data, new_row)

    } else {
      cat(paste0("  ✗ Failed: ", pooled$error, "\n"))
      log_message(paste0("ERROR: Failed to pool ", method, " - ", pooled$error))
    }

  }, error = function(e) {
    cat(paste0("  ✗ Error: ", e$message, "\n"))
    log_message(paste0("ERROR: ", method, " - ", e$message))
  })
}

# Save summary CSV
if (nrow(summary_data) > 0) {
  write.csv(summary_data, summary_file, row.names = FALSE)
  log_message(paste0("Saved summary to ", summary_file))
}

# Save all pooled results
saveRDS(pooled_results, file.path(pooled_cache_dir, "all_pooled_results.rds"))
log_message("Saved all pooled results")

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("POOLING COMPLETE\n")
cat(paste(rep("=", 60), collapse = ""), "\n")
```

## Summary Table

```{r summary-table}
if (length(pooled_results) > 0) {
  # Create formatted summary table
  summary_table <- create_summary_table(pooled_results)

  # Display with kable
  # Add FMI column if available
  col_names <- c("Method", "Odds Ratio", "95% CI", "P-value", "N Imputations")
  if ("FMI" %in% colnames(summary_table)) {
    col_names <- c(col_names, "FMI")
  }

  kable(summary_table, 
        caption = "Pooled Effect Estimates Across Propensity Score Methods (Using mice/miceadds)",
        col.names = col_names,
        align = c("l", "r", "c", "r", "r", if("FMI" %in% colnames(summary_table)) "r" else NULL)) %>%
    kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                  full_width = FALSE)

  # Save to CSV
  write.csv(summary_table, 
            file.path(reanalysis_data_dir, "ps_methods_pooled_results.csv"),
            row.names = FALSE)

  log_message("Summary table saved to CSV")
} else {
  cat("No pooled results available for summary table\n")
}
```

## Forest Plot

```{r forest-plot}
#| fig.height: 10
#| fig.width: 8

if (length(pooled_results) > 0) {
  # Format data for forest plot
  forest_data <- format_for_forest_plot(pooled_results)

  # Add colors for specific methods
  forest_data$color <- "black"
  if ("twang_gbm" %in% forest_data$method) {
    forest_data$color[forest_data$method == "twang_gbm"] <- "blue"
  }

  # Create the plot
  forest_plot <- ggplot(forest_data, aes(x = reorder(method_label, or), y = or)) +
    geom_pointrange(aes(ymin = lower, ymax = upper, color = color), size = 0.8) +
    geom_point(aes(color = color), size = 3) +
    geom_text(aes(label = sprintf("%.2f", or), color = color), 
              vjust = -1.2, hjust = 0.5, size = 4) +
    coord_flip() +
    scale_color_identity() +
    scale_y_continuous(breaks = seq(0.5, max(forest_data$upper, 1.5), by = 0.1),
                       trans = "log",
                       labels = function(x) sprintf("%.1f", x)) +
    scale_x_discrete(expand = c(0.075, 0.075)) +
    theme_minimal(base_size = 14) +
    labs(
      x = element_blank(),
      y = "Odds Ratio",
      title = "Propensity Score Method Comparison",
      subtitle = "Pooled estimates across multiple imputations using mice package (Rubin's rules)",
      caption = "All methods use ATT (Average Treatment Effect in the Treated) estimand"
    ) +
    theme(
      plot.title = element_text(hjust = 0.5, face = "bold", margin = margin(b = 10)),
      plot.subtitle = element_text(hjust = 0.5, margin = margin(b = 15)),
      legend.position = "none",
      axis.line = element_line(color = "gray80"),
      axis.text = element_text(color = "gray30"),
      axis.ticks = element_line(color = "gray80"),
      axis.title.x = element_text(margin = margin(t = 10)),
      plot.caption = element_text(margin = margin(t = 15), hjust = 0),
      plot.margin = margin(20, 20, 30, 20)
    ) +
    geom_hline(yintercept = 1, linetype = "dashed", color = "gray70")

  print(forest_plot)

  # Save the plot
  ggsave(
    filename = file.path(reanalysis_data_dir, "forest_plot_pooled_results.png"),
    plot = forest_plot,
    width = 10,
    height = 8,
    dpi = 300
  )

  log_message("Forest plot saved")
} else {
  cat("No pooled results available for forest plot\n")
}
```

## Method Comparison

```{r method-comparison}
if (length(pooled_results) > 0) {
  cat("\n", paste(rep("=", 50), collapse = ""), "\n")
  cat("METHOD COMPARISON\n")
  cat(paste(rep("=", 50), collapse = ""), "\n\n")

  # Use twang_gbm as reference if available
  reference_method <- if ("twang_gbm" %in% names(pooled_results)) "twang_gbm" else names(pooled_results)[1]

  ref_or <- pooled_results[[reference_method]]$or
  cat(paste0("Reference method: ", reference_method, " (OR = ", sprintf("%.3f", ref_or), ")\n\n"))

  # Compare all methods
  comparison_data <- data.frame()
  for (method in names(pooled_results)) {
    res <- pooled_results[[method]]
    rel_diff <- 100 * (res$or / ref_or - 1)

    cat(sprintf("%-20s: OR = %.3f (%+.1f%% vs reference)\n",
                method, res$or, rel_diff))

    comparison_data <- rbind(comparison_data, data.frame(
      method = method,
      or = res$or,
      relative_diff = rel_diff,
      stringsAsFactors = FALSE
    ))
  }

  # Save comparison
  write.csv(comparison_data, 
            file.path(reanalysis_data_dir, "method_comparison.csv"),
            row.names = FALSE)

  log_message("Method comparison saved")
}
```

## Session Info

```{r session-info}
log_message("Pooling analysis complete")
finalize_log(success = TRUE)
sessionInfo()
```