### 2025-09-08 reanalysis-9-outcome_models.qmd

---
title: "Reanalysis Step 9: Fit Outcome Models"
subtitle: "Fit outcome models for all propensity score methods across imputations"
author: "Erik Westlund"
date: "`r Sys.Date()`"
format: 
  html:
    toc: true
    toc-depth: 2
    code-fold: true
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")  # Load formula definitions and helper functions
source("alt_specs.R")  # Load strategy pattern implementation
setup_analysis(seed = 2025)

library(ggplot2)
library(knitr)

# Initialize logging
log_file <- init_log("alternative_ps_methods")

# Check for required files
required_files <- c(
  "imputed_datasets.rds",
  "variable_lists.rds"
)

missing_files <- c()
for (file in required_files) {
  if (!file.exists(file.path(reanalysis_data_dir, file))) {
    missing_files <- c(missing_files, file)
  }
}

if (length(missing_files) > 0) {
  stop(paste("Required files not found. Please run previous steps first:\n",
             "- Step 1 (Data Preparation) creates: variable_lists.rds\n",
             "- Step 2 (Multiple Imputation) creates: imputed_datasets.rds\n",
             "Missing files:", paste(missing_files, collapse = ", ")))
}

# Load imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))

# ANALYSIS MODE: "test" or "final"
analysis_mode <- "final"  # Change to "test" for quick testing with 2 imputations

# Get configuration based on mode
config <- get_analysis_config(analysis_mode)

# Cache configuration
use_cache <- TRUE
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")  # Now using unified ps_cache
outcome_cache_dir <- file.path(reanalysis_data_dir, "outcome_cache")  # Separate outcome cache

log_message(paste0("Analysis mode: ", analysis_mode))
log_message(paste0("Processing up to ", config$n_imputations, " of ", n_imputations, " available imputations"))
log_message(paste0("Caching: ", ifelse(use_cache, "ENABLED", "DISABLED")))
log_message(paste0("Cache directory: ", cache_dir))
```

**Analysis Configuration:**
- Mode: `r analysis_mode`
- Maximum imputations to process: `r config$n_imputations`
- Available imputations: `r n_imputations`
- Cache: `r ifelse(use_cache, "Enabled", "Disabled")`

## Define Formulas

```{r define-formulas}
# Get PS formula using reanalysis version (includes dr_severity)
ps_formula <- get_matching_formula_reanalysis()

# Get outcome formula
outcome_formulas <- get_analysis_formulas_reanalysis()
outcome_formula <- outcome_formulas$full

log_message("Propensity Score Formula:")
print(ps_formula)

log_message("Outcome Formula:")
print(outcome_formula)
```

## Available Methods

```{r show-methods}
# Get list of available methods
available_methods <- names(get_ps_methods(config))

cat("\nAvailable PS methods to run:\n")
for (method in available_methods) {
  cat(" -", method, "\n")
}

# Define all available methods
all_methods <- c(
  "nearest_glm",
  "nearest_gam", 
  "nearest_gbm",
  "nearest_lasso",
  "nearest_rpart",
  "nearest_mahalanobis",
  "subclass_glm",
  "cbps",
  "entropy",
  "bart",
  "twang_gbm"  # Will load from reanalysis-4 if available
)

# Force rerun flag - set to TRUE to reprocess all methods even if cached
force_rerun <- FALSE  # Set to TRUE to rerun everything

log_message(paste0("Will process up to ", length(all_methods), " methods"))
log_message(paste0("Force rerun: ", ifelse(force_rerun, "YES", "NO")))
```

## Fit Outcome Models for All Methods

```{r fit-outcome-models}
# Only fit outcome models - no pooling in this step
# Get PS methods configuration
ps_methods <- get_ps_methods(config)

# Track progress
all_outcome_results <- list()
success_count <- 0
failure_count <- 0

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("FITTING OUTCOME MODELS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Process each method
for (method_name in all_methods) {
  cat(paste0("\nProcessing ", method_name, "...\n"))
  method_outcome_results <- list()
  n_fitted <- 0

  # Process each imputation
  for (i in 1:config$n_imputations) {
    # Check if outcome model already cached
    outcome_cache_file <- file.path(outcome_cache_dir, 
                                   paste0("outcome_", method_name, "_imp", i, ".rds"))

    if (file.exists(outcome_cache_file) && !force_rerun) {
      cat(paste0("  Imputation ", i, ": Using cached outcome model\n"))
      method_outcome_results[[i]] <- readRDS(outcome_cache_file)
      n_fitted <- n_fitted + 1
      next
    }

    # Load PS results
    ps_cache_file <- file.path(cache_dir, paste0(method_name, "_imp", i, ".rds"))

    if (!file.exists(ps_cache_file)) {
      cat(paste0("  Imputation ", i, ": No PS results found, skipping\n"))
      next
    }

    ps_result <- readRDS(ps_cache_file)

    if (!ps_result$success) {
      cat(paste0("  Imputation ", i, ": PS method failed, skipping\n"))
      next
    }

    # Get the imputed dataset
    imp_data <- imputed_datasets[[i]]

    # Fit outcome model
    cat(paste0("  Imputation ", i, ": Fitting outcome model...\n"))

    outcome_result <- tryCatch({
      fit_outcome_model_cached(
        ps_result = ps_result,
        outcome_formula = outcome_formula,
        data = imp_data,
        method_name = method_name,
        imputation = i,
        cache_dir = outcome_cache_dir
      )
    }, error = function(e) {
      log_message(paste0("ERROR: ", method_name, " imp", i, " - ", e$message))
      list(success = FALSE, error = e$message)
    })

    if (outcome_result$success) {
      method_outcome_results[[i]] <- outcome_result
      n_fitted <- n_fitted + 1
      cat(paste0("    ✓ Outcome model fitted and cached\n"))
    } else {
      cat(paste0("    ✗ Failed: ", outcome_result$error, "\n"))
    }

    # Periodic garbage collection
    if (i %% 5 == 0) {
      gc(verbose = FALSE)
    }
  }

  # Store results for this method
  all_outcome_results[[method_name]] <- method_outcome_results

  if (n_fitted > 0) {
    success_count <- success_count + 1
    cat(paste0("✓ ", method_name, ": Fitted ", n_fitted, " outcome models\n"))
    log_message(paste0(method_name, ": Successfully fitted ", n_fitted, " outcome models"))
  } else {
    failure_count <- failure_count + 1
    cat(paste0("✗ ", method_name, ": No outcome models fitted\n"))
    log_message(paste0("WARNING: ", method_name, " - No outcome models fitted"))
  }
}

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("OUTCOME MODEL FITTING COMPLETE\n")
cat(paste0("Successful methods: ", success_count, "\n"))
cat(paste0("Failed methods: ", failure_count, "\n"))
cat(paste(rep("=", 60), collapse = ""), "\n")

# Save summary of what was fitted
summary_info <- data.frame(
  method = names(all_outcome_results),
  n_models = sapply(all_outcome_results, length),
  timestamp = Sys.time()
)
write.csv(summary_info, 
          file.path(outcome_cache_dir, "outcome_fitting_summary.csv"),
          row.names = FALSE)

log_message(paste0("Outcome model fitting complete. ",
                  "Successful methods: ", success_count,
                  ", Failed methods: ", failure_count))
```

## Outcome Model Fitting Summary

```{r outcome-summary}
# Summarize what was fitted
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("OUTCOME MODEL FITTING SUMMARY\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Check what's in the outcome cache
if (dir.exists(outcome_cache_dir)) {
  outcome_files <- list.files(outcome_cache_dir, pattern = "outcome_.*_imp\\d+\\.rds")

  if (length(outcome_files) > 0) {
    # Extract method names from filenames
    methods_in_cache <- unique(gsub("outcome_(.*)_imp\\d+\\.rds", "\\1", outcome_files))

    cat("Methods with cached outcome models:\n")
    for (method in sort(methods_in_cache)) {
      method_files <- grep(paste0("outcome_", method, "_imp"), outcome_files, value = TRUE)
      cat(sprintf("  %-20s: %d models\n", method, length(method_files)))
    }

    cat(paste0("\nTotal outcome models cached: ", length(outcome_files), "\n"))
  } else {
    cat("No outcome models found in cache.\n")
  }
} else {
  cat("Outcome cache directory does not exist.\n")
}
```

## Summary

This step fits outcome models for **`r length(all_methods)` different propensity score methods** across **`r config$n_imputations` multiply imputed datasets**.

### Analysis Details:
- **Estimand:** ATT (Average Treatment Effect in the Treated)
- **Imputation method:** Multiple imputation using mice
- **Outcome model:** Survey-weighted logistic regression with cluster-robust SEs
- **Results saved to:** `outcome_cache/` directory

### Next Steps:
Run `reanalysis-10-pooling_results.qmd` to:
1. Pool results across imputations using Rubin's rules
2. Generate forest plots and comparison tables
3. Create publication-ready summaries

## Session Info

```{r session-info}
# Finalize logging
log_message("Analysis completed successfully")
finalize_log(success = TRUE)

# Print session information
sessionInfo()
```