### 2025-09-08 reanalysis-11-forest_plot.qmd
refactor forest plot

---
title: "Step 11: Forest Plot Visualization"
subtitle: "Generate forest plot from pooled propensity score results"
author: "Analysis Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 2
    code-fold: true
    code-summary: "Show code"
    fig-width: 10
    fig-height: 8
    fig-dpi: 300
    embed-resources: true
    self-contained: true
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
source("alt_specs.R")
setup_analysis(seed = 2025)

library(ggplot2)
library(knitr)
library(kableExtra)

# Initialize logging
log_file <- init_log("forest_plot")
log_message("Starting forest plot generation")

# Configuration
pooled_cache_dir <- file.path(reanalysis_data_dir, "ps_cache", "pooled_results")
```

## Load Pooled Results

```{r load-results}
# Load the pooled results from file 10
pooled_results_file <- file.path(pooled_cache_dir, "all_pooled_results.rds")

if (!file.exists(pooled_results_file)) {
  stop(paste0("Pooled results not found at: ", pooled_results_file, "\n",
              "Please run reanalysis-10-pooling_results.qmd first"))
}

pooled_results <- readRDS(pooled_results_file)
log_message(paste0("Loaded pooled results for ", length(pooled_results), " methods"))

# Show which methods we have
cat("Methods with pooled results:\n")
for (method in names(pooled_results)) {
  res <- pooled_results[[method]]
  if (res$success) {
    cat(sprintf("  %-20s: OR = %.3f (%.3f - %.3f), p = %.4f\n",
                method, res$or, res$or_ci_lower, res$or_ci_upper, res$p_value))
  } else {
    cat(sprintf("  %-20s: FAILED\n", method))
  }
}
```

## Format Data for Plotting

```{r format-data}
# Use the existing format function
forest_data <- format_for_forest_plot(pooled_results)

# Add colors for specific methods
forest_data$color <- "black"
if ("twang_gbm" %in% forest_data$method) {
  forest_data$color[forest_data$method == "twang_gbm"] <- "blue"
}

# Sort by OR for better visualization
forest_data <- forest_data[order(forest_data$or), ]

# Display the data
kable(forest_data[, c("method_label", "or", "lower", "upper", "p_value")],
      caption = "Data for Forest Plot",
      col.names = c("Method", "OR", "Lower CI", "Upper CI", "P-value"),
      digits = 3)
```

## Create Forest Plot

```{r forest-plot}
#| fig.height: 10
#| fig.width: 8

# Create the forest plot
forest_plot <- ggplot(forest_data, aes(x = reorder(method_label, or), y = or)) +
  # Point estimates with confidence intervals
  geom_pointrange(aes(ymin = lower, ymax = upper, color = color), 
                  size = 0.8, linewidth = 0.8) +
  # Larger points for estimates
  geom_point(aes(color = color), size = 3) +
  # Add OR values as text
  geom_text(aes(label = sprintf("%.2f", or), color = color), 
            vjust = -1.2, hjust = 0.5, size = 4) +
  # Flip coordinates for horizontal forest plot
  coord_flip() +
  # Use the color mapping
  scale_color_identity() +
  # Log scale for OR with specific breaks
  scale_y_continuous(
    breaks = c(0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.7),
    trans = "log",
    labels = function(x) sprintf("%.1f", x),
    limits = c(min(forest_data$lower) * 0.95, max(forest_data$upper) * 1.05)
  ) +
  # Expand x axis slightly for readability
  scale_x_discrete(expand = c(0.075, 0.075)) +
  # Minimal theme
  theme_minimal(base_size = 14) +
  # Labels
  labs(
    x = element_blank(),
    y = "Odds Ratio (95% CI)",
    title = "Propensity Score Method Comparison",
    subtitle = paste0("Pooled estimates across ", 
                     ifelse(length(unique(forest_data$n_imputations)) == 1,
                            paste0(unique(forest_data$n_imputations)[1], " imputations"),
                            "multiple imputations"),
                     " using Rubin's rules"),
    caption = paste0("All methods use ATT (Average Treatment Effect in the Treated) estimand.\n",
                    ifelse("twang_gbm" %in% forest_data$method,
                          "twang GBM (shown in blue) is the primary method reported in the paper.",
                          ""))
  ) +
  # Theme customization - using explicit ggplot2:: namespace to avoid conflicts
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", 
                             margin = ggplot2::margin(b = 10)),
    plot.subtitle = element_text(hjust = 0.5, 
                                 margin = ggplot2::margin(b = 15)),
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "gray90", linetype = "dotted"),
    axis.line = element_line(color = "gray80"),
    axis.text = element_text(color = "gray30"),
    axis.text.y = element_text(size = 12),
    axis.ticks = element_line(color = "gray80"),
    axis.title.x = element_text(margin = ggplot2::margin(t = 10)),
    plot.caption = element_text(margin = ggplot2::margin(t = 15), hjust = 0),
    plot.margin = ggplot2::margin(20, 20, 30, 20)
  ) +
  # Reference line at OR = 1
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", linewidth = 0.75)

# Display the plot
print(forest_plot)

# Save the plot
plot_file <- file.path(reanalysis_data_dir, "forest_plot_ps_methods.png")
ggsave(
  filename = plot_file,
  plot = forest_plot,
  width = 10,
  height = 8,
  dpi = 300
)

log_message(paste0("Forest plot saved to: ", plot_file))
cat("\nForest plot saved to:", plot_file, "\n")
```

## Statistical Summary

```{r summary-stats}
# Calculate summary statistics
or_values <- forest_data$or
or_range <- range(or_values)
or_median <- median(or_values)
or_iqr <- IQR(or_values)

cat("\n", paste(rep("=", 50), collapse = ""), "\n")
cat("SUMMARY STATISTICS\n")
cat(paste(rep("=", 50), collapse = ""), "\n\n")

cat(sprintf("Number of methods: %d\n", nrow(forest_data)))
cat(sprintf("OR range: %.3f to %.3f\n", or_range[1], or_range[2]))
cat(sprintf("Median OR: %.3f\n", or_median))
cat(sprintf("IQR of ORs: %.3f\n", or_iqr))

# Find methods with smallest and largest effects
min_idx <- which.min(or_values)
max_idx <- which.max(or_values)

cat(sprintf("\nSmallest effect: %s (OR = %.3f)\n", 
            forest_data$method_label[min_idx], or_values[min_idx]))
cat(sprintf("Largest effect: %s (OR = %.3f)\n", 
            forest_data$method_label[max_idx], or_values[max_idx]))

# Reference to twang if available
if ("twang_gbm" %in% forest_data$method) {
  twang_idx <- which(forest_data$method == "twang_gbm")
  twang_or <- forest_data$or[twang_idx]

  cat(sprintf("\ntwang GBM (reference): OR = %.3f\n", twang_or))

  # Calculate how many methods are within 10% of twang
  within_10pct <- sum(abs(or_values / twang_or - 1) <= 0.10)
  cat(sprintf("Methods within 10%% of twang: %d of %d (%.1f%%)\n", 
              within_10pct, length(or_values), 
              100 * within_10pct / length(or_values)))
}

# Check for significant results
sig_results <- sum(forest_data$p_value < 0.05)
cat(sprintf("\nStatistically significant (p < 0.05): %d of %d methods\n",
            sig_results, nrow(forest_data)))
```

## Alternative Forest Plot (Sorted by Effect Size)

```{r forest-plot-sorted}
#| fig.height: 10
#| fig.width: 8

# Create alternative plot sorted by effect size
forest_plot_sorted <- ggplot(forest_data, aes(x = reorder(method_label, -or), y = or)) +
  geom_pointrange(aes(ymin = lower, ymax = upper, color = color), 
                  size = 0.8, linewidth = 0.8) +
  geom_point(aes(color = color), size = 3) +
  geom_text(aes(label = sprintf("%.2f", or), color = color), 
            vjust = -1.2, hjust = 0.5, size = 4) +
  coord_flip() +
  scale_color_identity() +
  scale_y_continuous(
    breaks = c(0.8, 0.9, 1.0, 1.1, 1.2, 1.3, 1.4, 1.5, 1.6, 1.7),
    trans = "log",
    labels = function(x) sprintf("%.1f", x),
    limits = c(min(forest_data$lower) * 0.95, max(forest_data$upper) * 1.05)
  ) +
  scale_x_discrete(expand = c(0.075, 0.075)) +
  theme_minimal(base_size = 14) +
  labs(
    x = element_blank(),
    y = "Odds Ratio (95% CI)",
    title = "Propensity Score Methods - Sorted by Effect Size",
    subtitle = "Largest to smallest odds ratios",
    caption = "Methods sorted from largest to smallest treatment effect estimate"
  ) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold", 
                             margin = ggplot2::margin(b = 10)),
    plot.subtitle = element_text(hjust = 0.5, 
                                 margin = ggplot2::margin(b = 15)),
    legend.position = "none",
    panel.grid.major.y = element_blank(),
    panel.grid.minor = element_blank(),
    panel.grid.major.x = element_line(color = "gray90", linetype = "dotted"),
    axis.line = element_line(color = "gray80"),
    axis.text = element_text(color = "gray30"),
    axis.text.y = element_text(size = 12),
    axis.ticks = element_line(color = "gray80"),
    axis.title.x = element_text(margin = ggplot2::margin(t = 10)),
    plot.caption = element_text(margin = ggplot2::margin(t = 15), hjust = 0),
    plot.margin = ggplot2::margin(20, 20, 30, 20)
  ) +
  geom_hline(yintercept = 1, linetype = "dashed", color = "gray50", linewidth = 0.75)

print(forest_plot_sorted)

# Save the sorted plot
ggsave(
  filename = file.path(reanalysis_data_dir, "forest_plot_ps_methods_sorted.png"),
  plot = forest_plot_sorted,
  width = 10,
  height = 8,
  dpi = 300
)
```

## Session Info

```{r session-info}
log_message("Forest plot generation complete")
finalize_log(success = TRUE)
sessionInfo()
```