### 2025-09-08 reanalysis-13-marginal_effects.qmd
ame

---
title: "Step 13: Marginal Effects Analysis with Multiple Imputation"
subtitle: "Calculate average marginal effects for twang model across subgroups"
author: "Analysis Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    self-contained: true
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
source("alt_specs.R")
setup_analysis(seed = 2025)

library(margins)
library(survey)
library(knitr)
library(kableExtra)

# Initialize logging
log_file <- init_log("marginal_effects")
log_message("Starting marginal effects analysis for twang models")

# Configuration
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
marginal_effects_dir <- file.path(reanalysis_data_dir, "marginal_effects")

# Create directory for marginal effects results
if (!dir.exists(marginal_effects_dir)) {
  dir.create(marginal_effects_dir, recursive = TRUE)
  log_message(paste0("Created marginal effects directory: ", marginal_effects_dir))
}

# Load imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))

# Get outcome formula
outcome_formulas <- get_analysis_formulas_reanalysis()
outcome_formula <- outcome_formulas$full
```

## Define Subgroups

```{r define-subgroups}
# Define the subgroups of interest (matching analysis.Rmd)
subgroups <- list(
  overall = list(
    name = "Overall (Average Marginal Effect)",
    subset_condition = NULL,  # No subsetting for overall
    description = "Average marginal effect across entire sample"
  ),

  pdr_any_tx = list(
    name = "PDR on any treatment",
    subset_condition = expression(dr_severity == "PDR" & 
                                 (PRP_flag == 1 | anti_VEGF == 1 | 
                                  focal_laser_flag == 1 | other_inject == 1)),
    description = "Patients with PDR receiving any treatment"
  ),

  pdr_anti_vegf_only = list(
    name = "PDR on anti-VEGF only",
    subset_condition = expression(dr_severity == "PDR" & 
                                 PRP_flag == 0 & anti_VEGF == 1 & 
                                 focal_laser_flag == 0 & other_inject == 0),
    description = "Patients with PDR receiving anti-VEGF only"
  ),

  pdr_prp_only = list(
    name = "PDR on PRP only",
    subset_condition = expression(dr_severity == "PDR" & 
                                 PRP_flag == 1 & anti_VEGF == 0 & 
                                 focal_laser_flag == 0 & other_inject == 0),
    description = "Patients with PDR receiving PRP only"
  )
)

# Display subgroup definitions
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("SUBGROUPS FOR MARGINAL EFFECTS ANALYSIS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("- ", sg$name, "\n"))
  cat(paste0("  ", sg$description, "\n\n"))
}
```

## Test on Single Imputation

```{r test-single}
# Test the approach on a single imputation first
test_imp <- 1
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat(paste0("TESTING ON IMPUTATION ", test_imp, "\n"))
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Load twang weights for this imputation
twang_file <- file.path(cache_dir, paste0("twang_gbm_imp", test_imp, ".rds"))
if (!file.exists(twang_file)) {
  stop(paste0("twang results not found for imputation ", test_imp, 
              ". Please run reanalysis-4-propensity_scores.qmd first"))
}

twang_result <- readRDS(twang_file)
test_data <- imputed_datasets[[test_imp]]

# Add weights to data
test_data$twang_att_w <- twang_result$weights

# Fit the outcome model with weights
design <- svydesign(ids = ~e_mrn_deidentified, 
                   weights = ~twang_att_w, 
                   data = test_data)
model <- svyglm(formula = outcome_formula, 
               design = design, 
               family = quasibinomial())

# Calculate marginal effects for each subgroup
test_results <- list()

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("\nCalculating marginal effects for: ", sg$name, "\n"))

  if (is.null(sg$subset_condition)) {
    # Overall marginal effect
    mfx <- margins(model, design = design, variables = "ever_lapse_binary")
    test_results[[sg_name]] <- summary(mfx)
    cat(paste0("  N = ", nrow(test_data), "\n"))
  } else {
    # Subset the data and design
    subset_data <- subset(test_data, eval(sg$subset_condition))
    subset_design <- subset(design, eval(sg$subset_condition))

    if (nrow(subset_data) > 0) {
      mfx <- margins(model, 
                    design = subset_design, 
                    data = subset_data,
                    variables = "ever_lapse_binary")
      test_results[[sg_name]] <- summary(mfx)
      cat(paste0("  N = ", nrow(subset_data), "\n"))
    } else {
      cat("  WARNING: No observations in this subgroup\n")
      test_results[[sg_name]] <- NULL
    }
  }

  # Display results if available
  if (!is.null(test_results[[sg_name]])) {
    mfx_summary <- test_results[[sg_name]]
    if ("ever_lapse_binary" %in% rownames(mfx_summary)) {
      ame <- mfx_summary["ever_lapse_binary", "AME"]
      se <- mfx_summary["ever_lapse_binary", "SE"]
      cat(sprintf("  AME = %.4f (SE = %.4f)\n", ame, se))
    }
  }
}

cat("\n✓ Test on single imputation successful\n")
```

## Calculate Marginal Effects for All Imputations

```{r calculate-all}
# Now run for all imputations
all_mfx_results <- list()

# Initialize storage for each subgroup
for (sg_name in names(subgroups)) {
  all_mfx_results[[sg_name]] <- list()
}

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("CALCULATING MARGINAL EFFECTS ACROSS ALL IMPUTATIONS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Process each imputation
for (i in 1:n_imputations) {
  if (i %% 10 == 1) {
    cat(paste0("Processing imputations ", i, "-", min(i+9, n_imputations), "...\n"))
  }

  # Load twang weights
  twang_file <- file.path(cache_dir, paste0("twang_gbm_imp", i, ".rds"))
  if (!file.exists(twang_file)) {
    cat(paste0("  WARNING: Skipping imputation ", i, " - twang results not found\n"))
    next
  }

  twang_result <- readRDS(twang_file)
  imp_data <- imputed_datasets[[i]]
  imp_data$twang_att_w <- twang_result$weights

  # Create survey design and fit model
  design <- svydesign(ids = ~e_mrn_deidentified, 
                     weights = ~twang_att_w, 
                     data = imp_data)

  tryCatch({
    model <- svyglm(formula = outcome_formula, 
                   design = design, 
                   family = quasibinomial())

    # Calculate marginal effects for each subgroup
    for (sg_name in names(subgroups)) {
      sg <- subgroups[[sg_name]]

      if (is.null(sg$subset_condition)) {
        # Overall marginal effect
        mfx <- margins(model, design = design, variables = "ever_lapse_binary")
        mfx_summary <- summary(mfx)
      } else {
        # Subset marginal effect
        subset_data <- subset(imp_data, eval(sg$subset_condition))

        if (nrow(subset_data) > 0) {
          subset_design <- subset(design, eval(sg$subset_condition))
          mfx <- margins(model, 
                        design = subset_design, 
                        data = subset_data,
                        variables = "ever_lapse_binary")
          mfx_summary <- summary(mfx)
        } else {
          mfx_summary <- NULL
        }
      }

      # Store results
      if (!is.null(mfx_summary) && "ever_lapse_binary" %in% rownames(mfx_summary)) {
        all_mfx_results[[sg_name]][[i]] <- list(
          ame = mfx_summary["ever_lapse_binary", "AME"],
          se = mfx_summary["ever_lapse_binary", "SE"],
          n = ifelse(is.null(sg$subset_condition), 
                    nrow(imp_data),
                    nrow(subset(imp_data, eval(sg$subset_condition)))),
          success = TRUE
        )
      } else {
        all_mfx_results[[sg_name]][[i]] <- list(success = FALSE)
      }
    }

  }, error = function(e) {
    cat(paste0("  ERROR in imputation ", i, ": ", e$message, "\n"))
    for (sg_name in names(subgroups)) {
      all_mfx_results[[sg_name]][[i]] <- list(success = FALSE, error = e$message)
    }
  })

  # Periodic garbage collection
  if (i %% 10 == 0) {
    gc(verbose = FALSE)
  }
}

cat("\n✓ Marginal effects calculation complete\n")

# Save intermediate results
saveRDS(all_mfx_results, file.path(marginal_effects_dir, "all_mfx_results.rds"))
log_message("Saved all marginal effects results")
```

## Pool Marginal Effects Using Rubin's Rules

```{r pool-mfx}
# Pool the marginal effects across imputations
pooled_mfx <- list()

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("POOLING MARGINAL EFFECTS USING RUBIN'S RULES\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("\n", sg$name, ":\n"))

  # Extract successful results
  sg_results <- all_mfx_results[[sg_name]]
  valid_results <- sg_results[sapply(sg_results, function(x) 
    !is.null(x) && x$success)]

  if (length(valid_results) < 2) {
    cat("  Insufficient valid imputations for pooling\n")
    pooled_mfx[[sg_name]] <- list(success = FALSE)
    next
  }

  # Extract AMEs and SEs
  ames <- sapply(valid_results, function(x) x$ame)
  ses <- sapply(valid_results, function(x) x$se)
  ns <- sapply(valid_results, function(x) x$n)

  # Apply Rubin's rules
  m <- length(ames)
  q_bar <- mean(ames)         # Combined estimate
  u_bar <- mean(ses^2)        # Within-imputation variance
  b <- var(ames)              # Between-imputation variance
  t <- u_bar + (1 + 1/m) * b  # Total variance
  se_pooled <- sqrt(t)

  # Degrees of freedom
  gamma <- ((1 + 1/m) * b) / t
  df <- (m - 1) / gamma^2

  # Confidence intervals and p-value
  t_crit <- qt(0.975, df)
  ci_lower <- q_bar - t_crit * se_pooled
  ci_upper <- q_bar + t_crit * se_pooled
  z_stat <- q_bar / se_pooled
  p_value <- 2 * pnorm(-abs(z_stat))

  # Average sample size
  avg_n <- mean(ns)

  # Store pooled results
  pooled_mfx[[sg_name]] <- list(
    name = sg$name,
    ame = q_bar,
    se = se_pooled,
    ci_lower = ci_lower,
    ci_upper = ci_upper,
    z_stat = z_stat,
    p_value = p_value,
    n_imputations = m,
    avg_n = avg_n,
    df = df,
    success = TRUE
  )

  # Display results
  cat(sprintf("  AME = %.4f (SE = %.4f)\n", q_bar, se_pooled))
  cat(sprintf("  95%% CI: (%.4f, %.4f)\n", ci_lower, ci_upper))
  cat(sprintf("  P-value = %.4f\n", p_value))
  cat(sprintf("  N imputations = %d\n", m))
  cat(sprintf("  Average N = %.0f\n", avg_n))
}

# Save pooled results
saveRDS(pooled_mfx, file.path(marginal_effects_dir, "pooled_mfx_results.rds"))
log_message("Saved pooled marginal effects")
```

## Summary Table

```{r summary-table}
# Create summary table
summary_df <- do.call(rbind, lapply(names(pooled_mfx), function(sg_name) {
  res <- pooled_mfx[[sg_name]]
  if (res$success) {
    data.frame(
      Subgroup = res$name,
      AME = sprintf("%.4f", res$ame),
      SE = sprintf("%.4f", res$se),
      `95% CI` = sprintf("(%.4f, %.4f)", res$ci_lower, res$ci_upper),
      `P-value` = ifelse(res$p_value < 0.001, "<0.001", 
                         sprintf("%.3f", res$p_value)),
      `Avg N` = sprintf("%.0f", res$avg_n),
      `N Imp` = res$n_imputations,
      stringsAsFactors = FALSE,
      check.names = FALSE
    )
  }
}))

# Display table
kable(summary_df,
      caption = "Pooled Average Marginal Effects of Lapsing by Subgroup",
      align = c("l", "r", "r", "c", "r", "r", "r")) %>%
  kable_styling(bootstrap_options = c("striped", "hover", "condensed"),
                full_width = FALSE) %>%
  add_header_above(c(" " = 1, 
                     "Marginal Effect" = 2,
                     "Statistical Inference" = 2,
                     "Sample" = 2))

# Save to CSV
write.csv(summary_df, 
          file.path(marginal_effects_dir, "marginal_effects_summary.csv"),
          row.names = FALSE)
log_message("Saved summary table to CSV")
```

## Interpretation

```{r interpretation}
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("INTERPRETATION OF RESULTS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

cat("Average Marginal Effect (AME) represents the average change in\n")
cat("probability of vision impairment/blindness associated with lapsing\n")
cat("from care, holding all other variables at their observed values.\n\n")

cat("Key findings:\n\n")

# Overall effect
if ("overall" %in% names(pooled_mfx) && pooled_mfx$overall$success) {
  overall_ame <- pooled_mfx$overall$ame
  overall_pct <- overall_ame * 100
  cat(sprintf("1. Overall: Lapsing is associated with a %.1f percentage point\n", overall_pct))
  cat(sprintf("   increase in probability of vision impairment (AME = %.4f)\n\n", overall_ame))
}

# PDR subgroups
pdr_groups <- c("pdr_any_tx", "pdr_anti_vegf_only", "pdr_prp_only")
for (sg_name in pdr_groups) {
  if (sg_name %in% names(pooled_mfx) && pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    pct_change <- sg$ame * 100
    cat(sprintf("2. %s:\n", sg$name))
    cat(sprintf("   AME = %.4f (%.1f percentage points)\n", sg$ame, pct_change))
    cat(sprintf("   Average N = %.0f patients\n\n", sg$avg_n))
  }
}

# Statistical significance
cat("Statistical significance (p < 0.05):\n")
for (sg_name in names(pooled_mfx)) {
  if (pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    sig <- ifelse(sg$p_value < 0.05, "Yes", "No")
    cat(sprintf("  - %s: %s (p = %.4f)\n", sg$name, sig, sg$p_value))
  }
}
```

## Export Results

```{r export}
# Create comprehensive export
export_list <- list(
  pooled_results = pooled_mfx,
  summary_table = summary_df,
  metadata = list(
    n_imputations = n_imputations,
    subgroups = names(subgroups),
    method = "twang_gbm",
    formula = deparse(outcome_formula),
    date = Sys.Date()
  )
)

# Save as RDS for R users
saveRDS(export_list, 
        file.path(marginal_effects_dir, "marginal_effects_export.rds"))

# Create text summary for sharing
summary_text <- c(
  "MARGINAL EFFECTS ANALYSIS RESULTS",
  paste(rep("=", 60), collapse = ""),
  paste0("Method: twang GBM with IPW"),
  paste0("Number of imputations: ", n_imputations),
  paste0("Date: ", Sys.Date()),
  "",
  "POOLED AVERAGE MARGINAL EFFECTS:",
  paste(rep("-", 40), collapse = "")
)

for (sg_name in names(pooled_mfx)) {
  if (pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    summary_text <- c(summary_text,
      "",
      sg$name,
      sprintf("  AME: %.4f (SE = %.4f)", sg$ame, sg$se),
      sprintf("  95%% CI: (%.4f, %.4f)", sg$ci_lower, sg$ci_upper),
      sprintf("  P-value: %.4f", sg$p_value),
      sprintf("  Average N: %.0f", sg$avg_n)
    )
  }
}

# Save text summary
writeLines(summary_text, 
          file.path(marginal_effects_dir, "marginal_effects_summary.txt"))

log_message("Results exported")
cat("\n✓ Results exported to:", marginal_effects_dir, "\n")
```

## Session Info

```{r session-info}
log_message("Marginal effects analysis complete")
finalize_log(success = TRUE)
sessionInfo()
```