### 2025-09-08 reanalysis-13-marginal_effects.qmd
ame bug

---
title: "Step 13: Marginal Effects Analysis with Multiple Imputation"
subtitle: "Calculate average marginal effects for twang model across subgroups"
author: "Analysis Team"
date: "`r Sys.Date()`"
format:
  html:
    toc: true
    toc-depth: 3
    code-fold: true
    code-summary: "Show code"
    embed-resources: true
    self-contained: true
---

## Setup

```{r setup}
#| include: false
source("dependencies.R")
source("functions.R")
source("alt_specs.R")
setup_analysis(seed = 2025)

library(margins)
library(survey)
library(knitr)
library(kableExtra)

# Initialize logging
log_file <- init_log("marginal_effects")
log_message("Starting marginal effects analysis for twang models")

# Configuration
cache_dir <- file.path(reanalysis_data_dir, "ps_cache")
marginal_effects_dir <- file.path(reanalysis_data_dir, "marginal_effects")

# Create directory for marginal effects results
if (!dir.exists(marginal_effects_dir)) {
  dir.create(marginal_effects_dir, recursive = TRUE)
  log_message(paste0("Created marginal effects directory: ", marginal_effects_dir))
}

# Load imputed datasets
imputed_datasets <- readRDS(file.path(reanalysis_data_dir, "imputed_datasets.rds"))
n_imputations <- length(imputed_datasets)

# Load variable lists
variable_lists <- readRDS(file.path(reanalysis_data_dir, "variable_lists.rds"))

# Get outcome formula
outcome_formulas <- get_analysis_formulas_reanalysis()
outcome_formula <- outcome_formulas$full
```

## Define Subgroups

```{r define-subgroups}
# Define all subgroups from twang-analysis.Rmd
subgroups <- list(
  # Overall population
  overall = list(
    name = "Overall Population",
    subset_condition = NULL,  # No subsetting for overall
    description = "Average marginal effect across entire sample"
  ),

  # PDR subgroups - using dr_severity == "PDR"
  pdr_any_tx = list(
    name = "PDR, any treatment",
    subset_condition = expression(dr_severity == "PDR" & 
                                 (PRP_flag == 1 | anti_VEGF == 1 | 
                                  focal_laser_flag == 1 | other_inject == 1)),
    description = "Patients with PDR receiving any treatment"
  ),

  pdr_anti_vegf_only = list(
    name = "PDR, on anti-VEGF only",
    subset_condition = expression(dr_severity == "PDR" & 
                                 PRP_flag == 0 & anti_VEGF == 1 & 
                                 focal_laser_flag == 0 & other_inject == 0),
    description = "Patients with PDR receiving anti-VEGF only"
  ),

  pdr_prp_only = list(
    name = "PDR, on PRP only",
    subset_condition = expression(dr_severity == "PDR" & 
                                 PRP_flag == 1 & anti_VEGF == 0 & 
                                 focal_laser_flag == 0 & other_inject == 0),
    description = "Patients with PDR receiving PRP only"
  ),

  # No DR or NPDR combined groups - using dr_severity levels
  npdr_nodr_no_tx = list(
    name = "No DR or NPDR, no treatment",
    subset_condition = expression(dr_severity %in% c("No_DR", "NPDR") & 
                                 (PRP_flag == 0 & anti_VEGF == 0 & 
                                  focal_laser_flag == 0 & other_inject == 0)),
    description = "Patients with No DR or NPDR not on treatment"
  ),

  npdr_nodr_any_tx = list(
    name = "No DR or NPDR, any treatment",
    subset_condition = expression(dr_severity %in% c("No_DR", "NPDR") & 
                                 (PRP_flag == 1 | anti_VEGF == 1 | 
                                  focal_laser_flag == 1 | other_inject == 1)),
    description = "Patients with No DR or NPDR on any treatment"
  ),

  # NPDR overall groups - using dr_severity == "NPDR"
  npdr_no_tx = list(
    name = "NPDR, no treatment",
    subset_condition = expression(dr_severity == "NPDR" & 
                                 (PRP_flag == 0 & anti_VEGF == 0 & 
                                  focal_laser_flag == 0 & other_inject == 0)),
    description = "Patients with NPDR not on treatment"
  ),

  npdr_any_tx = list(
    name = "NPDR, any treatment",
    subset_condition = expression(dr_severity == "NPDR" & 
                                 (PRP_flag == 1 | anti_VEGF == 1 | 
                                  focal_laser_flag == 1 | other_inject == 1)),
    description = "Patients with NPDR on any treatment"
  )

  # Note: NPDR severity-specific subgroups (Mild/Moderate/Severe) removed
  # because dr_severity was imputed at coarse level (No_DR/NPDR/PDR only)
)

# Display subgroup definitions
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("SUBGROUPS FOR MARGINAL EFFECTS ANALYSIS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("- ", sg$name, "\n"))
  cat(paste0("  ", sg$description, "\n\n"))
}
```

## Test on Single Imputation

```{r test-single}
# Test the approach on a single imputation first
test_imp <- 1
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat(paste0("TESTING ON IMPUTATION ", test_imp, "\n"))
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Load twang weights for this imputation
twang_file <- file.path(cache_dir, paste0("twang_gbm_imp", test_imp, ".rds"))
if (!file.exists(twang_file)) {
  stop(paste0("twang results not found for imputation ", test_imp, 
              ". Please run reanalysis-4-propensity_scores.qmd first"))
}

twang_result <- readRDS(twang_file)
test_data <- imputed_datasets[[test_imp]]

# Add weights to data
test_data$twang_att_w <- twang_result$weights

# Fit the outcome model with weights
design <- svydesign(ids = ~e_mrn_deidentified, 
                   weights = ~twang_att_w, 
                   data = test_data)
model <- svyglm(formula = outcome_formula, 
               design = design, 
               family = quasibinomial())

# Calculate marginal effects for each subgroup
test_results <- list()

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("\nCalculating marginal effects for: ", sg$name, "\n"))

  if (is.null(sg$subset_condition)) {
    # Overall marginal effect
    mfx <- margins(model, design = design, variables = "ever_lapse_binary")
    test_results[[sg_name]] <- summary(mfx)
    cat(paste0("  N = ", nrow(test_data), "\n"))
  } else {
    # Subset the data and design
    subset_data <- subset(test_data, eval(sg$subset_condition))
    subset_design <- subset(design, eval(sg$subset_condition))

    if (nrow(subset_data) > 0) {
      mfx <- margins(model, 
                    design = subset_design, 
                    data = subset_data,
                    variables = "ever_lapse_binary")
      test_results[[sg_name]] <- summary(mfx)
      cat(paste0("  N = ", nrow(subset_data), "\n"))
    } else {
      cat("  WARNING: No observations in this subgroup\n")
      test_results[[sg_name]] <- NULL
    }
  }

  # Display results if available
  if (!is.null(test_results[[sg_name]])) {
    mfx_summary <- test_results[[sg_name]]

    # Debug output for first subgroup
    if (sg_name == names(subgroups)[1]) {
      cat("  Debug - margins output structure:\n")
      cat("    Rows:", paste(rownames(mfx_summary), collapse=", "), "\n")
      cat("    Cols:", paste(colnames(mfx_summary), collapse=", "), "\n")
      if ("factor" %in% colnames(mfx_summary)) {
        cat("    Factor column:", paste(mfx_summary$factor, collapse=", "), "\n")
      }
    }

    # Margins output has factor name in 'factor' column, not rownames
    # Look for ever_lapse_binary in the factor column
    if ("factor" %in% colnames(mfx_summary)) {
      lapse_rows <- grep("ever_lapse_binary", mfx_summary$factor)
      if (length(lapse_rows) > 0) {
        # Use the first matching row
        lapse_row <- lapse_rows[1]
        ame <- mfx_summary[lapse_row, "AME"]
        se <- mfx_summary[lapse_row, "SE"]
        cat(sprintf("  AME = %.4f (SE = %.4f)\n", ame, se))
      } else {
        cat("  WARNING: Could not find ever_lapse_binary in factor column\n")
      }
    } else {
      cat("  WARNING: No 'factor' column in margins output\n")
    }
  }
}

cat("\n✓ Test on single imputation successful\n")
```

## Calculate Marginal Effects for All Imputations

```{r calculate-all}
# Check if we have cached results to resume from
cached_results_file <- file.path(marginal_effects_dir, "all_mfx_results.rds")
if (file.exists(cached_results_file)) {
  cat("Loading cached marginal effects results...\n")
  all_mfx_results <- readRDS(cached_results_file)

  # Check which imputations are complete
  completed_imps <- c()
  for (i in 1:n_imputations) {
    if (length(all_mfx_results[[1]]) >= i && 
        !is.null(all_mfx_results[[1]][[i]])) {
      completed_imps <- c(completed_imps, i)
    }
  }

  if (length(completed_imps) > 0) {
    cat(paste0("Found ", length(completed_imps), " completed imputations\n"))
    cat(paste0("Will resume from imputation ", max(completed_imps) + 1, "\n\n"))
  }
} else {
  # Initialize new storage
  all_mfx_results <- list()

  # Initialize storage for each subgroup
  for (sg_name in names(subgroups)) {
    all_mfx_results[[sg_name]] <- list()
  }
  completed_imps <- c()
}

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("CALCULATING MARGINAL EFFECTS ACROSS ALL IMPUTATIONS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

# Process each imputation
for (i in 1:n_imputations) {
  # Skip if already completed
  if (i %in% completed_imps) {
    next
  }

  if (i %% 10 == 1) {
    cat(paste0("Processing imputations ", i, "-", min(i+9, n_imputations), "...\n"))
  }

  # Load twang weights
  twang_file <- file.path(cache_dir, paste0("twang_gbm_imp", i, ".rds"))
  if (!file.exists(twang_file)) {
    cat(paste0("  WARNING: Skipping imputation ", i, " - twang results not found\n"))
    next
  }

  twang_result <- readRDS(twang_file)
  imp_data <- imputed_datasets[[i]]
  imp_data$twang_att_w <- twang_result$weights

  # Create survey design and fit model
  design <- svydesign(ids = ~e_mrn_deidentified, 
                     weights = ~twang_att_w, 
                     data = imp_data)

  tryCatch({
    model <- svyglm(formula = outcome_formula, 
                   design = design, 
                   family = quasibinomial())

    # Calculate marginal effects for each subgroup
    for (sg_name in names(subgroups)) {
      sg <- subgroups[[sg_name]]

      tryCatch({
        if (is.null(sg$subset_condition)) {
          # Overall marginal effect
          mfx <- margins(model, design = design, variables = "ever_lapse_binary")
          mfx_summary <- summary(mfx)
        } else {
          # Subset marginal effect
          subset_data <- subset(imp_data, eval(sg$subset_condition))

          if (nrow(subset_data) > 0) {
            subset_design <- subset(design, eval(sg$subset_condition))
            mfx <- margins(model, 
                          design = subset_design, 
                          data = subset_data,
                          variables = "ever_lapse_binary")
            mfx_summary <- summary(mfx)
          } else {
            mfx_summary <- NULL
          }
        }

        # Store results - margins output has factor name in 'factor' column
        if (!is.null(mfx_summary)) {
          # Debug: show what we got for first imp and first subgroup
          if (i == 1 && sg_name == names(subgroups)[1]) {
            cat("  Debug: mfx_summary structure:\n")
            cat("    Cols:", paste(colnames(mfx_summary), collapse=", "), "\n")
            if ("factor" %in% colnames(mfx_summary)) {
              cat("    Factor values:", paste(mfx_summary$factor, collapse=", "), "\n")
            }
          }

          # Look for ever_lapse_binary in the factor column
          if ("factor" %in% colnames(mfx_summary)) {
            lapse_rows <- grep("ever_lapse_binary", mfx_summary$factor)
            if (length(lapse_rows) > 0) {
              # Use the first matching row (should only be one)
              lapse_row <- lapse_rows[1]
              all_mfx_results[[sg_name]][[i]] <- list(
                ame = mfx_summary[lapse_row, "AME"],
                se = mfx_summary[lapse_row, "SE"],
                n = ifelse(is.null(sg$subset_condition), 
                          nrow(imp_data),
                          nrow(subset(imp_data, eval(sg$subset_condition)))),
                success = TRUE
              )
            } else {
              all_mfx_results[[sg_name]][[i]] <- list(
                success = FALSE,
                error = "Could not find ever_lapse_binary in factor column"
              )
            }
          } else {
            # Fallback: if no factor column, try using first row
            if (nrow(mfx_summary) == 1) {
              all_mfx_results[[sg_name]][[i]] <- list(
                ame = mfx_summary[1, "AME"],
                se = mfx_summary[1, "SE"],
                n = ifelse(is.null(sg$subset_condition), 
                          nrow(imp_data),
                          nrow(subset(imp_data, eval(sg$subset_condition)))),
                success = TRUE
              )
            } else {
              all_mfx_results[[sg_name]][[i]] <- list(
                success = FALSE,
                error = "No 'factor' column and multiple rows in margins output"
              )
            }
          }
        } else {
          all_mfx_results[[sg_name]][[i]] <- list(
            success = FALSE,
            error = "No data in subset"
          )
        }
      }, error = function(e) {
        all_mfx_results[[sg_name]][[i]] <- list(
          success = FALSE,
          error = e$message
        )
        if (i == 1) {
          cat(paste0("    ERROR in ", sg_name, ": ", e$message, "\n"))
        }
      })
    }

  }, error = function(e) {
    cat(paste0("  ERROR in imputation ", i, ": ", e$message, "\n"))
    for (sg_name in names(subgroups)) {
      all_mfx_results[[sg_name]][[i]] <- list(success = FALSE, error = e$message)
    }
  })

  # Save intermediate results every 5 imputations
  if (i %% 5 == 0) {
    saveRDS(all_mfx_results, cached_results_file)
    cat(paste0("  [Checkpoint] Saved results through imputation ", i, "\n"))
  }

  # Periodic garbage collection
  if (i %% 10 == 0) {
    gc(verbose = FALSE)
  }
}

cat("\n✓ Marginal effects calculation complete\n")

# Save final results
saveRDS(all_mfx_results, cached_results_file)
log_message("Saved all marginal effects results")
```

## Pool Marginal Effects Using Rubin's Rules

```{r pool-mfx}
# Check if we have pooled results cached
pooled_results_file <- file.path(marginal_effects_dir, "pooled_mfx_results.rds")

if (file.exists(pooled_results_file)) {
  cat("Loading cached pooled results...\n")
  pooled_mfx <- readRDS(pooled_results_file)
} else {
  # Pool the marginal effects across imputations
  pooled_mfx <- list()
}

cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("POOLING MARGINAL EFFECTS USING RUBIN'S RULES\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

for (sg_name in names(subgroups)) {
  sg <- subgroups[[sg_name]]
  cat(paste0("\n", sg$name, ":\n"))

  # Extract successful results
  sg_results <- all_mfx_results[[sg_name]]
  valid_results <- sg_results[sapply(sg_results, function(x) 
    !is.null(x) && x$success)]

  if (length(valid_results) < 2) {
    cat("  Insufficient valid imputations for pooling\n")
    pooled_mfx[[sg_name]] <- list(success = FALSE)
    next
  }

  # Extract AMEs and SEs
  ames <- sapply(valid_results, function(x) x$ame)
  ses <- sapply(valid_results, function(x) x$se)
  ns <- sapply(valid_results, function(x) x$n)

  # Apply Rubin's rules
  m <- length(ames)
  q_bar <- mean(ames)         # Combined estimate
  u_bar <- mean(ses^2)        # Within-imputation variance
  b <- var(ames)              # Between-imputation variance
  t <- u_bar + (1 + 1/m) * b  # Total variance
  se_pooled <- sqrt(t)

  # Degrees of freedom
  gamma <- ((1 + 1/m) * b) / t
  df <- (m - 1) / gamma^2

  # Confidence intervals and p-value
  t_crit <- qt(0.975, df)
  ci_lower <- q_bar - t_crit * se_pooled
  ci_upper <- q_bar + t_crit * se_pooled
  z_stat <- q_bar / se_pooled
  p_value <- 2 * pnorm(-abs(z_stat))

  # Average sample size
  avg_n <- mean(ns)

  # Store pooled results
  pooled_mfx[[sg_name]] <- list(
    name = sg$name,
    ame = q_bar,
    se = se_pooled,
    ci_lower = ci_lower,
    ci_upper = ci_upper,
    z_stat = z_stat,
    p_value = p_value,
    n_imputations = m,
    avg_n = avg_n,
    df = df,
    success = TRUE
  )

  # Display results
  cat(sprintf("  AME = %.4f (SE = %.4f)\n", q_bar, se_pooled))
  cat(sprintf("  95%% CI: (%.4f, %.4f)\n", ci_lower, ci_upper))
  cat(sprintf("  P-value = %.4f\n", p_value))
  cat(sprintf("  N imputations = %d\n", m))
  cat(sprintf("  Average N = %.0f\n", avg_n))
}

# Save pooled results if not already cached
if (!file.exists(pooled_results_file)) {
  saveRDS(pooled_mfx, pooled_results_file)
  log_message("Saved pooled marginal effects")
} else {
  log_message("Using cached pooled marginal effects")
}
```

## Summary Table

```{r summary-table}
# Create summary table
if (length(pooled_mfx) > 0) {
  summary_rows <- lapply(names(pooled_mfx), function(sg_name) {
    res <- pooled_mfx[[sg_name]]
    if (!is.null(res) && !is.null(res$success) && res$success) {
      data.frame(
        Subgroup = res$name,
        AME = sprintf("%.4f", res$ame),
        SE = sprintf("%.4f", res$se),
        `95% CI` = sprintf("(%.4f, %.4f)", res$ci_lower, res$ci_upper),
        `P-value` = ifelse(res$p_value < 0.001, "<0.001", 
                           sprintf("%.3f", res$p_value)),
        `Avg N` = sprintf("%.0f", res$avg_n),
        `N Imp` = res$n_imputations,
        stringsAsFactors = FALSE,
        check.names = FALSE
      )
    } else {
      NULL
    }
  })

  # Remove NULL entries and combine
  summary_rows <- summary_rows[!sapply(summary_rows, is.null)]

  if (length(summary_rows) > 0) {
    summary_df <- do.call(rbind, summary_rows)

    # Display table - simplified to avoid kable errors
    print(kable(summary_df,
          caption = "Pooled Average Marginal Effects of Lapsing by Subgroup",
          align = c("l", "r", "r", "c", "r", "r", "r"),
          digits = 4))

    # Save to CSV
    write.csv(summary_df, 
              file.path(marginal_effects_dir, "marginal_effects_summary.csv"),
              row.names = FALSE)
    log_message("Saved summary table to CSV")
  } else {
    cat("No successful results to display in summary table\n")
    log_message("No successful pooled results for summary table")
  }
} else {
  cat("No pooled results available for summary table\n")
  log_message("No pooled results found")
}
```

## Interpretation

```{r interpretation}
cat("\n", paste(rep("=", 60), collapse = ""), "\n")
cat("INTERPRETATION OF RESULTS\n")
cat(paste(rep("=", 60), collapse = ""), "\n\n")

cat("Average Marginal Effect (AME) represents the average change in\n")
cat("probability of vision impairment/blindness associated with lapsing\n")
cat("from care, holding all other variables at their observed values.\n\n")

cat("Key findings:\n\n")

# Overall effect
if ("overall" %in% names(pooled_mfx) && pooled_mfx$overall$success) {
  overall_ame <- pooled_mfx$overall$ame
  overall_pct <- overall_ame * 100
  cat(sprintf("1. Overall: Lapsing is associated with a %.1f percentage point\n", overall_pct))
  cat(sprintf("   increase in probability of vision impairment (AME = %.4f)\n\n", overall_ame))
}

# PDR subgroups
pdr_groups <- c("pdr_any_tx", "pdr_anti_vegf_only", "pdr_prp_only")
for (sg_name in pdr_groups) {
  if (sg_name %in% names(pooled_mfx) && pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    pct_change <- sg$ame * 100
    cat(sprintf("2. %s:\n", sg$name))
    cat(sprintf("   AME = %.4f (%.1f percentage points)\n", sg$ame, pct_change))
    cat(sprintf("   Average N = %.0f patients\n\n", sg$avg_n))
  }
}

# Statistical significance
cat("Statistical significance (p < 0.05):\n")
for (sg_name in names(pooled_mfx)) {
  if (pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    sig <- ifelse(sg$p_value < 0.05, "Yes", "No")
    cat(sprintf("  - %s: %s (p = %.4f)\n", sg$name, sig, sg$p_value))
  }
}
```

## Forest Plots

```{r forest-plots}
#| fig.width: 10
#| fig.height: 8

# Prepare data for forest plot
forest_data <- do.call(rbind, lapply(names(pooled_mfx), function(sg_name) {
  res <- pooled_mfx[[sg_name]]
  if (res$success) {
    data.frame(
      label = res$name,
      AME = res$ame,
      SE = res$se,
      Lower = res$ci_lower,
      Upper = res$ci_upper,
      p = res$p_value,
      n = res$avg_n,
      group_type = ifelse(sg_name == "overall", "overall", "subgroup"),
      stringsAsFactors = FALSE
    )
  }
}))

# Add sample sizes to labels
forest_data$label_with_n <- paste0(forest_data$label, " (n=", round(forest_data$n), ")")

# Define order for plots (reversed for coord_flip)
subgroup_order <- rev(c(
  "No DR or NPDR, no treatment", "No DR or NPDR, any treatment", 
  "NPDR, no treatment", "NPDR, any treatment", 
  "Mild NPDR, no treatment", "Mild NPDR, any treatment",
  "Moderate NPDR, no treatment", "Moderate NPDR, any treatment",
  "Severe NPDR, no treatment", "Severe NPDR, any treatment",
  "PDR, any treatment", "PDR, on anti-VEGF only", "PDR, on PRP only"
))

# Version 1: Subgroups only
data_subgroups <- forest_data[forest_data$group_type == "subgroup", ]
data_subgroups$Labels <- factor(data_subgroups$label, levels = subgroup_order)

plot_subgroups <- ggplot(data_subgroups, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  geom_point(size = 3, color = "#2563eb") +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper), height = 0.2, color = "#2563eb") +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.01, max(forest_data$Upper) * 1.1),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "Average Marginal Effects of Lapse Across Selected Groups",
    subtitle = "Pooled across multiple imputations using Rubin's rules",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("Based on ", unique(forest_data$n_imputations)[1], " imputations")
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = ggplot2::margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = ggplot2::margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = ggplot2::margin(b = 20)),
    plot.subtitle = element_text(size = 11, hjust = 0.5, margin = ggplot2::margin(b = 10)),
    plot.margin = ggplot2::margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Version 2: With overall population included
all_order <- c("Overall Population", subgroup_order)
forest_data$Labels <- factor(forest_data$label, levels = rev(all_order))
forest_data$point_color <- ifelse(forest_data$group_type == "overall", "#dc2626", "#2563eb")
forest_data$point_size <- ifelse(forest_data$group_type == "overall", 4, 3)

plot_with_overall <- ggplot(forest_data, aes(x = AME, y = Labels)) +
  geom_vline(xintercept = 0, color = "#be123c", linetype = "solid", size = 0.5) +
  geom_hline(yintercept = length(subgroup_order) + 0.5, 
             color = "gray50", linetype = "dashed", alpha = 0.5) +
  geom_point(aes(color = point_color, size = point_size)) +
  geom_errorbarh(aes(xmin = Lower, xmax = Upper, color = point_color), height = 0.2) +
  geom_text(
    aes(x = Lower, label = sprintf("%.3f", Lower)), 
    hjust = 1.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(x = Upper, label = sprintf("%.3f", Upper)), 
    hjust = -0.2, size = 3, color = "black"
  ) +
  geom_text(
    aes(label = sprintf("%.3f", AME)), 
    hjust = -0.2, vjust = -1, size = 3, color = "black"
  ) +
  scale_color_identity() +
  scale_size_identity() +
  theme_minimal() +
  scale_x_continuous(
    limits = c(-0.01, max(forest_data$Upper) * 1.1),
    breaks = scales::pretty_breaks(n = 6)
  ) +
  labs(
    title = "Average Marginal Effects of Lapse Across All Groups",
    subtitle = "Pooled across multiple imputations using Rubin's rules",
    x = "Average Marginal Effect (AME)",
    y = "Treatment Group",
    caption = paste0("Based on ", unique(forest_data$n_imputations)[1], " imputations\n",
                    "Overall population shown in red, subgroups in blue")
  ) +
  theme(
    axis.text.x = element_text(size = 10),
    axis.text.y = element_text(size = 10), 
    axis.title.y = element_text(size = 12, face = "bold", margin = ggplot2::margin(r = 20)),
    axis.title.x = element_text(size = 12, face = "bold", margin = ggplot2::margin(t = 20)),
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5, margin = ggplot2::margin(b = 20)),
    plot.subtitle = element_text(size = 11, hjust = 0.5, margin = ggplot2::margin(b = 10)),
    plot.margin = ggplot2::margin(10, 20, 30, 20),
    panel.grid.major.y = element_line(color = "gray", linetype = "dotted"),
    plot.caption = element_text(size = 10, face = "italic")
  )

# Display plots
print(plot_subgroups)
print(plot_with_overall)

# Save plots
ggsave(file.path(marginal_effects_dir, "forest_plot_subgroups_only.png"), 
       plot_subgroups, width = 10, height = 8, dpi = 300, bg = "white")
ggsave(file.path(marginal_effects_dir, "forest_plot_with_overall.png"), 
       plot_with_overall, width = 10, height = 8.5, dpi = 300, bg = "white")

log_message("Forest plots saved")
```

## Export Results

```{r export}
# Create comprehensive export
export_list <- list(
  pooled_results = pooled_mfx,
  summary_table = summary_df,
  metadata = list(
    n_imputations = n_imputations,
    subgroups = names(subgroups),
    method = "twang_gbm",
    formula = deparse(outcome_formula),
    date = Sys.Date()
  )
)

# Save as RDS for R users
saveRDS(export_list, 
        file.path(marginal_effects_dir, "marginal_effects_export.rds"))

# Create text summary for sharing
summary_text <- c(
  "MARGINAL EFFECTS ANALYSIS RESULTS",
  paste(rep("=", 60), collapse = ""),
  paste0("Method: twang GBM with IPW"),
  paste0("Number of imputations: ", n_imputations),
  paste0("Date: ", Sys.Date()),
  "",
  "POOLED AVERAGE MARGINAL EFFECTS:",
  paste(rep("-", 40), collapse = "")
)

for (sg_name in names(pooled_mfx)) {
  if (pooled_mfx[[sg_name]]$success) {
    sg <- pooled_mfx[[sg_name]]
    summary_text <- c(summary_text,
      "",
      sg$name,
      sprintf("  AME: %.4f (SE = %.4f)", sg$ame, sg$se),
      sprintf("  95%% CI: (%.4f, %.4f)", sg$ci_lower, sg$ci_upper),
      sprintf("  P-value: %.4f", sg$p_value),
      sprintf("  Average N: %.0f", sg$avg_n)
    )
  }
}

# Save text summary
writeLines(summary_text, 
          file.path(marginal_effects_dir, "marginal_effects_summary.txt"))

log_message("Results exported")
cat("\n✓ Results exported to:", marginal_effects_dir, "\n")
```

## Session Info

```{r session-info}
log_message("Marginal effects analysis complete")
finalize_log(success = TRUE)
sessionInfo()
```